### 2016-09-23 manuscript/tracer16.tex
initial import of manuscript

\documentclass{bioinfo}
\copyrightyear{2012}
\pubyear{2012}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

\usepackage{graphicx}
\usepackage{subfigure} 
% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{url}

\begin{document}
\firstpage{1}

% Title must be 150 characters or less
\title[Tracer 1.6]{Bayesian MCMC diagnostics and summarization using Tracer 1.6}

\author[Rambaut \textit{et~al}]{ Andrew Rambaut\,$^{1}$, Alexei J.~Drummond\,$^{2,3}$, Dong Xie\,$^{2,3}$, Marc A.~Suchard\,$^{4,5,6}$}

\address{
$^{1}$Institute of Evolutionary Biology, University of Edinburgh, Edinburgh, UK
$^{2}$Department of Computer Science, University of Auckland, Auckland, NZ\\
$^{3}$Allan Wilson Centre for Molecular Ecology and Evolution, University of Auckland, Auckland, NZ\\
$^{4,5}$Departments of Biomathematics and Human Genetics, David Geffen School of Medicine at UCLA, and \\
$^{6}$Department of Biostatistics, UCLA Fielding School of Public Health, University of California, Los Angeles, USA \\
}

\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\maketitle

% Please keep the abstract between 250 and 300 words
\begin{abstract}

\section{Motivation:}
Computational evolutionary biology, statistical phylogenetics and coalescent-based population genetics are becoming increasingly central to the analysis and understanding of molecular sequence data. 
\section{Results:}
We describe here a software package, Tracer version 1.6, which ... 
\section{Availability:}
Tracer is open-source under the GNU lesser general public license and available at \url{http://beast-mcmc.googlecode.com}  and  \href{http://tree.ed.ac.uk/software/tracer}{\url{http://tree.ed.ac.uk}}.

\section{Contact:} 
\href{a.rambaut@ed.ac.uk}{\url{a.rambaut@ed.ac.uk}}, 
\href{alexei@cs.auckland.ac.nz}{\url{alexei@cs.auckland.ac.nz}}
and 
\href{msuchard@ucla.edu}{\url{msuchard@ucla.edu}}

\end{abstract}

\section*{Introduction}

Blah blah

\begin{itemize}
\item Importance of posterior summaries (as scientifically relevant marginalization of the full posterior distribution)
\item Full posterior distribution is high-dimensional and (often) difficult to visualize
\item Posterior simulator agnostic -- highly used with \textsc{MrBayes} as well
\end{itemize}

\section*{Design and Implementation}

Blah blah

New features:
\begin{itemize}
\item Kernel density estimates
\item HME
\item AICM
\end{itemize}

\subsubsection*{Graphical user-interface:} Blah blah

\subsubsection*{Diagnostics:} Blah blah

\subsubsection*{Demographic reconstruction:}

\subsubsection*{Model selection:}

\section*{Example}

Even users of \textsc{MrBayes} find Tracer handy.

\section*{Availability and Future Directions}

We make the Tracer package available in both executable and source code forms.   Tracer requires Java version 1.5 or greater and executables for Windows, Mac OS and Linux platforms are located at \url{http://beast.bio.ed.ac.uk} which serves as the main page for the package. This page also links to a sizable list of self-contained, step-by-step tutorials covering basic to advance usage of Tracer to summarize posterior distributions of a large set of phylogenetic models simulated using BEAST.  For example, popular tutorials describe how to use Tracer to generate marginal parameter summaries and infer population dynamics trajectories over time.

GoogleCode houses the Tracer's version-controlled source code within \url{http://beast-mcmc.googlecode.com} and links to two GoogleGroup discussion groups related to Tracer.  
The first is the ``beast-users" group (\url{http://groups.google.com/group/beast-users}) with over 1,500 members. 
%At the time of writing, forty-seven developers belong to the ``beast-dev" group that facilitates BEAST development across three continents.

Future development directions for Tracer focus on \ldots

Feature list:
\begin{itemize}
\item KDEs
\item cheap marginal likelihood estimators
\item demographic trajectory reconstruction
\end{itemize}

\section*{Taken from the Tracer website}

Tracer is a program for analysing the trace files generated by Bayesian MCMC runs (that is, the continuous parameter values sampled from the chain). It can be used to analyse runs of BEAST, MrBayes, LAMARC and possibly other MCMC programs.

Although Tracer can be used with programs other than BEAST, users are strongly advised to join the BEAST mailing-list. This will be used to announce new versions and advise users about bugs and problems.

You can join the mailing list here:
http://groups.google.com/group/beast-users

The website for BEAST (and Tracer) is here:
http://beast.bio.ed.ac.uk/

At present there is no detailed manual for this application, you will simply have to play around and see what happens. Basically you can select the trace file in the top left of the window, the individual parameter in the bottom left and the analysis appears on the right.

There are 4 analysis tabs to choose from:

Estimates - this shows the mean, stdev, confidence intervals and other statistics about the selected parameter. A frequency distribution will also be plotted.
Density - this shows the Bayesian posterior density plot for the selected parameter.
Joint-Marginal - this only appears if exactly 2 parameters are choosen (hold down shift to select multiple parameters). It then plots one against the other to look at their joint-marginal distribution.
Trace - this shows the trace of the parameter against state or generation number. Use this to check mixing, choose a suitable burn-in and look for trends that might suggest problems with convergence.
Multiple parameters can be selected by holding down the shift key. This will overlay the plots for the different parameters allowing comparisons to be made. You can also select multiple trace files as well to compare different runs. If multiple trace files have the same trace names then a "Combined" trace will automatically appear. This can be selected as well as the individual trace files.

You can also select the "Demographic Analysis" from the Analysis menu - This plots the distribution of demographic population sizes over time for a number of models (constant size, exponential growth \& logistic growth) that are available in BEAST. This involves you selecting the traces for each parameter of the model. You should only select the model that was actually run under BEAST (e.g., if you ran an exponential growth model, you shouldn't plot the constant population size model).

The "Analysis" menu also contains options for performing Bayesian Skyline reconstructions and for calculating Bayes Factors between runs.

The "Print" function in the "File" menu will print the current graph or table and the "Export Data" function can be used to export the data from the plots for use in another graphic package.

To export the currently displayed graphic use the "Export PDF" function in the "File" menu.

\section*{Acknowledgments}

We thank the National Evolutionary Synthesis Center (NESCent) for sponsoring a working group (Software for Bayesian Evolutionary Analysis) that facilitated the development of Tracer version 1.6. 
This work was supported in part by funding from the Marsden Trust, NSF (DMS 0856099), NIH (R01 GM086887, R01 HG006139), The Royal Society of London, BBSRC (BB/H011285/1) and the Wellcome Trust (WT092807MA).

%\section*{References}
% The bibtex filename
\bibliographystyle{natbib}
\bibliography{tracer16}

\end{document}

### 2016-09-26 manuscript/tracer16.tex
small changes and addition of bib

@@
 \address{
-$^{1}$Institute of Evolutionary Biology, University of Edinburgh, Edinburgh, UK
+$^{1}$Institute of Evolutionary Biology, University of Edinburgh, Edinburgh, UK\\
 $^{2}$Department of Computer Science, University of Auckland, Auckland, NZ\\
-$^{3}$Allan Wilson Centre for Molecular Ecology and Evolution, University of Auckland, Auckland, NZ\\
+$^{3}$Centre for Computational Evolution, University of Auckland, Auckland, NZ\\
 $^{4,5}$Departments of Biomathematics and Human Genetics, David Geffen School of Medicine at UCLA, and \\
@@
 
-Even users of \textsc{MrBayes} find Tracer handy.
+Even users of \textsc{MrBayes} and \textsc{RevBayes} find Tracer handy.
 
@@
 Tracer is a program for analysing the trace files generated by Bayesian MCMC runs (that is, the continuous parameter values sampled from the chain).
-It can be used to analyse runs of BEAST, MrBayes, LAMARC and possibly other MCMC programs.
+It can be used to analyse runs of BEAST \citep{drummond2007beast,drummond2012bayesian}, BEAST2 \citep{bouckaert2014beast2}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes}, LAMARC \citep{kuhner2006lamarc}, Migrate \citep{beerli2006comparison} and possibly other MCMC programs.
 
-Although Tracer can be used with programs other than BEAST, users are strongly advised to join the BEAST mailing-list.
-This will be used to announce new versions and advise users about bugs and problems.
+Although Tracer can be used with programs other than BEAST, users may find it useful to join the BEAST users mailing list.
+This is used to announce new versions and advise users about bugs and problems.
 
@@
 Density - this shows the Bayesian posterior density plot for the selected parameter.
-Joint-Marginal - this only appears if exactly 2 parameters are choosen (hold down shift to select multiple parameters).
-It then plots one against the other to look at their joint-marginal distribution.
+Joint-Marginal - this only appears if exactly 2 parameters are chosen (hold down shift to select multiple parameters). It then plots one against the other to look at their joint-marginal distribution.
 Trace - this shows the trace of the parameter against state or generation number. Use this to check mixing, choose a suitable burn-in and look for trends that might suggest problems with convergence.

### 2016-10-03 manuscript/tracer16.tex
add a subsection and few figures

@@
 
-\subsubsection*{Graphical user-interface:} Blah blah
+\subsubsection*{Graphical user-interface:}
+
+Tracer can analyse either a single MCMC log or combine the shared traces from multiple logs.
+Traces panel provides the list of the statistical summary from the sampling parameters' values at all states logged during a MCMC process.
+
+There are 4 analysis tabs to choose from a selected parameter:
+
+%TODO
+Estimates - this shows the mean, stdev, confidence intervals and other statistics about the selected parameter. A frequency distribution will also be plotted.
+Density - this shows the Bayesian posterior density plot for the selected parameter.
+Joint-Marginal - this only appears if exactly 2 parameters are chosen (hold down shift to select multiple parameters). It then plots one against the other to look at their joint-marginal distribution.
+Trace - this shows the trace of the parameter against state or generation number. Use this to check mixing, choose a suitable burn-in and look for trends that might suggest problems with convergence.
+Multiple parameters can be selected by holding down the shift key.
+This will overlay the plots for the different parameters allowing comparisons to be made.
+You can also select multiple trace files as well to compare different runs.
+If multiple trace files have the same trace names then a "Combined" trace will automatically appear.
+This can be selected as well as the individual trace files.
+
+%TODO better figures
+
+\begin{figure}[H]
+\includegraphics[width=.5\textwidth]{../figures/gopher-log.png}  
+\caption{The }
+\label{fig:front:page}
+\end{figure}
+
+\begin{figure}[H]
+\includegraphics[width=.5\textwidth]{../figures/estimates.pdf}  
+\caption{The estimates}
+\label{fig:estimates}
+\end{figure}
+
+\begin{figure}[H]
+\includegraphics[width=.5\textwidth]{../figures/comp-95HPD.pdf}  
+\caption{The comparison of  95\% HPD intervals from multi-trace}
+\label{fig:comp95HPD}
+\end{figure}
+
+\begin{figure}[H]
+\includegraphics[width=.5\textwidth]{../figures/trace.pdf}  
+\caption{The trace}
+\label{fig:trace}
+\end{figure}
 
@@
 \subsubsection*{Model selection:}
+
+\subsubsection*{Conditional posterior distribution:}
 

### 2017-02-03 src/beast/tracer/analysis/BayesianSkylineDialog.java
working Tracer imported from BEAST1 after commit 20fe946d852c2f0375a1aa41309b521a933d64d1

@@ -1,0 +1,887 @@
+/*
+ * BayesianSkylineDialog.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.analysis;
+
+import dr.app.gui.components.RealNumberField;
+import dr.app.gui.components.WholeNumberField;
+import dr.app.gui.util.LongTask;
+import beast.tracer.trace.TraceDistribution;
+import beast.tracer.trace.TraceList;
+import dr.stats.Variate;
+import jam.framework.DocumentFrame;
+import jam.panels.OptionsPanel;
+import jebl.evolution.coalescent.IntervalList;
+import jebl.evolution.coalescent.Intervals;
+import jebl.evolution.io.ImportException;
+import jebl.evolution.io.NewickImporter;
+import jebl.evolution.io.NexusImporter;
+import jebl.evolution.io.TreeImporter;
+import jebl.evolution.trees.RootedTree;
+
+import javax.swing.*;
+import javax.swing.border.EmptyBorder;
+import javax.swing.event.ChangeEvent;
+import javax.swing.event.ChangeListener;
+import java.awt.*;
+import java.awt.event.ActionEvent;
+import java.awt.event.ActionListener;
+import java.awt.event.ItemEvent;
+import java.awt.event.ItemListener;
+import java.io.BufferedReader;
+import java.io.File;
+import java.io.FileReader;
+import java.util.ArrayList;
+import java.util.HashSet;
[…]

### 2017-02-03 src/beast/tracer/analysis/DemographicDialog.java

@@ -1,0 +1,772 @@
+/*
+ * DemographicDialog.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.analysis;
+
+import dr.app.gui.components.RealNumberField;
+import dr.app.gui.components.WholeNumberField;
+import dr.app.gui.util.LongTask;
+import dr.evolution.coalescent.ConstExpConst;
+import dr.evolution.coalescent.TwoEpochDemographic;
+import dr.evolution.util.Units;
+import beast.tracer.trace.TraceDistribution;
+import beast.tracer.trace.TraceList;
+import dr.stats.Variate;
+import jam.framework.DocumentFrame;
+import jam.panels.OptionsPanel;
+import jebl.evolution.coalescent.*;
+
+import javax.swing.*;
+import javax.swing.border.EmptyBorder;
+import javax.swing.event.ChangeEvent;
+import javax.swing.event.ChangeListener;
+import java.awt.event.ActionEvent;
+import java.awt.event.ActionListener;
+import java.awt.event.ItemEvent;
+import java.awt.event.ItemListener;
+import java.util.ArrayList;
+import java.util.List;
+
+public class DemographicDialog {
+
+    private JFrame frame;
+
+    private JComboBox demographicCombo;
+    private WholeNumberField binCountField;
+
+    public static String[] demographicModels = {
+            "Constant Population",
+            "Exponential Growth (Growth Rate)",
+            "Exponential Growth (Doubling Time)",
+            "Logistic Growth (Growth Rate)",
[…]

### 2017-02-03 src/beast/tracer/analysis/ExtendedBayesianSkylineDialog.java

@@ -1,0 +1,877 @@
+/*
+ * ExtendedBayesianSkylineDialog.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.analysis;
+
+import dr.app.gui.components.RealNumberField;
+import dr.app.gui.components.WholeNumberField;
+import dr.app.gui.util.LongTask;
+import beast.tracer.trace.TraceDistribution;
+import beast.tracer.trace.TraceList;
+import dr.stats.Variate;
+import jam.framework.DocumentFrame;
+import jam.panels.OptionsPanel;
+import jebl.evolution.coalescent.IntervalList;
+import jebl.evolution.coalescent.Intervals;
+import jebl.evolution.io.ImportException;
+import jebl.evolution.io.NewickImporter;
+import jebl.evolution.io.NexusImporter;
+import jebl.evolution.io.TreeImporter;
+import jebl.evolution.trees.RootedTree;
+
+import javax.swing.*;
+import javax.swing.border.EmptyBorder;
+import javax.swing.event.ChangeEvent;
+import javax.swing.event.ChangeListener;
+import java.awt.*;
+import java.awt.event.ActionEvent;
+import java.awt.event.ActionListener;
+import java.awt.event.ItemEvent;
+import java.awt.event.ItemListener;
+import java.io.BufferedReader;
+import java.io.File;
+import java.io.FileReader;
+import java.util.ArrayList;
+import java.util.HashSet;
[…]

### 2017-02-03 src/beast/tracer/analysis/LineagesThroughTimeDialog.java

@@ -1,0 +1,565 @@
+/*
+ * LineagesThroughTimeDialog.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.analysis;
+
+import dr.app.gui.components.RealNumberField;
+import dr.app.gui.components.WholeNumberField;
+import dr.app.gui.util.LongTask;
+import beast.tracer.trace.TraceDistribution;
+import beast.tracer.trace.TraceList;
+import dr.stats.Variate;
+import jam.framework.DocumentFrame;
+import jam.panels.OptionsPanel;
+import jebl.evolution.graphs.Node;
+import jebl.evolution.io.ImportException;
+import jebl.evolution.io.NewickImporter;
+import jebl.evolution.io.NexusImporter;
+import jebl.evolution.io.TreeImporter;
+import jebl.evolution.trees.RootedTree;
+
+import javax.swing.*;
+import javax.swing.border.EmptyBorder;
+import javax.swing.event.ChangeEvent;
+import javax.swing.event.ChangeListener;
+import java.awt.*;
+import java.awt.event.ActionEvent;
+import java.awt.event.ActionListener;
+import java.io.BufferedReader;
+import java.io.File;
+import java.io.FileReader;
+import java.util.Arrays;
+import java.util.List;
+
+public class LineagesThroughTimeDialog {
+
+    private JFrame frame;
+
+    private final JButton button = new JButton("Choose File...");
+    private ActionListener buttonListener;
+
[…]

### 2017-02-03 src/beast/tracer/analysis/NewTemporalAnalysisDialog.java

@@ -1,0 +1,125 @@
+/*
+ * NewTemporalAnalysisDialog.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.analysis;
+
+import dr.app.gui.components.RealNumberField;
+import dr.app.gui.components.WholeNumberField;
+import jam.framework.DocumentFrame;
+import jam.panels.OptionsPanel;
+
+import javax.swing.*;
+import javax.swing.border.EmptyBorder;
+
+public class NewTemporalAnalysisDialog {
+
+    private JFrame frame;
+
+    private JTextField titleField;
+
+    private WholeNumberField binCountField;
+
+    private RealNumberField minTimeField;
+    private RealNumberField maxTimeField;
+
+    private OptionsPanel optionPanel;
+
+    public NewTemporalAnalysisDialog(JFrame frame) {
+        this.frame = frame;
+
+        titleField = new JTextField();
+        titleField.setColumns(32);
+
+        binCountField = new WholeNumberField(2, 2000);
+        binCountField.setValue(100);
+        binCountField.setColumns(4);
+
+        maxTimeField = new RealNumberField(0.0, Double.MAX_VALUE);
+        maxTimeField.setColumns(12);
+
+        minTimeField = new RealNumberField(0.0, Double.MAX_VALUE);
+        minTimeField.setColumns(12);
+
+        optionPanel = new OptionsPanel(12, 12);
+    }
+
+    public int showDialog() {
+
+        setArguments();
+
[…]

### 2017-02-03 src/beast/tracer/analysis/SkyGridDialog.java

@@ -1,0 +1,645 @@
+/*
+ * SkyGridDialog.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.analysis;
+
+import dr.app.gui.components.RealNumberField;
+import dr.app.gui.components.WholeNumberField;
+import dr.app.gui.util.LongTask;
+import beast.tracer.trace.TraceDistribution;
+import beast.tracer.trace.TraceType;
+import beast.tracer.trace.TraceList;
+import dr.stats.Variate;
+import jam.framework.DocumentFrame;
+import jam.panels.OptionsPanel;
+
+import javax.swing.*;
+import javax.swing.border.EmptyBorder;
+import javax.swing.event.ChangeEvent;
+import javax.swing.event.ChangeListener;
+import java.awt.*;
+import java.awt.event.ActionEvent;
+import java.awt.event.ActionListener;
+import java.io.File;
+import java.util.ArrayList;
+import java.util.HashSet;
+import java.util.List;
+import java.util.Set;
+
+public class SkyGridDialog {
+
+    private JFrame frame;
+
+    private String[][] argumentGuesses = {
+            {"popsize"},
+            {"cutoff"},
+    };
+
+    private String[] argumentNames = new String[]{
+            "Population Size",
+            "Grid Height",
+    };
+
+    //    private final JButton button = new JButton("Choose File...");
+    private ActionListener buttonListener;
+
[…]

### 2017-02-03 src/beast/tracer/analysis/TimeDensityDialog.java

@@ -1,0 +1,244 @@
+/*
+ * TimeDensityDialog.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.analysis;
+
+import dr.app.gui.components.RealNumberField;
+import dr.app.gui.util.LongTask;
+import beast.tracer.trace.TraceList;
+import dr.stats.Variate;
+import dr.util.FrequencyDistribution;
+import jam.panels.OptionsPanel;
+
+import javax.swing.*;
+import javax.swing.border.EmptyBorder;
+import java.awt.event.ActionEvent;
+import java.awt.event.ActionListener;
+import java.util.List;
+
+public class TimeDensityDialog {
+
+    private JFrame frame;
+
+    private JComboBox traceCombo;
+    private String timeTrace = "None selected";
+
+    private RealNumberField ageOfYoungestField = new RealNumberField();
+
+    private OptionsPanel optionPanel;
+
+    public TimeDensityDialog(JFrame frame) {
+        this.frame = frame;
+
+        traceCombo = new JComboBox();
+
+        ageOfYoungestField.setValue(0.0);
+        ageOfYoungestField.setColumns(12);
+
+        optionPanel = new OptionsPanel(12, 12);
+    }
+
+    private int findArgument(JComboBox comboBox, String argument) {
+        for (int i = 0; i < comboBox.getItemCount(); i++) {
+            String item = ((String) comboBox.getItemAt(i)).toLowerCase();
[…]

### 2017-02-03 src/beast/tracer/analysis/TraitThroughTimeDialog.java

@@ -1,0 +1,581 @@
+/*
+ * TraitThroughTimeDialog.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.analysis;
+
+import dr.app.gui.components.RealNumberField;
+import dr.app.gui.components.WholeNumberField;
+import dr.app.gui.util.LongTask;
+import beast.tracer.trace.TraceDistribution;
+import beast.tracer.trace.TraceList;
+import dr.stats.Variate;
+import jam.framework.DocumentFrame;
+import jam.panels.OptionsPanel;
+import jebl.evolution.graphs.Node;
+import jebl.evolution.io.ImportException;
+import jebl.evolution.io.NewickImporter;
+import jebl.evolution.io.NexusImporter;
+import jebl.evolution.io.TreeImporter;
+import jebl.evolution.trees.RootedTree;
+
+import javax.swing.*;
+import javax.swing.border.EmptyBorder;
+import javax.swing.event.ChangeEvent;
+import javax.swing.event.ChangeListener;
+import java.awt.*;
+import java.awt.event.ActionEvent;
+import java.awt.event.ActionListener;
+import java.io.BufferedReader;
+import java.io.File;
+import java.io.FileReader;
+import java.util.List;
+
+public class TraitThroughTimeDialog {
+
+    private JFrame frame;
+
+    private final JButton button = new JButton("Choose File...");
+    private ActionListener buttonListener;
+
[…]

### 2017-02-03 src/beast/tracer/trace/ArrayTraceList.java

@@ -1,0 +1,122 @@
+/*
+ * ArrayTraceList.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.trace;
+
+import java.util.HashMap;
+import java.util.List;
+import java.util.Map;
+
+/**
+ * @author Alexei Drummond
+ */
+public class ArrayTraceList extends AbstractTraceList {
+
+    String name;
+    Map<String, Integer> traceIndex = new HashMap<String, Integer>();
+    int burnin = 0;
+    List<Trace> traces;
+    int stepSize;
+
+    public ArrayTraceList(String name, List<Trace> traces, int burnin) {
+        this.name = name;
+        this.traces = traces;
+        this.burnin = burnin;
+
+        for (int i = 0; i < traces.size(); i++) {
+            traceIndex.put(traces.get(i).getName(), i);
+        }
+
+        Trace stateTrace = traces.get(0);
+        this.stepSize = (int) Math.round((Double) stateTrace.getValue(1) - (Double) stateTrace.getValue(0));
+    }
+
+    public String getName() {
+        return name;
+    }
+
+    public int getTraceCount() {
+        return traces.size();
+    }
+
+    public int getTraceIndex(String name) {
+        Integer index = traceIndex.get(name);
+        if (index == null) return -1;
+        return index;
+    }
+
+    public String getTraceName(int index) {
+        return traces.get(index).getName();
+    }
+
+    /**
+     * @return the burn-in for this trace list (the number of sampled states to discard)
+     */
[…]

### 2017-02-03 src/beast/tracer/trace/CnCsPerSiteAnalysis.java

@@ -1,0 +1,475 @@
+/*
+ * CnCsPerSiteAnalysis.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.trace;
+
+import beast.tracer.trace.LogFileTraces;
+import beast.tracer.trace.TraceDistribution;
+import beast.tracer.trace.TraceException;
+import beast.tracer.trace.TraceList;
+import dr.util.*;
+import dr.xml.*;
+
+import java.io.File;
+import java.io.FileNotFoundException;
+import java.util.ArrayList;
+import java.util.Collections;
+import java.util.List;
+import java.util.StringTokenizer;
+
+/**
+ * @author Philippe Lemey
+ * @author Marc A. Suchard
+ */
+public class CnCsPerSiteAnalysis implements Citable {
+    public static final String CNCS_PER_SITE_ANALYSIS = "cNcSPerSiteAnalysis";
+    public static final String BURN_IN = "burnin";
+    public static final String CUTOFF = "cutoff";
+    public static final String INCLUDE_SIGNIFICANT_SYMBOL = "includeSymbol";
+    public static final String INCLUDE_PVALUES = "includePValues";
+    public static final String INCLUDE_SITE_CLASSIFICATION = "includeClassification";
+    public static final String SEPARATOR_STRING = "separator";
+    public static final String INCLUDE_SIMULATION_OUTCOME = "simulationOutcome";
[…]

### 2017-02-03 src/beast/tracer/trace/CnCsToDnDsPerSiteAnalysis.java

@@ -1,0 +1,719 @@
+/*
+ * CnCsToDnDsPerSiteAnalysis.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.trace;
+
+import dr.math.EmpiricalBayesPoissonSmoother;
+import dr.stats.DiscreteStatistics;
+import dr.util.*;
+import dr.xml.*;
+
+import java.io.File;
+import java.io.FileNotFoundException;
+import java.util.ArrayList;
+import java.util.Collections;
+import java.util.List;
+import java.util.StringTokenizer;
+
+/**
+ * @author Philippe Lemey
+ * @author Marc A. Suchard
+ */
+public class CnCsToDnDsPerSiteAnalysis implements Citable {
+
+    public static final String CNCS_TO_DNDS_PER_SITE_ANALYSIS = "cNcSTodNdSPerSiteAnalysis";
+    public static final String BURN_IN = "burnin";
+    public static final String CUTOFF = "cutoff";
+    public static final String PROPORTION = "proportion";
+    public static final String INCLUDE_SIGNIFICANT_SYMBOL = "includeSymbol";
+    public static final String INCLUDE_SIGNIFICANCE_LEVEL = "includeLevel";
+    public static final String INCLUDE_SITE_CLASSIFICATION = "includeClassification";
+    public static final String SIGNIFICANCE_TEST = "test";
+    public static final String SEPARATOR_STRING = "separator";
[…]

### 2017-02-03 src/beast/tracer/trace/DensityEstimate.java

@@ -1,0 +1,132 @@
+/*
+ * DensityEstimate.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.trace;
+
+import beast.tracer.ui.chart.Axis;
+import beast.tracer.ui.chart.LinearAxis;
+import dr.stats.Variate;
+import dr.util.FrequencyDistribution;
+
+/**
+ * Creates a simple histogram-based density estimate
+ *
+ * @author Andrew Rambaut
+ * @version $Id$
+ */
+public class DensityEstimate {
+
+    public DensityEstimate(Double[] samples, int minimumBinCount) {
+        this.samples = new Variate.D(samples);
+
+        calculateDensity(this.samples, minimumBinCount);
+    }
+
+    protected void calculateDensity(Variate.D data, int minimumBinCount) {
+        frequencyDistribution = calculateFrequencies(samples, minimumBinCount);
+
+        xCoordinates = new Variate.D();
+        yCoordinates = new Variate.D();
+
+        double x = frequencyDistribution.getLowerBound() - frequencyDistribution.getBinSize();
+
+        xCoordinates.add(x + (frequencyDistribution.getBinSize() / 2.0));
+        yCoordinates.add(0.0);
+        x += frequencyDistribution.getBinSize();
+
+        for (int i = 0; i < frequencyDistribution.getBinCount(); i++) {
+            xCoordinates.add(x + (frequencyDistribution.getBinSize() / 2.0));
[…]

### 2017-02-03 src/beast/tracer/trace/DensityRectangle.java

@@ -1,0 +1,254 @@
+/*
+ * DensityRectangle.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.trace;
+
+import beast.tracer.trace.LogFileTraces;
+import beast.tracer.trace.TraceException;
+
+import javax.swing.*;
+import java.awt.*;
+import java.awt.geom.Point2D;
+import java.awt.geom.Rectangle2D;
+import java.io.File;
+import java.io.IOException;
+import java.util.ArrayList;
+import java.util.List;
+
+/**
+ * @author Alexei Drummond
+ * @version $Id: DensityRectangle.java,v 1.4 2006/08/26 19:24:41 rambaut Exp $
+ */
+public class DensityRectangle {
+
+    Rectangle2D rectangle;
+    List<Point2D> points;
+    int pointCount;
+    double area;
+
+    static int THRESHOLD = 80;
+
+    beast.tracer.trace.DensityRectangle[] quadrants = null;
+
+
+    public DensityRectangle(double x, double y, double w, double h) {
+
+        rectangle = new Rectangle2D.Double(x, y, w, h);
+        area = w * h;
+        points = new ArrayList<Point2D>();
+    }
+
+    public boolean contains(Point2D point) {
+
+        return rectangle.contains(point);
+    }
+
+    public boolean contains(double x, double y) {
+
+        return rectangle.contains(x, y);
+    }
+
+    public void addPoint(Point2D point) {
[…]

### 2017-02-03 src/beast/tracer/trace/DnDsPerSiteAnalysis.java

@@ -1,0 +1,634 @@
+/*
+ * DnDsPerSiteAnalysis.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.trace;
+
+import beast.tracer.trace.TraceDistribution;
+import dr.stats.DiscreteStatistics;
+import dr.util.*;
+import dr.xml.*;
+
+import java.io.File;
+import java.io.FileNotFoundException;
+import java.util.ArrayList;
+import java.util.Collections;
+import java.util.List;
+import java.util.StringTokenizer;
+
+/**
+ * @author Philippe Lemey
+ * @author Marc A. Suchard
+ */
+public class DnDsPerSiteAnalysis implements Citable {
+    public static final String DNDS_PER_SITE_ANALYSIS = "dNdSPerSiteAnalysis";
+    public static final String BURN_IN = "burnin";
+    public static final String CUTOFF = "cutoff";
+    public static final String PROPORTION = "proportion";
+    public static final String INCLUDE_SIGNIFICANT_SYMBOL = "includeSymbol";
+    public static final String INCLUDE_SIGNIFICANCE_LEVEL = "includeLevel";
+    public static final String INCLUDE_SITE_CLASSIFICATION = "includeClassification";
+    public static final String SIGNIFICANCE_TEST = "test";
+    public static final String SEPARATOR_STRING = "separator";
+    public static final String INCLUDE_SIMULATION_OUTCOME = "simulationOutcome";
[…]

### 2017-02-03 src/beast/tracer/trace/FilteredTraceList.java

@@ -1,0 +1,107 @@
+/*
+ * FilteredTraceList.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.trace;
+
+import beast.tracer.trace.Filter;
+import beast.tracer.trace.TraceList;
+
+/**
+ * not available for CombinedTraces yet
+ */
+public abstract class FilteredTraceList implements TraceList {
+
+//    protected boolean[] selected; // length = values[].length = getValuesSize, all must be true initially
+
+//    private void createSelected() { // will init in updateSelected()
+//        if (getTrace(0) != null) {
+//            selected = new boolean[getTrace(0).getValuesSize()];
+//        } else {
+//            throw new RuntimeException("Cannot initial filters ! getTrace(0) failed !");
+//        }
+//    }
+
+//    private void initSelected() {
+//        for (int i = 0; i < selected.length; i++) {
+//            selected[i] = true;
+//        }
+//    }
+
+    public boolean hasFilter(int traceIndex) {
+//        if (selected == null) return false;
+        return getTrace(traceIndex).getFilter() != null;
+    }
+
+    public void setFilter(int traceIndex, Filter filter) {
+//        if (selected == null) createSelected();
+        getTrace(traceIndex).setFilter(filter);
+        refreshStatistics();
+    }
+
+    public Filter getFilter(int traceIndex) {
+//        if (selected == null) return null;
+        return getTrace(traceIndex).getFilter();
[…]

### 2017-02-03 src/beast/tracer/trace/GeneralizedSteppingStoneSamplingAnalysis.java

@@ -1,0 +1,303 @@
+/*
+ * GeneralizedSteppingStoneSamplingAnalysis.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.trace;
+
+import beast.tracer.trace.TraceException;
+import dr.util.Attribute;
+import dr.util.FileHelpers;
+import dr.xml.*;
+
+import java.io.BufferedWriter;
+import java.io.File;
+import java.io.FileNotFoundException;
+import java.io.FileWriter;
+import java.util.*;
+
+/**
+ * @author Guy Baele
+ */
+
+public class GeneralizedSteppingStoneSamplingAnalysis {
+	
+    public static final String GENERALIZED_STEPPING_STONE_SAMPLING_ANALYSIS = "generalizedSteppingStoneSamplingAnalysis";
+    public static final String RESULT_FILE_NAME = "resultsFileName";
+    public static final String THETA_COLUMN = "thetaColumn";
+    public static final String SOURCE_COLUMN = "sourceColumn";
+    public static final String DESTINATION_COLUMN = "destinationColumn";
+    public static final String FORMAT = "%5.5g";
+    
+    private final String sourceName, destinationName;
+    private final List<Double> thetaSample;
+    private final List<Double> sourceSample;
+    private final List<Double> destinationSample;
+    private boolean logBayesFactorCalculated = false;
+    private double logBayesFactor;
+    private List<Double> maxLogLikelihood;
[…]

### 2017-02-03 src/beast/tracer/trace/LogFileTraceExporter.java

@@ -1,0 +1,98 @@
+/*
+ * LogFileTraceExporter.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.trace;
+
+import beast.tracer.trace.TraceDistribution;
+import dr.util.TabularData;
+
+import java.io.File;
+import java.io.IOException;
+
+/**
+ * Export trace analysis data such as mean,median,HPD and ESS of trace variables.
+ *
+ * @author Joseph Heled
+ *         Date: 25/10/2007
+ */
+public class LogFileTraceExporter extends TabularData {
+    private final LogFileTraces analysis;
+    private final String[] rows = {"mean", "median", "hpdLower", "hpdUpper", "ESS"};
+    beast.tracer.trace.TraceDistribution[] distributions;
+
+    public LogFileTraceExporter(File file, int burnin) throws TraceException, IOException {
+
+        analysis = new LogFileTraces(file.getCanonicalPath(), file);
+        analysis.loadTraces();
+        if (burnin >= 0) {
+            analysis.setBurnIn(burnin);
+        }
+
+        distributions = new beast.tracer.trace.TraceDistribution[nColumns()];
+    }
+
+    public int nColumns() {
+        return analysis.getTraceCount();
+    }
+
+    public String columnName(int nColumn) {
+        return analysis.getTraceName(nColumn);
+    }
+
+    public int nRows() {
+        return rows.length;
+    }
+
[…]

### 2017-02-03 src/beast/tracer/trace/LogFileTraces.java

@@ -1,0 +1,520 @@
+/*
+ * LogFileTraces.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+package beast.tracer.trace;
+
+import java.io.*;
+import java.util.ArrayList;
+import java.util.List;
+import java.util.StringTokenizer;
+import java.util.TreeMap;
+
+/**
+ * A class that stores a set of traces from a single chain
+ *
+ * @author Andrew Rambaut
+ * @author Alexei Drummond
+ * @version $Id: LogFileTraces.java,v 1.4 2006/11/30 17:39:29 rambaut Exp $
+ */
+
+public class LogFileTraces extends AbstractTraceList {
+
+    public LogFileTraces(String name, File file) {
+        this.name = name;
+        this.file = file;
+    }
+
+    /**
+     * @return the name of this traceset
+     */
+    public String getName() {
+        return name;
+    }
+
+    public File getFile() {
+        return file;
+    }
+
+    /**
+     * @return the last state in the chain
+     */
+    public long getMaxState() {
+        return lastState;
+    }
+
+    public boolean isIncomplete() {
+        return false;
+    }
+
+    /**
+     * @return the number of states excluding the burnin
+     */
+    public int getStateCount() {
+        // This is done as two integer divisions to ensure the same rounding for
+        // the burnin...
+        return (int) (((lastState - firstState) / stepSize) - (getBurnIn() / stepSize) + 1);
+    }
+
+    /**
[…]

### 2017-02-03 src/beast/tracer/trace/MarginalLikelihoodAnalysis.java

@@ -1,0 +1,358 @@
+/*
+ * MarginalLikelihoodAnalysis.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+package beast.tracer.trace;
+
+import dr.math.LogTricks;
+import dr.math.MathUtils;
+import dr.util.TaskListener;
+
+import java.util.ArrayList;
+import java.util.List;
+
+/**
+ * @author Marc Suchard
+ * @author Alexei Drummond
+ *         <p/>
+ *         Source translated from model_P.c (a component of BAli-Phy by Benjamin Redelings and Marc Suchard
+ */
+public class MarginalLikelihoodAnalysis {
+
+    private final String traceName;
+    private final List<Double> sample;
+    private final int burnin;
+    private final String analysisType; // "harmonic" for harmonic mean, "smoothed" for smoothed harmonic mean, "aicm" for AICM, "arithmetic" for arithmetic mean 
+    private final int bootstrapLength;
+
+    private boolean marginalLikelihoodCalculated = false;
+    private double logMarginalLikelihood;
+    private double bootstrappedSE;
+
+
+//    public MarginalLikelihoodAnalysis(double[] sample, String traceName, int burnin) {
+//        this(sample, traceName, burnin, false, 1000);
+//    }
+
+    public String getTraceName() {
+        return traceName;
+    }
+
+    public int getBurnin() {
+        return burnin;
+    }
+
+    /**
+     * Constructor
+     *
+     * @param sample
+     * @param traceName       used for 'toString' display purposes only
[…]

### 2017-02-03 src/beast/tracer/trace/OldDnDsPerSiteAnalysis.java

@@ -1,0 +1,486 @@
+/*
+ * OldDnDsPerSiteAnalysis.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.trace;
+
+import beast.tracer.trace.LogFileTraces;
+import beast.tracer.trace.TraceException;
+import dr.inferencexml.trace.MarginalLikelihoodAnalysisParser;
+import dr.math.EmpiricalBayesPoissonSmoother;
+import dr.stats.DiscreteStatistics;
+import dr.util.Attribute;
+import dr.util.FileHelpers;
+import dr.util.HeapSort;
+import dr.xml.*;
+
+import java.io.File;
+import java.io.FileNotFoundException;
+import java.io.IOException;
+import java.util.List;
+
+/**
+ * @author Philippe Lemey
+ * @author Marc A. Suchard
+ */
+public class OldDnDsPerSiteAnalysis {
+    public static final String DNDS_PERSITE_ANALYSIS = "olddNdSPerSiteAnalysis";
+    public static final String COND_SPERSITE_COLUMNS = "conditionalSperSite";
+    public static final String UNCOND_SPERSITE_COLUMNS = "unconditionalSperSite";
+    public static final String COND_NPERSITE_COLUMNS = "conditionalNperSite";
+    public static final String UNCOND_NPERSITE_COLUMNS = "unconditionalNperSite";
+    //public static final String ALIGNMENT = "alignment";
+
[…]

### 2017-02-03 src/beast/tracer/trace/PathSamplingAnalysis.java

@@ -1,0 +1,294 @@
+/*
+ * PathSamplingAnalysis.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.trace;
+
+import beast.tracer.trace.LogFileTraces;
+import beast.tracer.trace.TraceException;
+import dr.inferencexml.trace.MarginalLikelihoodAnalysisParser;
+import dr.util.Attribute;
+import dr.util.FileHelpers;
+import dr.xml.*;
+
+import java.io.BufferedWriter;
+import java.io.File;
+import java.io.FileNotFoundException;
+import java.io.FileWriter;
+import java.util.*;
+
+/**
+ * @author Marc A. Suchard
+ * @author Alexander Alekseyenko
+ * @author Guy Baele
+ */
+public class PathSamplingAnalysis {
+
+    public static final String PATH_SAMPLING_ANALYSIS = "pathSamplingAnalysis";
+    public static final String RESULT_FILE_NAME = "resultsFileName";
+    public static final String LIKELIHOOD_COLUMN = "likelihoodColumn";
+    public static final String THETA_COLUMN = "thetaColumn";
+    public static final String FORMAT = "%5.5g";
+
+    PathSamplingAnalysis(String logLikelihoodName, List<Double> logLikelihoodSample, List<Double> thetaSample) {
+        this.logLikelihoodSample = logLikelihoodSample;
+        this.logLikelihoodName = logLikelihoodName;
[…]

### 2017-02-03 src/beast/tracer/trace/SteppingStoneSamplingAnalysis.java

@@ -1,0 +1,275 @@
+/*
+ * SteppingStoneSamplingAnalysis.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.trace;
+
+import beast.tracer.trace.LogFileTraces;
+import beast.tracer.trace.TraceException;
+import dr.util.Attribute;
+import dr.util.FileHelpers;
+import dr.xml.*;
+
+import java.io.BufferedWriter;
+import java.io.File;
+import java.io.FileNotFoundException;
+import java.io.FileWriter;
+import java.util.*;
+
+/**
+ * @author Guy Baele
+ */
+
+public class SteppingStoneSamplingAnalysis {
+	
+    public static final String STEPPING_STONE_SAMPLING_ANALYSIS = "steppingStoneSamplingAnalysis";
+    public static final String RESULT_FILE_NAME = "resultsFileName";
+    public static final String LIKELIHOOD_COLUMN = "likelihoodColumn";
+    public static final String THETA_COLUMN = "thetaColumn";
+    public static final String FORMAT = "%5.5g";
+    
+    private final String logLikelihoodName;
+    private final List<Double> logLikelihoodSample;
+    private final List<Double> thetaSample;
+    private boolean logBayesFactorCalculated = false;
+    private double logBayesFactor;
+    private List<Double> maxLogLikelihood;
+    private List<Double> mlContribution;
+    private List<Double> orderedTheta;
+    
[…]

### 2017-02-03 src/beast/tracer/trace/Trace.java

@@ -1,0 +1,294 @@
+/*
+ * Trace.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.trace;
+
+import beast.tracer.trace.Filter;
+import beast.tracer.trace.TraceCorrelation;
+import beast.tracer.trace.TraceType;
+
+import java.util.*;
+
+/**
+ * A simple class that stores a trace for a single statistic
+ *
+ * @author Andrew Rambaut
+ * @author Alexei Drummond
+ * @version $Id: Trace.java,v 1.11 2005/07/11 14:07:26 rambaut Exp $
+ */
+public class Trace<T> { // TODO get rid of generic to make things easy
+
+//    public static final int INITIAL_SIZE = 1000;
+//    public static final int INCREMENT_SIZE = 1000;
+
+    // use <Double> for integer, but traceType must = ORDINAL, because of legacy issue at analyseCorrelationContinuous
+    protected beast.tracer.trace.TraceType traceType = beast.tracer.trace.TraceType.REAL;
+    protected List<T> values = new ArrayList<T>(); // TODO change to String only, and parse to double, int or string in getValues according to trace type
+    //    protected int valueCount = 0;
+    protected String name;
+
+//    private Object[] range;
+
+    public Trace(String name) { // traceType = TraceFactory.TraceType.DOUBLE;
+        this.name = name;
+    }
+
+    public Trace(String name, beast.tracer.trace.TraceType traceType) {
[…]

### 2017-02-03 src/beast/tracer/trace/TraceDistribution.java

@@ -1,0 +1,351 @@
+/*
+ * TraceDistribution.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.trace;
+
+import beast.tracer.trace.TraceType;
+import dr.stats.DiscreteStatistics;
+import dr.util.HeapSort;
+
+import java.util.*;
+
+/**
+ * A class that stores the distribution statistics for a trace
+ *
+ * @author Andrew Rambaut
+ * @author Alexei Drummond
+ * @version $Id: TraceDistribution.java,v 1.1.1.2 2006/04/25 23:00:09 rambaut Exp $
+ */
+public class TraceDistribution<T> {
+    private TraceType traceType;
+
+    public TraceDistribution(List<T> values, TraceType traceType) {
+        this.traceType = traceType;
+        initStatistics(values, 0.95);
+    }
+
+    public TraceDistribution(List<T> values, TraceType traceType, double ESS) {
+        this(values, traceType);
+        this.ESS = ESS;
+    }
+
+    public TraceType getTraceType() {
+        return traceType;
+    }
+
+    public void setTraceType(TraceType traceType) {
+        this.traceType = traceType;
+    }
+
+    public boolean isValid() {
+        return isValid;
+    }
+
+    public double getMean() {
+        return mean;
+    }
+
+    public double getVariance() {
+        return variance;
+    }
+
+    public double getStdError() {
+        return stdError;
+    }
+
[…]

### 2017-02-03 src/beast/tracer/trace/TraceList.java

@@ -1,0 +1,137 @@
+/*
+ * TraceList.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.trace;
+
+import beast.tracer.trace.TraceCorrelation;
+import beast.tracer.trace.TraceDistribution;
+
+import java.util.List;
+
+/**
+ * An interface and default class that stores a set of traces from a single chain
+ *
+ * @author Andrew Rambaut
+ * @author Alexei Drummond
+ * @version $Id: TraceList.java,v 1.2 2006/05/22 15:52:06 rambaut Exp $
+ */
+
+public interface TraceList {
+
+    /**
+     * @return the name of this trace list
+     */
+    String getName();
+
+    /**
+     * @return the number of traces in this trace list
+     */
+    int getTraceCount();
+
+    /**
+     * @param name trace name
+     * @return the index of the trace with the given name
+     */
+    int getTraceIndex(String name);
+
+    /**
+     * @param index trace ordinal index
+     * @return the name of the trace with the given index
+     */
+    String getTraceName(int index);
+
+    /**
+     * @return the burn-in for this trace list (the number of actual states to discard)
+     */
+    int getBurnIn();
+
+    /**
+     * @return the number of states in the traces (without the burnin)
+     */
+    int getStateCount();
+
+    /**
+     * @return the number of states in the burnin
+     */
+    int getBurninStateCount();
+
+    /**
+     * @return the size of the step between states
[…]

### 2017-02-03 src/beast/tracer/trace/TracesListener.java

@@ -1,0 +1,36 @@
+/*
+ * TracesListener.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.trace;
+
+/**
+ * @author Alexei Drummond
+ */
+public interface TracesListener {
+
+    void traceNames(String[] names);
+
+    void traceRow(int state, double[] values);
+}

### 2017-02-03 src/beast/tracer/trace/TracesReader.java

@@ -1,0 +1,147 @@
+/*
+ * TracesReader.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.trace;
+
+import java.io.Reader;
+import java.util.ArrayList;
+import java.util.List;
+import java.util.StringTokenizer;
+
+/**
+ * @author Alexei Drummond
+ */
+public class TracesReader {
+
+    public TracesReader() {
+    }
+
+    public void addTracesListener(TracesListener listener) {
+        listeners.add(listener);
+    }
+
+    public void removeTracesListener(TracesListener listener) {
+        listeners.remove(listener);
+    }
+
+    /**
+     * Reads all the traces in a file, but does not store in memory.
+     *
+     * @param r the reader to read traces from
+     * @throws TraceException      when trace contents is not valid
+     * @throws java.io.IOException low level problems with file
+     */
+    public void readTraces(Reader r) throws TraceException, java.io.IOException {
+
+        LogFileTraces.TrimLineReader reader = new LogFileTraces.TrimLineReader(r);
+
+        // Read through to first token
+        StringTokenizer tokens = reader.tokenizeLine();
+
+        // read over empty lines
+        while (!tokens.hasMoreTokens()) {
+            tokens = reader.tokenizeLine();
+        }
+
+        // skip the first column which should be the state number
+        String token = tokens.nextToken();
+
+        // lines starting with [ are ignored, assuming comments in MrBayes file
[…]

### 2017-02-03 src/beast/tracer/ui/chart/AreaPlot.java

@@ -1,0 +1,162 @@
+/*
+ * AreaPlot.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+import dr.stats.Variate;
+
+import java.awt.*;
+import java.awt.geom.GeneralPath;
+import java.util.List;
+
+/**
+ * Description:	An area plot.
+ *
+ * @author Andrew Rambaut
+ * @author Alexei Drummond
+ * @version $Id: AreaPlot.java,v 1.4 2005/05/24 20:25:59 rambaut Exp $
+ */
+
+public class AreaPlot extends Plot.AbstractPlot {
+
+    protected Variate.N xData2 = null;
+    protected Variate.N yData2 = null;
+
+    /**
+     * Constructor
+     */
+    public AreaPlot(Variate.N xData, Variate.N yData) {
+        super(xData, yData);
+    }
+
+    /**
+     * Constructor
+     */
+    public AreaPlot(List<Double> xData, List<Double> yData) {
+        super(xData, yData);
+    }
+
+    /**
+     * Constructor
+     */
+    public AreaPlot(Variate.N xData1, Variate.N yData1, Variate.N xData2, Variate.N yData2) {
+        super(xData1, yData1);
+        this.xData2 = xData2;
+        this.yData2 = yData2;
+    }
+
+    /**
+     * Constructor
+     */
+    public AreaPlot(List<Double> xData1, List<Double> yData1, List<Double> xData2, List<Double> yData2) {
[…]

### 2017-02-03 src/beast/tracer/ui/chart/Axis.java

@@ -1,0 +1,858 @@
+/*
+ * Axis.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+import dr.util.NumberFormatter;
+
+/**
+ * Axis.java
+ *
+ * Description:	Provides an interface for axis
+ * @author Andrew Rambaut
+ * @version	$Id: Axis.java,v 1.11 2005/05/24 20:25:59 rambaut Exp $
+ */
+
+public interface Axis {
+
+    // These constants are used for automatic scaling to select exactly
+    // where the axis starts and stops.
+    static public final int AT_MAJOR_TICK=0;
+    static public final int AT_MAJOR_TICK_PLUS=1;
+    static public final int AT_MINOR_TICK=2;
+    static public final int AT_MINOR_TICK_PLUS=3;
+    static public final int AT_DATA=4;
+    static public final int AT_ZERO=5;
+    static public final int AT_VALUE=6;
+
+    /**
+     *	Set axis flags
+     */
+    void setAxisFlags(int minAxisFlag, int maxAxisFlag);
+
+    /**
+     *	Set preferred number of ticks
+     */
+    void setPrefNumTicks(int prefNumMajorTicks, int prefNumMinorTicks);
+
+    /**
+     *	Set integer scale
+     */
+    void setIsDiscrete(boolean isDiscrete);
+
+    /**
+     *	Set show label for first tick
+     */
[…]

### 2017-02-03 src/beast/tracer/ui/chart/BlankAxis.java

@@ -1,0 +1,73 @@
+/*
+ * BlankAxis.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+
+public class BlankAxis extends Axis.AbstractAxis {
+
+    /**
+     * Empty constructor
+     */
+    public BlankAxis() {
+    }
+
+    /**
+     * Axis flag constructor
+     */
+    public BlankAxis(int minAxisFlag, int maxAxisFlag) {
+
+        setAxisFlags(minAxisFlag, maxAxisFlag);
+    }
+
+    /**
+     * Returns 0 as we want a blank axis
+     */
+    public int getMajorTickCount() {
+        return 0;
+    }
+
+    /**
+     * Returns 0 as we want a blank axis
+     */
+    public int getMinorTickCount(int majorTickNo) {
+        return 0;
+    }
+
+    /**
+     * Transform a value
+     */
+    public double transform(double value) {
+        return value;    // a linear transform !
+    }
+
+    /**
+     * Untransform a value
+     */
+    public double untransform(double value) {
+        return value;    // a linear transform !
+	}
+}
+

### 2017-02-03 src/beast/tracer/ui/chart/CategoryDensityPlot.java

@@ -1,0 +1,137 @@
+/*
+ * CategoryDensityPlot.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+import dr.stats.Variate;
+import dr.util.FrequencyDistribution;
+import beast.tracer.trace.TraceDistribution;
+
+import java.awt.*;
+import java.util.List;
+
+public class CategoryDensityPlot extends FrequencyPlot {
+    private int barCount = 0;
+    private int barId;
+    // for string[], passing the int[] storing the index of string[]
+
+    public CategoryDensityPlot(List<Double> data, int minimumBinCount, TraceDistribution traceDistribution,
+                               int barCount, int barId) {
+        super(traceDistribution);
+        this.barCount = barCount;
+        this.barId = barId;
+
+        setData(new Variate.D(data), minimumBinCount);
+    }
+
+    /**
+     * Set data
+     */
+    public void setData(Variate data, int minimumBinCount) {
+        raw = data;
+        FrequencyDistribution frequency = getFrequencyDistribution(data, minimumBinCount);
+
+        Variate.D xData = new Variate.D();
+        Variate.D yData = new Variate.D();
+
+        double x = frequency.getLowerBound();
+
+        for (int i = 0; i < frequency.getBinCount(); i++) {
+
+            xData.add(x);
[…]

### 2017-02-03 src/beast/tracer/ui/chart/ChartException.java

@@ -1,0 +1,37 @@
+/*
+ * ChartException.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+
+public class ChartException extends Exception {
+	/**
+	 *
+	 */
+	private static final long serialVersionUID = 5400944779931674299L;
+	public ChartException() { super(); }
+	public ChartException(String s) { super(s); }
+}
+

### 2017-02-03 src/beast/tracer/ui/chart/ChartSelector.java

@@ -1,0 +1,136 @@
+/*
+ * ChartSelector.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+import java.awt.*;
+import java.awt.event.MouseEvent;
+import java.awt.event.MouseListener;
+import java.awt.event.MouseMotionListener;
+import java.awt.geom.Point2D;
+import java.awt.geom.Rectangle2D;
+
+/**
+ * @author Andrew Rambaut
+ * @version $Id: TreePaneSelector.java 822 2007-10-26 13:50:26Z rambaut $
+ */
+public class ChartSelector implements MouseListener, MouseMotionListener {
+
+    public ChartSelector(JChart chart) {
+        this.chart = chart;
+        chart.addMouseListener(this);
+        chart.addMouseMotionListener(this);
+
+    }
+
+    public boolean isCrossHairCursor() {
+        return crossHairCursor;
+    }
+
+    public void setCrossHairCursor(boolean crossHairCursor) {
+        this.crossHairCursor = crossHairCursor;
+    }
+
+    private void setupCursor() {
+        if (crossHairCursor) {
+            chart.setCursor(Cursor.getPredefinedCursor(Cursor.CROSSHAIR_CURSOR));
+        } else {
+            chart.setCursor(Cursor.getPredefinedCursor(Cursor.DEFAULT_CURSOR));
+        }
+
+        chart.repaint();
[…]

### 2017-02-03 src/beast/tracer/ui/chart/DensityEstimatePlot.java

@@ -1,0 +1,98 @@
+/*
+ * DensityEstimatePlot.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+import beast.tracer.trace.DensityEstimate;
+import dr.stats.Variate;
+
+import java.awt.*;
+import java.awt.geom.GeneralPath;
+
+public class DensityEstimatePlot extends Plot.AbstractPlot {
+
+    public boolean isSolid() {
+        return solid;
+    }
+
+    public void setSolid(boolean solid) {
+        this.solid = solid;
+    }
+
+    boolean solid = true;
+
+    public DensityEstimatePlot(DensityEstimate densityEstimate) {
+        setData(densityEstimate.getXCoordinates(), densityEstimate.getYCoordinates());
+    }
+
+
+    public void setRelativeDensity(boolean relative) {
+//        relativeDensity = relative;
+//        setData(getRawData(), minimumBinCount);
+    }
+
+    /**
+     * Set bar fill style. Use a barPaint of null to not fill bar.
+     * Bar outline style is set using setLineStyle
+     */
+    public void setBarFillStyle(Paint barPaint) {
+        throw new IllegalArgumentException();
+    }
+
+    /**
+     * Paint data series
+     */
+    protected void paintData(Graphics2D g2, Variate.N xData, Variate.N yData) {
+
+        int n = xData.getCount();
+
[…]

### 2017-02-03 src/beast/tracer/ui/chart/DiscreteAxis.java

@@ -1,0 +1,99 @@
+/*
+ * DiscreteAxis.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+
+public class DiscreteAxis extends Axis.AbstractAxis {
+
+	private boolean originBetweenCategories;
+	private boolean showEveryCategory;
+
+	public DiscreteAxis(boolean originBetweenCategories, boolean showEveryCategory) {
+		super(AT_MAJOR_TICK, AT_MAJOR_TICK, true);
+
+		this.originBetweenCategories = originBetweenCategories;
+		this.showEveryCategory = showEveryCategory;
+	}
+
+	/**
+	*	Transform a value
+	*/
+	public double transform(double value) {
+		return value;	// a linear transform !
+	}
+
+	/**
+	*	Untransform a value
+	*/
+	public double untransform(double value) {
+		return value;	// a linear transform !
+	}
+
+	public void calibrate() {
+		majorTick=1;
+		minorTick=1;
+
+		minTick=minData;
+		maxTick=maxData;
+		majorTickCount = (int)((maxTick-minTick)/majorTick)+1;
+		minorTickCount = 0;
+
+		if (!showEveryCategory) {
+			while (majorTickCount > prefMajorTickCount) {
[…]

### 2017-02-03 src/beast/tracer/ui/chart/DiscreteJChart.java

@@ -1,0 +1,136 @@
+/*
+ * DiscreteJChart.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+import java.awt.*;
+import java.awt.geom.Line2D;
+import java.util.Map;
+
+/**
+ *
+ */
+public class DiscreteJChart extends JChart {
+
+    private Map<Integer, String> categoryDataMap;
+
+    public DiscreteJChart(Axis xAxis, Axis yAxis) {
+        super(xAxis, yAxis);
+    }
+
+    public void setXAxis(boolean isInteger, Map<Integer, String> categoryDataMap) {
+        if (categoryDataMap.isEmpty()) {
+            if (isInteger) {
+                super.setXAxis(new DiscreteAxis(true, true));
+            } else {
+                super.setXAxis(new LinearAxis(Axis.AT_MAJOR_TICK_PLUS, Axis.AT_MAJOR_TICK_PLUS));
+            }
+
+        } else {
+            super.setXAxis(new DiscreteAxis(true, true));
+        }
+
+        this.categoryDataMap = categoryDataMap;
+    }
+
+    public void setYAxis(boolean isInteger, Map<Integer, String> categoryDataMap) {
+        if (categoryDataMap.isEmpty()) {
+            if (isInteger) {
+                super.setYAxis(new DiscreteAxis(true, true));
+            } else {
+                super.setYAxis(new LinearAxis());
+            }
+
+        } else {
[…]

### 2017-02-03 src/beast/tracer/ui/chart/ErrorBarPlot.java

@@ -1,0 +1,115 @@
+/*
+ * ErrorBarPlot.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+import dr.stats.Variate;
+
+import java.awt.*;
+import java.awt.geom.GeneralPath;
+import java.util.List;
+
+/**
+ * Description:	An error bar plot (to be put under a scatter plot).
+ *
+ * @author Andrew Rambaut
+ * @version $Id: LinePlot.java,v 1.8 2005/05/24 20:25:59 rambaut Exp $
+ */
+
+public class ErrorBarPlot extends Plot.AbstractPlot {
+         public enum Orientation {
+             HORIZONTAL,
+             VERTICAL
+         }
+
+    /**
+     * Constructor
+     */
+    public ErrorBarPlot(Orientation orientation, Variate.N xData, Variate.N yData, Variate.N errorData) {
+        super(xData, yData);
+
+        this.errorData = errorData;
+        this.orientation = orientation;
+    }
+
+    /**
+     * Constructor
+     */
+    public ErrorBarPlot(Orientation orientation, List<Double> xData, List<Double> yData, List<Double> errorData) {
+        super(xData, yData);
+
+        this.errorData = new Variate.D(errorData);
+        this.orientation = orientation;
+    }
+
+    /**
+     * Paint data series
+     */
+    protected void paintData(Graphics2D g2, Variate.N xData, Variate.N yData) {
+
+        g2.setPaint(linePaint);
[…]

### 2017-02-03 src/beast/tracer/ui/chart/FrequencyPlot.java

@@ -1,0 +1,312 @@
+/*
+ * FrequencyPlot.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+import beast.tracer.trace.TraceDistribution;
+import beast.tracer.trace.TraceType;
+import dr.stats.Variate;
+import dr.util.FrequencyDistribution;
+
+import java.awt.*;
+import java.util.List;
+
+public class FrequencyPlot extends Plot.AbstractPlot {
+
+    protected Variate raw = null;
+
+//    protected Paint barPaint = Color.blue;
+    protected Paint barPaint = new Color(124, 164, 221);
+    private Paint quantilePaint = new Color(232, 114, 103);
+
+    private boolean hasQuantiles = false;
+    private double quantiles = 0;
+
+    private boolean hasIntervals = false;
+    private double upperInterval = 0.0;
+    private double lowerInterval = 0.0;
+
+    private boolean hasIncredibleSet = false;
+//    private TraceDistribution.CredibleSet credSet;
+
+    protected TraceDistribution traceDistribution = null;
+
+    protected FrequencyPlot(TraceDistribution traceDistribution) {
+        super();
+        this.traceDistribution = traceDistribution;
+    }
+
+    public FrequencyPlot(Variate.D data, int minimumBinCount) {
+        super();
+        setData(data, minimumBinCount);
+    }
+
[…]

### 2017-02-03 src/beast/tracer/ui/chart/KDESetupDialog.java

@@ -1,0 +1,215 @@
+/*
+ * KDESetupDialog.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+import dr.app.gui.components.RealNumberField;
+import jam.panels.OptionsPanel;
+
+import javax.swing.*;
+import javax.swing.border.EmptyBorder;
+import javax.swing.event.ChangeEvent;
+import javax.swing.event.ChangeListener;
+
+/**
+ * @author Marc A. Suchard
+ */
+public class KDESetupDialog {
+
+
+	private final JFrame frame;
+
+	private final JCheckBox logXAxis;
+	private final JCheckBox logYAxis;
+
+	private final JCheckBox manualXAxis;
+	private final JCheckBox manualYAxis;
+	private final RealNumberField minXValue;
+	private final RealNumberField maxXValue;
+	private final RealNumberField minYValue;
+	private final RealNumberField maxYValue;
+
+	private final boolean canLogXAxis;
+	private final boolean canLogYAxis;
+	private final int defaultMinXAxisFlag;
+	private final int defaultMaxXAxisFlag;
+	private final int defaultMinYAxisFlag;
[…]

### 2017-02-03 src/beast/tracer/ui/chart/LineDistributionPlot.java

@@ -1,0 +1,97 @@
+/*
+ * LineDistributionPlot.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+import dr.stats.Variate;
+
+import java.awt.*;
+
+public class LineDistributionPlot extends Plot.AbstractPlot {
+
+    private Variate gradients = null;
+    private Variate intercepts = null;
+
+    private int lineCount = 0;
+    private double x1, x2;
+    private double[] y1;
+    private double[] y2;
+
+    private boolean isCalibrated = false;
+
+    /**
+     * Constructor
+     */
+    public LineDistributionPlot(Variate gradients, Variate intercepts) {
+        this.gradients = gradients;
+        this.intercepts = intercepts;
+    }
+
+    /**
+     * Paint actual plot
+     */
+    public void paintPlot(Graphics2D g2, double xScale, double yScale,
+                          double xOffset, double yOffset) {
+
+        if (gradients == null || intercepts == null) {
+            return;
+        }
+
+        super.paintPlot(g2, xScale, yScale, xOffset, yOffset);
+
+        if (!isCalibrated) {
+            x1 = xAxis.getMinAxis();
+            x2 = xAxis.getMaxAxis();
+
+            lineCount = gradients.getCount();
+            y1 = new double[lineCount];
+            y2 = new double[lineCount];
+
[…]

### 2017-02-03 src/beast/tracer/ui/chart/LinePlot.java

@@ -1,0 +1,93 @@
+/*
+ * LinePlot.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+import dr.stats.Variate;
+
+import java.awt.*;
+import java.awt.geom.GeneralPath;
+import java.util.List;
+
+/**
+ * Description:	A line plot.
+ *
+ * @author Andrew Rambaut
+ * @author Alexei Drummond
+ * @version $Id: LinePlot.java,v 1.8 2005/05/24 20:25:59 rambaut Exp $
+ */
+
+public class LinePlot extends Plot.AbstractPlot {
+
+
+    /**
+     * Constructor
+     */
+    public LinePlot(Variate.N xData, Variate.N yData) {
+        super(xData, yData);
+    }
+
+    /**
+     * Constructor
+     */
+    public LinePlot(List<Double> xData, List<Double> yData) {
+        super(xData, yData);
+    }
+
+    /**
+     * Paint data series
+     */
+    protected void paintData(Graphics2D g2, Variate.N xData, Variate.N yData) {
+
+        double x = transformX(((Number) xData.get(0)).doubleValue());
+        double y = transformY(((Number) yData.get(0)).doubleValue());
+
+        GeneralPath path = new GeneralPath();
+        path.moveTo((float) x, (float) y);
+
+        int n = xData.getCount();
+        boolean failed = false;
+        for (int i = 1; i < n; i++) {
[…]

### 2017-02-03 src/beast/tracer/ui/chart/LogAxis.java

@@ -1,0 +1,224 @@
+/*
+ * LogAxis.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+public class LogAxis extends Axis.AbstractAxis {
+
+//	private int extraMinorTickCount;
+
+
+	/**
+	 * Empty constructor
+	 */
+	public LogAxis() {
+		super();
+        setSignficantFigures(0);
+    }
+
+	/**
+	 * Axis flag constructor
+	 */
+	public LogAxis(int minAxisFlag, int maxAxisFlag) {
+
+		setAxisFlags(minAxisFlag, maxAxisFlag);
+	}
+
+	/**
+	*	Set axis flags
+	*/
+	public void setAxisFlags(int minAxisFlag, int maxAxisFlag) {
+
+		if (minAxisFlag==AT_DATA || minAxisFlag==AT_ZERO) {
+			minAxisFlag = this.minAxisFlag;
+		}
+
+		if (maxAxisFlag==AT_DATA || maxAxisFlag==AT_ZERO) {
+			maxAxisFlag = this.maxAxisFlag;
+		}
+
+		super.setAxisFlags(minAxisFlag, maxAxisFlag);
+	}
+
+	/**
+	*	Set the range of the data.
+	*/
+	public void setRange(double minValue, double maxValue) {
+
+		if (maxValue <= 0.0)
+			maxValue = 1.0E-100;
+
[…]

### 2017-02-03 src/beast/tracer/ui/chart/NumericalDensityPlot.java

@@ -1,0 +1,191 @@
+/*
+ * NumericalDensityPlot.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+import beast.tracer.trace.TraceDistribution;
+import dr.stats.Variate;
+import dr.util.FrequencyDistribution;
+
+import java.awt.*;
+import java.awt.geom.GeneralPath;
+import java.awt.geom.Rectangle2D;
+
+public class NumericalDensityPlot extends FrequencyPlot {
+
+    boolean relativeDensity = true;
+    boolean pointsOnly = false;
+    int minimumBinCount;
+
+    public boolean isSolid() {
+        return solid;
+    }
+
+    public void setSolid(boolean solid) {
+        this.solid = solid;
+    }
+
+    boolean solid = true;
+
+    public NumericalDensityPlot(Variate.D data, int minimumBinCount) {
+        super(data, minimumBinCount);
+        this.minimumBinCount = minimumBinCount;
+    }
+
+//    public NumericalDensityPlot(double[] data, int minimumBinCount) {
+//        super(data, minimumBinCount);
+//        this.minimumBinCount = minimumBinCount;
+//    }
+
+    public NumericalDensityPlot(java.util.List<Double> data, int minimumBinCount, TraceDistribution traceD) {
+        super(data, minimumBinCount, traceD);
+        this.minimumBinCount = minimumBinCount;
+    }
+
[…]

### 2017-02-03 src/beast/tracer/ui/chart/Plot.java

@@ -1,0 +1,649 @@
+/*
+ * Plot.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+import dr.stats.Variate;
+
+import java.awt.*;
+import java.awt.geom.*;
+import java.util.Collection;
+import java.util.HashSet;
+import java.util.List;
+import java.util.Set;
+
+/**
+ * Plot.java
+ * <p/>
+ * Description:	Provides an interface for plots. Plots are elements of a
+ * chart that renders some data in a particular style. Multiple plots can
+ * be added to the same chart to get complex results.
+ *
+ * @author Andrew Rambaut
+ * @author Alexei Drummond
+ * @version $Id: Plot.java,v 1.23 2006/01/03 10:26:20 rambaut Exp $
+ */
+
+public interface Plot {
+
+    // These constants are used for automatic scaling to select exactly
+    // where the axis starts and stops.
+    static public final int NO_MARK = 0;
+    static public final int POINT_MARK = 1;
+    static public final int CROSS_MARK = 2;
+    static public final int PLUS_MARK = 3;
+    static public final int CIRCLE_MARK = 4;
+    static public final int SQUARE_MARK = 5;
+    static public final int DIAMOND_MARK = 6;
+
+    /**
+     * Set axes
+     */
+    void setAxes(Axis xAxis, Axis yAxis);
+
+    /**
[…]

### 2017-02-03 src/beast/tracer/ui/chart/RegressionPlot.java

@@ -1,0 +1,132 @@
+/*
+ * RegressionPlot.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+import dr.stats.Regression;
+import dr.stats.Variate;
+
+import java.awt.*;
+import java.util.List;
+
+public class RegressionPlot extends Plot.AbstractPlot {
+
+    private Regression regression;
+
+    /**
+     * Constructor
+     */
+    public RegressionPlot(Regression regression) {
+        super(regression.getXData(), regression.getYData());
+        this.regression = regression;
+    }
+
+    /**
+     * Constructor
+     */
+    public RegressionPlot(Variate.N xData, Variate.N yData, boolean forceOrigin) {
+        super(xData, yData);
+        setForceOrigin(forceOrigin);
+    }
+
+    /**
+     * Constructor
+     */
+    public RegressionPlot(List<Double> xData, List<Double> yData, boolean forceOrigin) {
+        super(xData, yData);
+        setForceOrigin(forceOrigin);
+    }
+
+    /**
+     * Set data
+     */
+    public void setData(List<Double> xData, List<Double> yData) {
+        super.setData(xData, yData);
+        regression = new Regression(this.xData, this.yData);
+    }
+
+    /**
+     * Set data
+     */
+    public void setData(Variate.N xData, Variate.N yData) {
+        super.setData(xData, yData);
+        regression = new Regression(this.xData, this.yData);
+    }
+
[…]

### 2017-02-03 src/beast/tracer/ui/chart/ScatterPlot.java

@@ -1,0 +1,210 @@
+/*
+ * ScatterPlot.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.chart;
+
+import dr.stats.Variate;
+
+import java.awt.*;
+import java.awt.geom.Point2D;
+import java.awt.geom.Rectangle2D;
+import java.util.List;
+import java.util.Set;
+import java.util.Vector;
+
+/**
+ * Description:	A scatter plot.
+ *
+ * @author Andrew Rambaut
+ * @author Alexei Drummond
+ * @version $Id: ScatterPlot.java,v 1.10 2005/05/24 20:25:59 rambaut Exp $
+ */
+
+public class ScatterPlot extends Plot.AbstractPlot {
+
+    private Vector<Rectangle2D> markBounds = null;
+
+    protected Stroke hilightedMarkStroke = new BasicStroke(0.5f);
+    protected Paint hilightedMarkPaint = Color.black;
+    protected Paint hilightedMarkFillPaint = Color.blue;
+
+    /**
+     * Constructor
+     */
+    public ScatterPlot(Variate.N xData, Variate.N yData) {
+        super(xData, yData);
+        setMarkStyle(CIRCLE_MARK, 5, new BasicStroke(1),
+                Color.black, Color.yellow);
+        setHilightedMarkStyle(new BasicStroke(1),
+                Color.black, Color.blue);
+    }
+
+    /**
+     * Constructor
[…]

### 2017-02-03 src/beast/tracer/ui/frame/AnalysisMenuFactory.java

@@ -1,0 +1,179 @@
+/*
+ * AnalysisMenuFactory.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.frame;
+
+import jam.framework.AbstractFrame;
+import jam.framework.MenuFactory;
+
+import javax.swing.*;
+
+/**
+ * @author rambaut
+ *         Date: Feb 24, 2005
+ *         Time: 5:12:11 PM
+ */
+public class AnalysisMenuFactory implements MenuFactory {
+
+    public static final String DEMOGRAPHIC_RECONSTRUCTION = "Demographic Reconstruction...";
+    public static final String BAYESIAN_SKYLINE_RECONSTRUCTION = "Bayesian Skyline Reconstruction...";
+    public static final String EXTENDED_BAYESIAN_SKYLINE_RECONSTRUCTION = "Extended Bayesian Skyline Reconstruction...";
+    public static final String GMRF_SKYRIDE_RECONSTRUCTION = "GMRF Skyride Reconstruction...";
+    public static final String SKY_GRID_RECONSTRUCTION = "SkyGrid Reconstruction...";
+    public static final String LINEAGES_THROUGH_TIME = "Lineages Through Time...";
+    public static final String TRAIT_THROUGH_TIME = "Trait Through Time...";
+
+    public static final String CREATE_TEMPORAL_ANALYSIS = "Create Temporal Analysis...";
[…]

### 2017-02-03 src/beast/tracer/ui/frame/AnalysisMenuHandler.java

@@ -1,0 +1,64 @@
+/*
+ * AnalysisMenuHandler.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.frame;
+
+import javax.swing.*;
+
+/**
+ * @author rambaut
+ *         Date: Feb 24, 2005
+ *         Time: 5:13:13 PM
+ */
+public interface AnalysisMenuHandler {
+
+    Action getDemographicAction();
+
+    Action getBayesianSkylineAction();
+
+    Action getExtendedBayesianSkylineAction();
+
+    Action getSkyGridAction();
+
+    Action getGMRFSkyrideAction();
+
+    Action getLineagesThroughTimeAction();
+
+    Action getTraitThroughTimeAction();
+
+    Action getCreateTemporalAnalysisAction();
+
+    Action getAddDemographicAction();
+
+    Action getAddBayesianSkylineAction();
+
+    Action getAddExtendedBayesianSkylineAction();
+
+    Action getAddTimeDensityAction();
+
+    Action getBayesFactorsAction();
+
+    Action getConditionalPosteriorDistAction();
+}

### 2017-02-03 src/beast/tracer/ui/frame/InstantiableTracerApp.java

@@ -1,0 +1,228 @@
+/*
+ * InstantiableTracerApp.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.frame;
+
+import beast.tracer.app.TracerApp;
+import dr.app.util.OSType;
+import beast.tracer.trace.LogFileTraces;
+import jam.framework.DocumentFrame;
+import jam.framework.SingleDocApplication;
+
+import javax.swing.*;
+import java.awt.*;
+import java.io.File;
+import java.security.Permission;
+import java.util.Locale;
+
+/**
+ *
+ * A class for calling an instance of Tracer within a BEAST (or any other Java) analysis.
+ * Allows a threaded call of Tracer (that won't terminate the JVM). A lot of the code is
+ * copied from TracerApp.java (but is a SingleDocApplication instead of a
+ * MultiDocApplication)
+ *
+ *
+ * @author Wai Lok Sibon Li
+ * @version $Id: InstantiableTracerApp.java,v 1.0 2012/09/17 15:23:33 sibon.li Exp $
+ */
+public class InstantiableTracerApp extends SingleDocApplication {
+    private boolean exiting;
+    public InstantiableTracerApp (String nameString, String aboutString, Icon icon,
+                         String websiteURLString, String helpURLString, boolean exiting) {
+            super(new TracerMenuBarFactory(), nameString, aboutString, icon,
+                    websiteURLString, helpURLString);
+
+            addPreferencesSection(new GeneralPreferencesSection());
[…]

### 2017-02-03 src/beast/tracer/ui/frame/TracerDefaultFileMenuFactory.java

@@ -1,0 +1,106 @@
+/*
+ * TracerDefaultFileMenuFactory.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.frame;
+
+import jam.framework.MenuFactory;
+import jam.framework.AbstractFrame;
+import jam.framework.Application;
+import jam.framework.MenuBarFactory;
+
+import javax.swing.*;
+import java.awt.event.KeyEvent;
+
+import dr.app.beauti.BeautiFrame;
+
+/**
+ * @author rambaut
+ *         Date: Dec 26, 2004
+ *         Time: 11:01:06 AM
+ */
+public class TracerDefaultFileMenuFactory implements MenuFactory {
+
+
+    public TracerDefaultFileMenuFactory() {
+    }
+
+    public String getMenuName() {
+        return "File";
+    }
+
+    public void populateMenu(JMenu menu, AbstractFrame frame) {
+
+        JMenuItem item;
+
+        Application application = Application.getApplication();
+        menu.setMnemonic('F');
+
+        item = new JMenuItem(application.getNewAction());
+        item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_N, MenuBarFactory.MENU_MASK));
+        menu.add(item);
+
+        // On Windows and Linux platforms, each window has its own menu so items which are not needed
+        // are simply missing. In contrast, on Mac, the menu is for the application so items should
[…]

### 2017-02-03 src/beast/tracer/ui/frame/TracerFileMenuHandler.java

@@ -1,0 +1,41 @@
+/*
+ * TracerFileMenuHandler.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.frame;
+
+import javax.swing.*;
+
+/**
+ * @author rambaut
+ *         Date: Feb 24, 2005
+ *         Time: 5:13:13 PM
+ */
+public interface TracerFileMenuHandler {
+
+	Action getExportDataAction();
+
+	Action getExportPDFAction();
+
+}

### 2017-02-03 src/beast/tracer/ui/frame/TracerFrame.java

@@ -1,0 +1,1797 @@
+/*
+ * TracerFrame.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.frame;
+
+import beast.tracer.ui.chart.ChartRuntimeException;
+import com.lowagie.text.Document;
+import com.lowagie.text.DocumentException;
+import com.lowagie.text.pdf.DefaultFontMapper;
+import com.lowagie.text.pdf.PdfContentByte;
+import com.lowagie.text.pdf.PdfTemplate;
+import com.lowagie.text.pdf.PdfWriter;
+import jam.framework.DocumentFrame;
+import jam.panels.ActionPanel;
+import jam.table.TableRenderer;
+import dr.app.gui.FileDrop;
+import dr.app.gui.table.TableEditorStopper;
+import dr.app.gui.util.LongTask;
+import beast.tracer.ui.panel.*;
+import beast.tracer.trace.*;
+import beast.tracer.analysis.*;
+
+import javax.swing.*;
+import javax.swing.border.Border;
+import javax.swing.event.ListSelectionEvent;
+import javax.swing.event.ListSelectionListener;
+import javax.swing.filechooser.FileNameExtensionFilter;
+import javax.swing.plaf.BorderUIResource;
+import javax.swing.table.AbstractTableModel;
+import javax.swing.table.JTableHeader;
+import java.awt.*;
+import java.awt.event.*;
[…]

### 2017-02-03 src/beast/tracer/ui/frame/TracerMenuBarFactory.java

@@ -1,0 +1,51 @@
+/*
+ * TracerMenuBarFactory.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.frame;
+
+import jam.framework.*;
+import jam.mac.*;
+
+import dr.app.util.OSType;
+
+public class TracerMenuBarFactory extends DefaultMenuBarFactory {
+
+    public TracerMenuBarFactory() {
+        if (OSType.isMac()) {
+            registerMenuFactory(new TracerMacFileMenuFactory());
+            registerMenuFactory(new MacEditMenuFactory());
+	        registerMenuFactory(new AnalysisMenuFactory());
+            registerMenuFactory(new MacWindowMenuFactory());
+            registerMenuFactory(new MacHelpMenuFactory());
+        } else {
+            registerMenuFactory(new TracerDefaultFileMenuFactory());
+            registerMenuFactory(new DefaultEditMenuFactory());
+	        registerMenuFactory(new AnalysisMenuFactory());
+            registerMenuFactory(new DefaultHelpMenuFactory());
+        }
+
+    }
+}
+

### 2017-02-03 src/beast/tracer/ui/panel/CombinedTraces.java

@@ -1,0 +1,268 @@
+/*
+ * CombinedTraces.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.panel;
+
+import beast.tracer.trace.*;
+
+import java.util.ArrayList;
+import java.util.List;
+
+
+/**
+ * A class for analysing multiple tracesets
+ *
+ * @author Andrew Rambaut
+ * @author Alexei Drummond
+ * @version $Id: CombinedTraces.java,v 1.3 2006/11/29 16:04:12 rambaut Exp $
+ */
+
+public class CombinedTraces extends FilteredTraceList { //implements TraceList {
+
+    public CombinedTraces(String name, LogFileTraces[] traceLists) throws TraceException {
+
+        if (traceLists == null || traceLists.length < 1) {
+            throw new TraceException("Must have at least 1 Traces object in a CombinedTraces");
+        }
+
+        this.name = name;
+        this.traceLists = new LogFileTraces[traceLists.length];
+        this.traceLists[0] = traceLists[0];
+
+        for (int i = 1; i < traceLists.length; i++) {
+            if (traceLists[i].getTraceCount() != traceLists[0].getTraceCount()) {
+                throw new TraceException("Cannot add to a CombinedTraces: the count of new traces do not match existing traces");
+            }
+
+            if (traceLists[i].getStepSize() != traceLists[0].getStepSize()) {
[…]

### 2017-02-03 src/beast/tracer/ui/panel/FilterAbstractPanel.java

@@ -1,0 +1,52 @@
+/*
+ * FilterAbstractPanel.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.panel;
+
+import javax.swing.*;
+
+/**
+ * @author Walter Xie
+ */
+public abstract class FilterAbstractPanel extends JPanel {
+
+    abstract String[] getSelectedValues();
+
+    boolean containsNullValue() {
+        if (getSelectedValues() == null || getSelectedValues().length < 1) return true;
+        for (String ob : getSelectedValues()) {
+            if (ob == null || ob.equals("")) return true;
+        }
+        return false;
+    }
+//
+//    FilterAbstractPanel getInstance(String[] a, String[] b, TraceFactory.TraceType traceType) {
+//        if (traceType == TraceFactory.TraceType.DOUBLE) {
+//            return new FilterContinuousPanel(a, b);
+//        } else {
+//            return new FilterDiscretePanel(a, b);
+//        }
+//    }
+}

### 2017-02-03 src/beast/tracer/ui/panel/FilterDialog.java

@@ -1,0 +1,131 @@
+/*
+ * FilterDialog.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.panel;
+
+import beast.tracer.ui.frame.TracerFrame;
+import dr.app.gui.util.LongTask;
+
+import javax.swing.*;
+import javax.swing.border.EmptyBorder;
+
+/**
+ * @author Andrew Rambaut
+ * @author Alexei Drummond
+ * @author Walter Xie
+ */
+public class FilterDialog extends JDialog {
+
+    private TracerFrame frame;
+    Object[] options = {"Apply Filter Changes", "Remove All Filters", "Cancel"};
+
+    public FilterDialog(TracerFrame frame) {
+        this.frame = frame;
+    }
+
+    public String showDialog(FilterListPanel filterListPanel, String previousMessage) {
+        String message = "";
+
+        JOptionPane optionPane = new JOptionPane(filterListPanel,
+                JOptionPane.PLAIN_MESSAGE,
+                JOptionPane.YES_NO_CANCEL_OPTION,
+                null,
+                options,
+                options[0]);
+        optionPane.setBorder(new EmptyBorder(12, 12, 12, 12));
+
+        final JDialog dialog = optionPane.createDialog(frame, "Filter Editor");
+//        dialog.setModal(true);
+//        dialog.setResizable(true);
+        dialog.pack();
+        dialog.setVisible(true);
+
+        Object result = optionPane.getValue();
[…]

### 2017-02-03 src/beast/tracer/ui/panel/FilterDiscretePanel.java

@@ -1,0 +1,109 @@
+/*
+ * FilterDiscretePanel.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.panel;
+
+import javax.swing.*;
+import javax.swing.plaf.BorderUIResource;
+import java.awt.*;
+import java.awt.event.ActionEvent;
+import java.awt.event.ActionListener;
+import java.util.Set;
+import java.util.TreeSet;
+
+/**
+ * @author Walter Xie
+ */
+public class FilterDiscretePanel extends FilterAbstractPanel {
+    JList allValues;
+    JList selectedValues;
+//        JButton selectButton;
+
+    FilterDiscretePanel(Set<String> allValuesSet, String[] selectedValuesArray) {
+        JPanel panel = new JPanel(new FlowLayout(FlowLayout.CENTER));
+
+        allValues = new JList(allValuesSet.toArray());
+        allValues.setVisibleRowCount(8);
+        allValues.setFixedCellWidth(80);
+        allValues.setFixedCellHeight(20);
+        allValues.setSelectionMode(ListSelectionModel.MULTIPLE_INTERVAL_SELECTION);
+        panel.add(new JScrollPane(allValues));
+
+//        if (selectedValuesArray != null) {
+//            int[] indices = new int[selectedValuesArray.length];
+//            for (int i = 0; i < indices.length; i++) {
+//                for (int j = 0; j < allValuesSet.length; j++) {
[…]

### 2017-02-03 src/beast/tracer/ui/panel/FrequencyPanel.java

@@ -1,0 +1,241 @@
+/*
+ * FrequencyPanel.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.panel;
+
+import beast.tracer.trace.*;
+import beast.tracer.ui.chart.*;
+import jam.framework.Exportable;
+
+import javax.swing.*;
+import java.awt.*;
+import java.awt.event.ActionEvent;
+import java.util.ArrayList;
+import java.util.HashMap;
+import java.util.List;
+import java.util.Map;
+
+
+/**
+ * A panel that displays frequency distributions of traces
+ *
+ * @author Andrew Rambaut
+ * @author Alexei Drummond
+ * @version $Id: FrequencyPanel.java,v 1.1.1.2 2006/04/25 23:00:09 rambaut Exp $
+ */
+public class FrequencyPanel extends JPanel implements Exportable {
+
+    private TraceList traceList = null;
+    private String traceName = null;
+
+    private class Settings {
+        ChartSetupDialog chartSetupDialog = null;
+        int minimumBins = 50;
+    }
+
+    private Settings currentSettings = new Settings();
+    private Map<String, Settings> settingsMap = new HashMap<String, Settings>();
+
+    JComboBox binsCombo = new JComboBox(
+           new Integer[]{10, 20, 50, 100, 200, 500, 1000});
+    JLabel labelBins;
+
[…]

### 2017-02-03 src/beast/tracer/ui/panel/IntervalsPanel.java

@@ -1,0 +1,106 @@
+/*
+ * IntervalsPanel.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.panel;
+
+import beast.tracer.ui.chart.JChartPanel;
+import beast.tracer.ui.chart.LinearAxis;
+import beast.tracer.trace.TraceDistribution;
+import beast.tracer.trace.TraceList;
+
+import javax.swing.*;
+import java.awt.*;
+
+import jam.framework.Exportable;
+
+
+/**
+ * A panel that displays frequency distributions of traces
+ *
+ * @author Andrew Rambaut
+ * @author Alexei Drummond
+ * @version $Id: IntervalsPanel.java,v 1.1.1.2 2006/04/25 23:00:09 rambaut Exp $
+ */
+public class IntervalsPanel extends JPanel implements Exportable {
+
+    private JIntervalsChart intervalsChart = new JIntervalsChart(new LinearAxis());
+    private JChartPanel chartPanel = new JChartPanel(intervalsChart, null, "", "");
+
+    /**
+     * Creates new IntervalsPanel
+     */
+    public IntervalsPanel() {
+        setOpaque(false);
+        setMinimumSize(new Dimension(300, 150));
+        setLayout(new BorderLayout());
+        add(chartPanel, BorderLayout.CENTER);
+    }
+
+    public void setTraces(TraceList[] traceLists, java.util.List<String> traceNames) {
+
[…]

### 2017-02-03 src/beast/tracer/ui/panel/JIntervalsChart.java

@@ -1,0 +1,155 @@
+/*
+ * JIntervalsChart.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.panel;
+
+import beast.tracer.ui.chart.*;
+
+import java.awt.*;
+import java.awt.geom.GeneralPath;
+import java.awt.geom.Line2D;
+import java.util.ArrayList;
+
+public class JIntervalsChart extends JChart {
+
+    //private static final int MAX_POINTS = 1000;
+
+    // private boolean isLinePlot = false;
+
+    private class Interval {
+        String name;
+        double value, upper, lower;
+        boolean bold;
+
+        Interval(String name, double value, double upper, double lower, boolean bold) {
+
+            this.name = name;
+            this.value = value;
+            this.upper = upper;
+            this.lower = lower;
+            this.bold = bold;
+        }
+    }
+
+    private final ArrayList<Interval> intervals = new ArrayList<Interval>();
+
+    public JIntervalsChart(Axis yAxis) {
+        super(new DiscreteAxis(true, true), yAxis);
+    }
+
+    public JIntervalsChart(Axis xAxis, Axis yAxis) {
+        super(xAxis, yAxis);
+        intervals.clear();
+    }
+
+    public void addIntervals(String name, double mean, double upper, double lower, boolean bold) {
+
+        intervals.add(new Interval(name, mean, upper, lower, bold));
+
+        xAxis.addRange(1, intervals.size());
+        yAxis.addRange(lower, upper);
+
[…]

### 2017-02-03 src/beast/tracer/ui/panel/RawTracePanel.java

@@ -1,0 +1,532 @@
+/*
+ * RawTracePanel.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.panel;
+
+import beast.tracer.trace.Trace;
+import beast.tracer.trace.TraceDistribution;
+import beast.tracer.trace.TraceType;
+import beast.tracer.trace.TraceList;
+import beast.tracer.ui.chart.*;
+import jam.framework.Exportable;
+
+import javax.sound.sampled.AudioFormat;
+import javax.sound.sampled.AudioSystem;
+import javax.sound.sampled.LineUnavailableException;
+import javax.sound.sampled.SourceDataLine;
+import javax.swing.*;
+import java.awt.*;
+import java.awt.event.ActionEvent;
+import java.util.ArrayList;
+import java.util.HashMap;
+import java.util.List;
+import java.util.Map;
+
+/**
+ * A panel that displays information about traces
+ *
+ * @author Andrew Rambaut
+ * @author Alexei Drummond
+ * @version $Id: RawTracePanel.java,v 1.2 2006/11/30 17:39:29 rambaut Exp $
+ */
+public class RawTracePanel extends JPanel implements Exportable {
+    public static int COLOUR_BY_TRACE = 0;
+    public static int COLOUR_BY_FILE = 1;
+    public static int COLOUR_BY_ALL = 2;
+
+    private static final Color[] paints = new Color[]{
+            Color.BLACK,
[…]

### 2017-02-03 src/beast/tracer/ui/panel/SummaryStatisticsPanel.java

@@ -1,0 +1,372 @@
+/*
+ * SummaryStatisticsPanel.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.panel;
+
+import beast.tracer.trace.TraceAnalysis;
+import beast.tracer.trace.TraceCorrelation;
+import beast.tracer.trace.TraceList;
+import beast.tracer.trace.TraceType;
+import jam.framework.Exportable;
+import jam.table.TableRenderer;
+
+import javax.swing.*;
+import javax.swing.plaf.BorderUIResource;
+import javax.swing.table.AbstractTableModel;
+import java.awt.*;
+import java.awt.event.MouseEvent;
+import java.awt.event.MouseMotionAdapter;
+
+
+public class SummaryStatisticsPanel extends JPanel implements Exportable {
+
+    static final String NAME_ROW = "name";
+    static final String MEAN_ROW = "mean";
+    static final String MODE_ROW = "mode";
+    static final String STDEV_ROW = "stderr of mean";
+    static final String STDEV = "stdev";
+    static final String FREQ_MODE_ROW = "frequency of mode";
+    static final String VARIANCE_ROW = "variance";
+    //    static final String STDEV_VAR_ROW = "stderr of variance";
+    static final String GEOMETRIC_MEAN_ROW = "geometric mean";
+    static final String MEDIAN_ROW = "median";
[…]

### 2017-02-03 src/beast/tracer/ui/panel/TableRenderer.java

@@ -1,0 +1,90 @@
+/*
+ * TableRenderer.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+/*
+ * TableRenderer.java
+ *
+ * (c) 2002-2005 BEAST Development Core Team
+ *
+ * This package may be distributed under the
+ * Lesser Gnu Public Licence (LGPL)
+ */
+package beast.tracer.ui.panel;
+
+import javax.swing.*;
+import javax.swing.plaf.BorderUIResource;
+import javax.swing.table.DefaultTableCellRenderer;
+import java.awt.*;
+
+
+class TableRenderer extends DefaultTableCellRenderer {
+	Color bg1 = Color.white;
+	Color bg2 = new Color(0xEE, 0xEE, 0xFF);
+	int alignment;
+
+	public TableRenderer(int alignment, int inset) {
+		super();
+		this.alignment = alignment;
+		if (alignment == SwingConstants.LEFT) {
+			setBorder(new BorderUIResource.EmptyBorderUIResource(new java.awt.Insets(0, inset, 0, 0)));
+		} else if (alignment == SwingConstants.RIGHT) {
+			setBorder(new BorderUIResource.EmptyBorderUIResource(new java.awt.Insets(0, 0, 0, inset)));
+		}
+	}
+
+	public Component getTableCellRendererComponent(
+		JTable table, Object value, boolean isSelected,
+		boolean hasFocus, int row, int column) {
+
+		setOpaque(true);
+		setText(value.toString());
+
+
[…]

### 2017-02-03 src/beast/tracer/ui/panel/TableScrollPane.java

@@ -1,0 +1,191 @@
+/*
+ * TableScrollPane.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.panel;
+
+import beast.tracer.trace.TraceAnalysis;
+import jam.framework.Exportable;
+
+import javax.swing.*;
+import javax.swing.table.DefaultTableModel;
+import javax.swing.table.JTableHeader;
+import java.awt.*;
+
+/**
+ * A panel that displays correlation plots of 2 traces in integer and category type
+ *
+ * @author Andrew Rambaut
+ * @author Alexei Drummond
+ * @author Walter Xie
+ * @version $Id: TableScrollPane.java,v 1.1.1.2 2006/04/25 23:00:09 rambaut Exp $
+ */
+public class TableScrollPane extends JScrollPane implements Exportable {
+
+    Object[] rowNames;
+    Object[] columnNames;
+    double[][] data;
+
+    ListModel lm = new RowHeaderModel();
+
+    DataTableModel dataTableModel = new DataTableModel();
+    JTable dataTable = new JTable(dataTableModel);
+    private boolean defaultNumberFormat = true;
+
+    public TableScrollPane() {
+        super();
+        setViewportView(dataTable);
+
+        setOpaque(false);
+    }
+
+    public void setTable(Object[] rowNames, Object[] columnNames, double[][] data, boolean defaultNumberFormat) {
+        this.rowNames = rowNames;
[…]

### 2017-02-03 src/beast/tracer/ui/panel/TraceOptions.java

@@ -1,0 +1,47 @@
+/*
+ * TraceOptions.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.panel;
+
+import dr.math.distributions.KernelDensityEstimatorDistribution;
+
+/**
+ * @author Andrew Rambaut
+ * @version $Id$
+ */
+public class TraceOptions {
+
+    public enum Type {
+        REAL,
+        INTEGER,
+        CATEGORICAL
+    }
+
+    private int histogramBinCount = 50;
+    private double lowerBound = Double.NEGATIVE_INFINITY;
+    private double upperBound = Double.POSITIVE_INFINITY;
+
+    private KernelDensityEstimatorDistribution.Type kdeType = KernelDensityEstimatorDistribution.Type.GAUSSIAN;
+}

### 2017-02-03 src/beast/tracer/ui/panel/TracePanel.java

@@ -1,0 +1,132 @@
+/*
+ * TracePanel.java
+ *
+ * Copyright (c) 2002-2015 Alexei Drummond, Andrew Rambaut and Marc Suchard
+ *
+ * This file is part of BEAST.
+ * See the NOTICE file distributed with this work for additional
+ * information regarding copyright ownership and licensing.
+ *
+ * BEAST is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU Lesser General Public License as
+ * published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ *  BEAST is distributed in the hope that it will be useful,
+ *  but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *  GNU Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with BEAST; if not, write to the
+ * Free Software Foundation, Inc., 51 Franklin St, Fifth Floor,
+ * Boston, MA  02110-1301  USA
+ */
+
+package beast.tracer.ui.panel;
+
+import beast.tracer.trace.TraceList;
+import jam.framework.Exportable;
+import jam.util.IconUtils;
+
+import javax.swing.*;
+import java.awt.*;
+
+/**
+ * A panel that displays information about traces
+ *
+ * @author Alexei Drummond
+ * @author Andrew Rambaut
+ * @version $Id: TracePanel.java,v 1.2 2006/05/15 23:32:30 rambaut Exp $
+ */
+public class TracePanel extends javax.swing.JPanel implements Exportable {
+
+    private final JTabbedPane tabbedPane = new JTabbedPane();
+
+    private final SummaryStatisticsPanel summaryPanel;
+    private final DensityPanel densityPanel;
+    private final JointDensityPanel jointDensityPanel;
+    private final RawTracePanel tracePanel;
+
+    private static final boolean USE_KDE = false;
+
+
+    /**
+     * Creates new form TracePanel
+     */
+    public TracePanel(JFrame parent) {
+
+        //JFrame parent1 = parent;
+
+        Icon traceIcon = new ImageIcon(IconUtils.getImage(TracePanel.class, "../images/trace-small-icon.gif"));
+        Icon frequencyIcon = new ImageIcon(IconUtils.getImage(TracePanel.class, "../images/frequency-small-icon.gif"));
[…]

### 2017-02-03 manuscript/tracer16.tex
mv figures into manuscript, use previous figures folder for Markdown files
3 trace types & conditional posterior distribution

@@
 \section{Availability:}
-Tracer is open-source under the GNU lesser general public license and available at \url{http://beast-mcmc.googlecode.com}  and  \href{http://tree.ed.ac.uk/software/tracer}{\url{http://tree.ed.ac.uk}}.
+Tracer is open-source under the GNU lesser general public license and available at 
+\url{http://beast-mcmc.googlecode.com}  %TODO \url{http://github.com/beast-dev/tracer}
+and  \href{http://tree.ed.ac.uk/software/tracer}{\url{http://tree.ed.ac.uk}}.
 
@@
 \begin{itemize}
+\item Three trace types: real, integer, string
 \item Kernel density estimates
@@
 \begin{figure}[H]
-\includegraphics[width=.5\textwidth]{../figures/gopher-log.png}  
+\includegraphics[width=.5\textwidth]{./figures/gopher-log.png}  
 \caption{The }
@@
 \begin{figure}[H]
-\includegraphics[width=.5\textwidth]{../figures/estimates.pdf}  
+\includegraphics[width=.5\textwidth]{./figures/estimates.pdf}  
 \caption{The estimates}
@@
 \begin{figure}[H]
-\includegraphics[width=.5\textwidth]{../figures/comp-95HPD.pdf}  
+\includegraphics[width=.5\textwidth]{./figures/comp-95HPD.pdf}  
 \caption{The comparison of  95\% HPD intervals from multi-trace}
@@
 \begin{figure}[H]
-\includegraphics[width=.5\textwidth]{../figures/trace.pdf}  
+\includegraphics[width=.5\textwidth]{./figures/trace.pdf}  
 \caption{The trace}
 \label{fig:trace}
+\end{figure}
+
+\begin{figure}[H]
+\includegraphics[width=.5\textwidth]{./figures/jointPrInt.png}  
+\caption{The Joint probability table of two integer traces}
+\label{fig:int:jointpr}
 \end{figure}
@@
 \subsubsection*{Conditional posterior distribution:}
+
+Fairly general solution to looking at conditional posterior distributions;
+
+Specifically, support for BSSVS forms of model averaging, in which some parameters are only in the likelihood when their submodel is "indicated" by some indicator function that is usually a discrete, integer or boolean variable.
+In that case the posterior of the parameter should not include the states when it was sampled only in the prior because the submodel it belonged to was "turned off".
+This is relevant for EBSP, Random Local Clocks model, microsatellite model averaging and relaxed clock model averaging methods all from my group in last couple of years.
+Also relevant for BSSVS in phylogeography as well depending on how the state is logged.
 

### 2017-07-17 src/tracer/application/TracerDefaultFileMenuFactory.java
cannot cast to 'TracerFileMenuHandler' error

@@ -25,13 +25,12 @@
 
 package tracer.application;
 
-import jam.framework.MenuFactory;
 import jam.framework.AbstractFrame;
 import jam.framework.Application;
 import jam.framework.MenuBarFactory;
+import jam.framework.MenuFactory;
 
 import javax.swing.*;
-import java.awt.event.ActionEvent;
 import java.awt.event.KeyEvent;
 
 /**
@@ -71,9 +70,9 @@ public class TracerDefaultFileMenuFactory implements MenuFactory {
                 menu.add(item);
             }
 
-            item = new JMenuItem(((TracerFileExtraMenuHandler)frame).getReloadAction());
-            item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_R, MenuBarFactory.MENU_MASK));
-            menu.add(item);
+//            item = new JMenuItem(((TracerFileExtraMenuHandler)frame).getReloadAction());
+//            item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_R, MenuBarFactory.MENU_MASK));
+//            menu.add(item);
 
             menu.addSeparator();
 
@@ -81,9 +80,9 @@ public class TracerDefaultFileMenuFactory implements MenuFactory {
             item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK));
             menu.add(item);
 
-            item = new JMenuItem(((TracerFileExtraMenuHandler)frame).getFullStatistics());
-            item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK + ActionEvent.SHIFT_MASK));
-            menu.add(item);
+//            item = new JMenuItem(((TracerFileExtraMenuHandler)frame).getFullStatistics());
+//            item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK + ActionEvent.SHIFT_MASK));
+//            menu.add(item);
 
             item = new JMenuItem(((TracerFileMenuHandler)frame).getExportPDFAction());
             item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK + KeyEvent.ALT_MASK));

### 2017-07-17 src/tracer/application/TracerMacFileMenuFactory.java

@@ -73,10 +73,10 @@ public class TracerMacFileMenuFactory implements MenuFactory {
                 menu.add(item);
             }
 
-            item = new JMenuItem(((TracerFileExtraMenuHandler)frame).getReloadAction());
-            item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_R, MenuBarFactory.MENU_MASK));
-            item.setEnabled(false);
-            menu.add(item);
+//            item = new JMenuItem(((TracerFileExtraMenuHandler)frame).getReloadAction());
+//            item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_R, MenuBarFactory.MENU_MASK));
+//            item.setEnabled(false);
+//            menu.add(item);
 
             menu.addSeparator();
 
@@ -84,9 +84,9 @@ public class TracerMacFileMenuFactory implements MenuFactory {
             item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK));
             menu.add(item);
 
-            item = new JMenuItem(((TracerFileExtraMenuHandler)frame).getFullStatistics());
-            item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK + ActionEvent.SHIFT_MASK));
-            menu.add(item);
+//            item = new JMenuItem(((TracerFileExtraMenuHandler)frame).getFullStatistics());
+//            item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK + ActionEvent.SHIFT_MASK));
+//            menu.add(item);
 
             item = new JMenuItem(((TracerFileMenuHandler)frame).getExportPDFAction());
             item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK + KeyEvent.ALT_MASK));

### 2017-07-17 manuscript/tracer16.tex
update manuscript

@@
 \section{Results:}
-We describe here a software package, Tracer version 1.6, which ... 
+%We describe here a software package, Tracer version 1.6, which ... 
+We therefore develop a software package Tracer (version 1.6) for analysing the trace files generated by Bayesian MCMC runs (that is, the continuous parameter values sampled from the chain).
+It can be used to analyse runs of BEAST \citep{drummond2007beast,drummond2012bayesian}, BEAST2 \citep{bouckaert2014beast2}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes}, LAMARC \citep{kuhner2006lamarc}, Migrate \citep{beerli2006comparison} and possibly other MCMC programs.
+
 \section{Availability:}
 Tracer is open-source under the GNU lesser general public license and available at 
-\url{http://beast-mcmc.googlecode.com}  %TODO \url{http://github.com/beast-dev/tracer}
+\url{https://github.com/beast-dev/tracer} 
 and  \href{http://tree.ed.ac.uk/software/tracer}{\url{http://tree.ed.ac.uk}}.
@@
 
-Blah blah
+The design of Tracer is illustrated in Figure~\ref{fig:tracer}.
+Tracer can analyse either a single MCMC log or combine the shared traces from multiple logs.
+Traces panel provides the list of the statistical summary from the sampling parameters' values at all states logged during a MCMC process.
+
+\begin{figure}[ht]
+\includegraphics[width=.38\textwidth]{./figures/tracer.pdf}  
+\caption{Tracer design}
+\label{fig:tracer}
+\end{figure}
+
+There are 4 analysis tabs to choose from a selected parameter:
+\begin{itemize}
+\item Estimates - this shows the mean, standard deviation, confidence intervals, effective sample size, and other statistics about the selected parameter.
+A frequency distribution will also be plotted.
+\item Density - this shows the Bayesian posterior density plot for the selected parameter.
+\item Joint-Marginal - this only appears if exactly 2 parameters are chosen. It then plots one against the other to look at their joint-marginal distribution.
+\item Trace - this shows the trace of the parameter against state or generation number.
+It can be used to check mixing, choose a suitable burn-in and look for trends that might suggest problems with convergence.
+\end{itemize}
+
+Multiple parameters can be selected, which will overlay the plots for the different parameters allowing comparisons to be made, such as Figure~\ref{fig:multitrace}.
+Similarly multiple trace files can be selected as well to compare different runs.
+If multiple trace files have the same trace names then a ``Combined'' trace will automatically appear.
+This can be selected as well as the individual trace files.
+
+\begin{figure}[ht]
+\includegraphics[width=.5\textwidth]{./figures/multitrace.png}  
+\caption{Multiple parameters}
+\label{fig:multitrace}
+\end{figure}
+
+You can also select the "Demographic Analysis" from the Analysis menu - This plots the distribution of demographic population sizes over time for a number of models (constant size, exponential growth \& logistic growth) that are available in BEAST.
+This involves you selecting the traces for each parameter of the model.
+You should only select the model that was actually run under BEAST (e.g., if you ran an exponential growth model, you shouldn't plot the constant population size model).
+
+The "Analysis" menu also contains options for performing Bayesian Skyline reconstructions and for calculating Bayes Factors between runs.
+
+\begin{figure}[ht]
+%\includegraphics[width=.5\textwidth]{./figures/fluBSP.pdf}  
+\subfigure[Bayesian Skyline]{\includegraphics[width = .22\textwidth]{./figures/fluBSP.pdf}}
+\subfigure[Lineages Though Time]{\includegraphics[width = .22\textwidth]{./figures/fluLTT.pdf}}
+\caption{The plot to estimate population sizes over time from ``Flu'' data, where time starts from year 1997 at the 0 position and the unit is year.}
+\label{fig:flu}
+\end{figure}
 
@@
 
-\subsubsection*{Graphical user-interface:}
-
-Tracer can analyse either a single MCMC log or combine the shared traces from multiple logs.
-Traces panel provides the list of the statistical summary from the sampling parameters' values at all states logged during a MCMC process.
-
-There are 4 analysis tabs to choose from a selected parameter:
-
-%TODO
-Estimates - this shows the mean, stdev, confidence intervals and other statistics about the selected parameter. A frequency distribution will also be plotted.
-Density - this shows the Bayesian posterior density plot for the selected parameter.
-Joint-Marginal - this only appears if exactly 2 parameters are chosen (hold down shift to select multiple parameters). It then plots one against the other to look at their joint-marginal distribution.
-Trace - this shows the trace of the parameter against state or generation number. Use this to check mixing, choose a suitable burn-in and look for trends that might suggest problems with convergence.
-Multiple parameters can be selected by holding down the shift key.
-This will overlay the plots for the different parameters allowing comparisons to be made.
-You can also select multiple trace files as well to compare different runs.
-If multiple trace files have the same trace names then a "Combined" trace will automatically appear.
-This can be selected as well as the individual trace files.
-
 %TODO better figures
-
-\begin{figure}[H]
-\includegraphics[width=.5\textwidth]{./figures/gopher-log.png}  
-\caption{The }
-\label{fig:front:page}
-\end{figure}
 
@@
 We make the Tracer package available in both executable and source code forms.
-Tracer requires Java version 1.5 or greater and executables for Windows, Mac OS and Linux platforms are located at \url{http://beast.bio.ed.ac.uk} which serves as the main page for the package.
+Tracer requires Java version 1.6 or greater and executables for Windows, Mac OS and Linux platforms are located at \url{http://beast.bio.ed.ac.uk} which serves as the main page for the package.
 This page also links to a sizable list of self-contained, step-by-step tutorials covering basic to advance usage of Tracer to summarize posterior distributions of a large set of phylogenetic models simulated using BEAST.
@@
 
-GoogleCode houses the Tracer's version-controlled source code within \url{http://beast-mcmc.googlecode.com} and links to two GoogleGroup discussion groups related to Tracer.  
+Github houses the Tracer's version-controlled source code within \url{https://github.com/beast-dev/tracer} and links to two GoogleGroup discussion groups related to Tracer.  
 The first is the ``beast-users" group (\url{http://groups.google.com/group/beast-users}) with over 1,500 members. 
@@
 
-Tracer is a program for analysing the trace files generated by Bayesian MCMC runs (that is, the continuous parameter values sampled from the chain).
+%Tracer is a program for analysing the trace files generated by Bayesian MCMC runs (that is, the continuous parameter values sampled from the chain).
 It can be used to analyse runs of BEAST \citep{drummond2007beast,drummond2012bayesian}, BEAST2 \citep{bouckaert2014beast2}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes}, LAMARC \citep{kuhner2006lamarc}, Migrate \citep{beerli2006comparison} and possibly other MCMC programs.
@@
 
-There are 4 analysis tabs to choose from:
+%There are 4 analysis tabs to choose from:
 
-Estimates - this shows the mean, stdev, confidence intervals and other statistics about the selected parameter. A frequency distribution will also be plotted.
-Density - this shows the Bayesian posterior density plot for the selected parameter.
-Joint-Marginal - this only appears if exactly 2 parameters are chosen (hold down shift to select multiple parameters). It then plots one against the other to look at their joint-marginal distribution.
-Trace - this shows the trace of the parameter against state or generation number. Use this to check mixing, choose a suitable burn-in and look for trends that might suggest problems with convergence.
-Multiple parameters can be selected by holding down the shift key.
+%Estimates - this shows the mean, stdev, confidence intervals and other statistics about the selected parameter. A frequency distribution will also be plotted.
+%Density - this shows the Bayesian posterior density plot for the selected parameter.
+%Joint-Marginal - this only appears if exactly 2 parameters are chosen (hold down shift to select multiple parameters).
+It then plots one against the other to look at their joint-marginal distribution.
+%Trace - this shows the trace of the parameter against state or generation number. Use this to check mixing, choose a suitable burn-in and look for trends that might suggest problems with convergence.
+%Multiple parameters can be selected by holding down the shift key.
 This will overlay the plots for the different parameters allowing comparisons to be made.
@@
 
-You can also select the "Demographic Analysis" from the Analysis menu - This plots the distribution of demographic population sizes over time for a number of models (constant size, exponential growth \& logistic growth) that are available in BEAST.
+%You can also select the "Demographic Analysis" from the Analysis menu - This plots the distribution of demographic population sizes over time for a number of models (constant size, exponential growth \& logistic growth) that are available in BEAST.
 This involves you selecting the traces for each parameter of the model.
@@
 
-The "Analysis" menu also contains options for performing Bayesian Skyline reconstructions and for calculating Bayes Factors between runs.
+%The "Analysis" menu also contains options for performing Bayesian Skyline reconstructions and for calculating Bayes Factors between runs.
 

### 2017-07-20 manuscript/tracer16.tex
write half of section design and implmentation

@@
 
-The design of Tracer is illustrated in Figure~\ref{fig:tracer}.
-Tracer can analyse either a single MCMC log or combine the shared traces from multiple logs.
-Traces panel provides the list of the statistical summary from the sampling parameters' values at all states logged during a MCMC process.
+The design of Tracer is illustrated in Figure~\ref{fig:tracer}. It can analyse either a single MCMC log or combine the shared traces from multiple logs. 
+The list of traces in each log are created from the sampling parameters logged during a MCMC process, which include their statistical summaries from their values at all states logged.
 
@@
 
-There are 4 analysis tabs to choose from a selected parameter:
-\begin{itemize}
-\item Estimates - this shows the mean, standard deviation, confidence intervals, effective sample size, and other statistics about the selected parameter.
-A frequency distribution will also be plotted.
-\item Density - this shows the Bayesian posterior density plot for the selected parameter.
-\item Joint-Marginal - this only appears if exactly 2 parameters are chosen. It then plots one against the other to look at their joint-marginal distribution.
-\item Trace - this shows the trace of the parameter against state or generation number.
-It can be used to check mixing, choose a suitable burn-in and look for trends that might suggest problems with convergence.
-\end{itemize}
+Tracer can handle 3 trace types $-$ variable types \citep{mendenhall2012introduction}: real, integer, and categorical.
+The real trace is used to handle continuous quantitative variables, which is the most common type in logged parameters.
+The integer is dealing with discrete quantitative variables, and the categorical trace is for categorical qualitative variables. 
+% check about defination 
+These three types can be automatically assigned when the log file is imported. But they can also be changed later.
 
-Multiple parameters can be selected, which will overlay the plots for the different parameters allowing comparisons to be made, such as Figure~\ref{fig:multitrace}.
+Multiple traces can be selected as shown in Figure~\ref{fig:multitrace}, which will overlay the plots for the different traces allowing comparisons to be made.
 Similarly multiple trace files can be selected as well to compare different runs.
-If multiple trace files have the same trace names then a ``Combined'' trace will automatically appear.
-This can be selected as well as the individual trace files.
+If multiple trace files have the same trace names then a ``Combined'' trace will automatically appear, which can be selected as well as the individual trace files.
 
@@
 \includegraphics[width=.5\textwidth]{./figures/multitrace.png}  
-\caption{Multiple parameters}
+\caption{A comparison of multiple traces}
 \label{fig:multitrace}
 \end{figure}
+
+There are four analysis tabs to choose from either a selected trace (sampling parameter) or multiple traces:
+
+\begin{figure}[ht]
+\subfigure[Estimates]{\includegraphics[width = .22\textwidth, height = 3cm]{./figures/frequency.pdf}}
+\subfigure[Density]{\includegraphics[width = .22\textwidth, height = 3cm]{./figures/multiKDE.pdf}}\\
+\subfigure[Joint-Marginal]{\includegraphics[width = .22\textwidth, height = 3cm]{./figures/joint-marginal.pdf}}
+\subfigure[Trace]{\includegraphics[width = .22\textwidth, height = 3cm]{./figures/trace.pdf}}
+\caption{Image exported from Tracer's four analysis tabs}
+\label{fig:4tabs}
+\end{figure}
+
+\begin{itemize}
+\item Estimates - this shows the mean, standard deviation, confidence intervals, effective sample size, and other statistics about the selected trace(s).
+A frequency distribution will also be plotted for a single selected trace, such as Figure~\ref{fig:4tabs}.a.
+Interval bars for quantitative traces (real or integer) or violin plots for categorical traces will be drawn for a comparison, such as Figure~\ref{fig:multitrace}, if multiple traces are selected.
+
+\item Density - this shows the Bayesian posterior density plot for the selected trace(s), if they are continuous, the kernel density estimates will be available, for example, in Figure~\ref{fig:4tabs}.b.
+
+\item Joint-Marginal - this only appears if exactly 2 traces are chosen. It then plots one against the other to look at their joint-marginal distribution, for example, in Figure~\ref{fig:4tabs}.c.
+
+% trace is duplicate to trace, use trajectory?
+\item Trace - this shows the trajectory of the selected trace(s) against state or generation number as it can be seen in Figure~\ref{fig:4tabs}.d.
+It can be used to check mixing, choose a suitable burn-in and look for trends that might suggest problems with convergence.
+
+\end{itemize}
+
+Tracer also provides demographic population analysis given a log or its tree log,
+For example, ``Demographic Analysis" plots the distribution of demographic population sizes over time for a number of models, such as constant size, exponential growth, and logistic growth, all of which are available in BEAST.
 
@@
 
+%Model selection HME and AICM have been removed 
 New features:
@@
 \item Kernel density estimates
-\item HME
-\item AICM
+\item Conditional posterior distribution
+%\item HME
+%\item AICM
 \end{itemize}
@@
 %TODO better figures
-
-\begin{figure}[H]
-\includegraphics[width=.5\textwidth]{./figures/estimates.pdf}  
-\caption{The estimates}
-\label{fig:estimates}
-\end{figure}
 
@@
 \begin{figure}[H]
-\includegraphics[width=.5\textwidth]{./figures/trace.pdf}  
-\caption{The trace}
-\label{fig:trace}
-\end{figure}
-
-\begin{figure}[H]
 \includegraphics[width=.5\textwidth]{./figures/jointPrInt.png}  
@@
 
+%\begin{figure}[ht]
+%\includegraphics[width=.5\textwidth]{./figures/frequency.pdf}  
+%\caption{A frequency distribution for a continuous trace}
+%\label{fig:freq}
+%\end{figure}
+%\begin{figure}[ht]
+%\includegraphics[width=.5\textwidth]{./figures/multiKDE.pdf}  
+%\caption{KDE of multiple traces}
+%\label{fig:multiKDE}
+%\end{figure}
+%\begin{figure}[ht]
+%\includegraphics[width=.5\textwidth]{./figures/joint-marginal.pdf}  
+%\caption{The joint-marginal distribution of two selected traces}
+%\label{fig:trace}
+%\end{figure}
+%\begin{figure}[ht]
+%\includegraphics[width=.5\textwidth]{./figures/trace.pdf}  
+%\caption{The selected traces against state}
+%\label{fig:trace}
+%\end{figure}
+
 \subsubsection*{Diagnostics:} Blah blah
@@
 
-\subsubsection*{Model selection:}
+%\subsubsection*{Model selection:}
 
@@
 
-\section*{Taken from the Tracer website}
+%\section*{Taken from the Tracer website}
 
@@
 
-Although Tracer can be used with programs other than BEAST, users may find it useful to join the BEAST users mailing list.
+%Although Tracer can be used with programs other than BEAST, users may find it useful to join the BEAST users mailing list.
 This is used to announce new versions and advise users about bugs and problems.
 
-You can join the mailing list here:
-http://groups.google.com/group/beast-users
+%You can join the mailing list here:
+%http://groups.google.com/group/beast-users
 
-The website for BEAST (and Tracer) is here:
-http://beast.bio.ed.ac.uk/
+%The website for BEAST (and Tracer) is here:
+%http://beast.bio.ed.ac.uk/
 
-At present there is no detailed manual for this application, you will simply have to play around and see what happens.
+%At present there is no detailed manual for this application, you will simply have to play around and see what happens.
 Basically you can select the trace file in the top left of the window, the individual parameter in the bottom left and the analysis appears on the right.
@@
 
-The "Print" function in the "File" menu will print the current graph or table and the "Export Data" function can be used to export the data from the plots for use in another graphic package.
+%The "Print" function in the "File" menu will print the current graph or table and the "Export Data" function can be used to export the data from the plots for use in another graphic package.
 
-To export the currently displayed graphic use the "Export PDF" function in the "File" menu.
+%To export the currently displayed graphic use the "Export PDF" function in the "File" menu.
 

### 2017-07-21 manuscript/tracer16.tex
complete design and implmentation

@@
 \includegraphics[width=.38\textwidth]{./figures/tracer.pdf}  
-\caption{Tracer design}
+\caption{The design of Tracer}
 \label{fig:tracer}
@@
 
-Tracer can handle 3 trace types $-$ variable types \citep{mendenhall2012introduction}: real, integer, and categorical.
-The real trace is used to handle continuous quantitative variables, which is the most common type in logged parameters.
-The integer is dealing with discrete quantitative variables, and the categorical trace is for categorical qualitative variables. 
+Tracer can handle 3 trace types, also known as variable types \citep{mendenhall2012introduction}: real, integer, and categorical.
+The real trace is used to handle quantitative continuous variables, which is the most common type in logged parameters.
+The integer is dealing with quantitative discrete variables, and the categorical trace is for qualitative categorical variables. 
 % check about defination 
@@
 \begin{figure}[ht]
-\subfigure[Estimates]{\includegraphics[width = .22\textwidth, height = 3cm]{./figures/frequency.pdf}}
-\subfigure[Density]{\includegraphics[width = .22\textwidth, height = 3cm]{./figures/multiKDE.pdf}}\\
-\subfigure[Joint-Marginal]{\includegraphics[width = .22\textwidth, height = 3cm]{./figures/joint-marginal.pdf}}
-\subfigure[Trace]{\includegraphics[width = .22\textwidth, height = 3cm]{./figures/trace.pdf}}
+\subfigure[Estimates]{\includegraphics[width = .223\textwidth, height = 3cm]{./figures/frequency.pdf}}
+\subfigure[Density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/multiKDE.pdf}}\\
+\subfigure[Joint-Marginal]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/joint-marginal.pdf}}
+\subfigure[Trace]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/trace.pdf}}
 \caption{Image exported from Tracer's four analysis tabs}
@@
 
-Tracer also provides demographic population analysis given a log or its tree log,
-For example, ``Demographic Analysis" plots the distribution of demographic population sizes over time for a number of models, such as constant size, exponential growth, and logistic growth, all of which are available in BEAST.
+In addition to illustrating joint-marginal distribution between two continuous traces, we newly invented a TangHuLu chart to visualise the joint probability between two integer or categorical traces using coloured bubbles, as shown in Figure~\ref{fig:tanghulu}.a.
+The size of circle is proportional to the joint probability, the blue coloured circle is in the credible set of given a probability threshold default to 0.95, the red is in the set not in the credible set.
+The tile background is coloured if there is any circle, in case the very small circle wouldn't be missed out.
+The coloured background can also help to show the area of the credible set and non-credible set.
 
-You can also select the "Demographic Analysis" from the Analysis menu - This plots the distribution of demographic population sizes over time for a number of models (constant size, exponential growth \& logistic growth) that are available in BEAST.
-This involves you selecting the traces for each parameter of the model.
-You should only select the model that was actually run under BEAST (e.g., if you ran an exponential growth model, you shouldn't plot the constant population size model).
+Furthermore, box-and-whisker plots are used to display the joint-marginal distribution between one continuous trace and one integer or categorical trace as shown in Figure~\ref{fig:tanghulu}.b.
 
-The "Analysis" menu also contains options for performing Bayesian Skyline reconstructions and for calculating Bayes Factors between runs.
+\begin{figure}[ht]
+\subfigure[TangHuLu chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/tanghulu.pdf}}
+\subfigure[boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/boxplot.pdf}}
+\caption{Visualisation of joint-marginal distribution (a) between two integer or categorical traces using TangHuLu chart, and (b) between one continuous trace and one integer or categorical trace using box-and-whisker plot.}
+\label{fig:tanghulu}
+\end{figure}
+
+Tracer also provides demographic reconstruction resulting in a graphical plot to estimate the distribution of effective population sizes over time, given a MCMC log from a Bayesian framework or additionally with the tree log depending on the model.
+For example, ``Demographic Reconstruction" performs reconstructions of the distribution for a number of models, such as constant size, exponential growth \citep{drummond2002estimating}, and logistic growth.
+%citation?
+Besides ``Bayesian Skyline Reconstruction'' reconstructs a generalised Bayesian skyline plot (e.g.
+Figure~\ref{fig:flu}.a) describing demographic history under canonical scenarios \citep{drummond2005bayesian},
+``GMRF Skyride Reconstruction'' and ``SkyGrid Reconstruction'' are respectively available for a temporal smoothing method using the Gaussian Markov random field (GMRF) model  \citep{minin2008smooth} and a generalisation of the GMRF model \citep{gill2012improving}.
+``Lineages Through Time'' plots quantiles of the number of lineages against time, which describes the change of diversification of all organisms over time. %citation?
+All these models are available in BEAST \citep{drummond2007beast}.
 
@@
 \subfigure[Bayesian Skyline]{\includegraphics[width = .22\textwidth]{./figures/fluBSP.pdf}}
-\subfigure[Lineages Though Time]{\includegraphics[width = .22\textwidth]{./figures/fluLTT.pdf}}
-\caption{The plot to estimate population sizes over time from ``Flu'' data, where time starts from year 1997 at the 0 position and the unit is year.}
+\subfigure[Lineages Through Time]{\includegraphics[width = .22\textwidth]{./figures/fluLTT.pdf}}
+\caption{The plot (a) performs Bayesian Skyline reconstruction to estimate population sizes over time and (b) Lineages Through Time , from ``Flu'' data, where time starts from year 1997 at the 0 position and the unit is year.}
 \label{fig:flu}
 \end{figure}
+
+Tracer recently offers a fairly general solution of looking at conditional posterior distributions.
+This, specifically, supports for BSSVS forms of model averaging, in which some parameters are only in the likelihood when their sub-model is ``indicated" by some indicator function that is usually a discrete, integer or boolean variable.
+In that case the posterior of the parameter should not include the states when it was sampled only in the prior because the sub-model it belonged to was ``turned off".
+This is relevant for EBSP, Random Local Clocks model, microsatellite model averaging and relaxed clock model averaging methods all from my group in last couple of years.
+It is also relevant for BSSVS in phylogeography as well depending on how the state is logged.
+
+% TODO more ?
+
+\section*{Example}
+
+Even users of \textsc{MrBayes} and \textsc{RevBayes} find Tracer handy.
+
+\section*{Availability and Future Directions}
+
+We make the Tracer package available in both executable and source code forms.
+Tracer requires Java version 1.6 or greater and executables for Windows, Mac OS and Linux platforms are located at \url{http://beast.bio.ed.ac.uk} which serves as the main page for the package.
+This page also links to a sizable list of self-contained, step-by-step tutorials covering basic to advance usage of Tracer to summarize posterior distributions of a large set of phylogenetic models simulated using BEAST.
+For example, popular tutorials describe how to use Tracer to generate marginal parameter summaries and infer population dynamics trajectories over time.
+
+Github houses the Tracer's version-controlled source code within \url{https://github.com/beast-dev/tracer} and links to two GoogleGroup discussion groups related to Tracer.  
+The first is the ``beast-users" group (\url{http://groups.google.com/group/beast-users}) with over 1,500 members. 
+%At the time of writing, forty-seven developers belong to the ``beast-dev" group that facilitates BEAST development across three continents.
+
+Future development directions for Tracer focus on \ldots
+
+\section*{Old Text}
+
+%You can also select the "Demographic Analysis" from the Analysis menu - This plots the distribution of demographic population sizes over time for a number of models (constant size, exponential growth \& logistic growth) that are available in BEAST.
+This involves you selecting the traces for each parameter of the model.
+You should only select the model that was actually run under BEAST (e.g., if you ran an exponential growth model, you shouldn't plot the constant population size model).
+
+%The "Analysis" menu also contains options for performing Bayesian Skyline reconstructions and for calculating Bayes Factors between runs.
+
+Feature list:
+\begin{itemize}
+\item KDEs
+\item cheap marginal likelihood estimators
+\item demographic trajectory reconstruction
+\end{itemize}
 
@@
 \end{itemize}
-
-%TODO better figures
 
@@
 
-Fairly general solution to looking at conditional posterior distributions;
+%Fairly general solution to looking at conditional posterior distributions;
 
-Specifically, support for BSSVS forms of model averaging, in which some parameters are only in the likelihood when their submodel is "indicated" by some indicator function that is usually a discrete, integer or boolean variable.
+%Specifically, support for BSSVS forms of model averaging, in which some parameters are only in the likelihood when their submodel is "indicated" by some indicator function that is usually a discrete, integer or boolean variable.
 In that case the posterior of the parameter should not include the states when it was sampled only in the prior because the submodel it belonged to was "turned off".
@@
 Also relevant for BSSVS in phylogeography as well depending on how the state is logged.
-
-\section*{Example}
-
-Even users of \textsc{MrBayes} and \textsc{RevBayes} find Tracer handy.
-
-\section*{Availability and Future Directions}
-
-We make the Tracer package available in both executable and source code forms.
-Tracer requires Java version 1.6 or greater and executables for Windows, Mac OS and Linux platforms are located at \url{http://beast.bio.ed.ac.uk} which serves as the main page for the package.
-This page also links to a sizable list of self-contained, step-by-step tutorials covering basic to advance usage of Tracer to summarize posterior distributions of a large set of phylogenetic models simulated using BEAST.
-For example, popular tutorials describe how to use Tracer to generate marginal parameter summaries and infer population dynamics trajectories over time.
-
-Github houses the Tracer's version-controlled source code within \url{https://github.com/beast-dev/tracer} and links to two GoogleGroup discussion groups related to Tracer.  
-The first is the ``beast-users" group (\url{http://groups.google.com/group/beast-users}) with over 1,500 members. 
-%At the time of writing, forty-seven developers belong to the ``beast-dev" group that facilitates BEAST development across three continents.
-
-Future development directions for Tracer focus on \ldots
-
-Feature list:
-\begin{itemize}
-\item KDEs
-\item cheap marginal likelihood estimators
-\item demographic trajectory reconstruction
-\end{itemize}
-
 %\section*{Taken from the Tracer website}

### 2017-07-21 src/tracer/traces/JointDensityPanel.java
the axis label of box plot is other way around #81

@@ -274,6 +274,12 @@ public class JointDensityPanel extends NTracesChartPanel {
                     pointsCheckBox.setVisible(false);
                     translucencyCheckBox.setVisible(false);
                 }
+
+                if (!td1.getTraceType().isIntegerOrBinary()) { // do not swap name for TangHuLu plot
+                    String swapName = name1;
+                    name1 = name2;
+                    name2 = swapName;
+                }
             }
         }
         setXLab(name1);

### 2017-07-24 manuscript/tracer16.tex
add introduction

@@
 %We describe here a software package, Tracer version 1.6, which ...
-We therefore develop a software package Tracer (version 1.6) for analysing the trace files generated by Bayesian MCMC runs (that is, the continuous parameter values sampled from the chain).
-It can be used to analyse runs of BEAST \citep{drummond2007beast,drummond2012bayesian}, BEAST2 \citep{bouckaert2014beast2}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes}, LAMARC \citep{kuhner2006lamarc}, Migrate \citep{beerli2006comparison} and possibly other MCMC programs.
 
@@
 Tracer is open-source under the GNU lesser general public license and available at 
-\url{https://github.com/beast-dev/tracer} 
-and  \href{http://tree.ed.ac.uk/software/tracer}{\url{http://tree.ed.ac.uk}}.
+\url{http://beast.bio.ed.ac.uk/Tracer} and \url{https://github.com/beast-dev/tracer}.
 
@@
 
-Blah blah
+Bayesian inference of phylogeny \citep{huelsenbeck2001bayesian, felsenstein2004inferring,yang2006computational} has been a popular approach to reconstruct the evolutionary history and discover the evolutionary relationships among organisms that could come from genes, genomes, individuals or species.
+The statistical framework is based on the posterior probability of phylogenetic trees given the data, and currently takes the advantage of Markov chain Monte Carlo (MCMC) to promote the computational speed of Bayesian estimation to reach the approximated optimisation.
+The posterior summaries, as well as as scientifically relevant marginalisation, %TODO
+are therefore critical important to understand a Bayesian phylogenetic analysis.
+%However, the full posterior distribution is high-dimensional and often difficult to visualise. 
+We then develop a software package Tracer (version 1.6) for visualising and analysing the trace files generated by Bayesian MCMC runs that is the parameter values sampled from the chain. 
+It can be used to analyse runs of BEAST \citep{drummond2007beast,drummond2012bayesian}, BEAST2 \citep{bouckaert2014beast2}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes}, LAMARC \citep{kuhner2006lamarc}, Migrate \citep{beerli2006comparison},  and possibly other MCMC programs.
+In addition, Tracer will be able to reconstruct the distribution of effective population sizes over time if the population genetic model was involved in the analysis.
 
-\begin{itemize}
-\item Importance of posterior summaries (as scientifically relevant marginalization of the full posterior distribution)
-\item Full posterior distribution is high-dimensional and (often) difficult to visualize
-\item Posterior simulator agnostic -- highly used with \textsc{MrBayes} as well
-\end{itemize}
+%\begin{itemize}
+%\item Importance of posterior summaries (as scientifically relevant marginalization of the full posterior distribution)
+%\item Full posterior distribution is high-dimensional and (often) difficult to visualize
+%\item Posterior simulator agnostic -- highly used with \textsc{MrBayes} as well
+%\end{itemize}
 
@@
 
-The design of Tracer is illustrated in Figure~\ref{fig:tracer}. It can analyse either a single MCMC log or combine the shared traces from multiple logs. 
+Tracer can analyse either a single MCMC log or combine the shared traces from multiple logs as designed in Figure~\ref{fig:tracer}. 
 The list of traces in each log are created from the sampling parameters logged during a MCMC process, which include their statistical summaries from their values at all states logged. 
+The effective sample size (ESS) is one of important measurements of a sampled continuous parameter, which tells the number of effectively independent draws from the posterior distribution that the Markov chain is equivalent to.
 
@@
 
-\begin{figure}[ht]
-\subfigure[Estimates]{\includegraphics[width = .223\textwidth, height = 3cm]{./figures/frequency.pdf}}
-\subfigure[Density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/multiKDE.pdf}}\\
-\subfigure[Joint-Marginal]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/joint-marginal.pdf}}
-\subfigure[Trace]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/trace.pdf}}
-\caption{Image exported from Tracer's four analysis tabs}
-\label{fig:4tabs}
-\end{figure}
-
 \begin{itemize}
-\item Estimates - this shows the mean, standard deviation, confidence intervals, effective sample size, and other statistics about the selected trace(s).
+\item Estimates - this shows the mean, standard deviation, confidence intervals, ESS, and other statistics about the selected trace(s).
 A frequency distribution will also be plotted for a single selected trace, such as Figure~\ref{fig:4tabs}.a.
@@
 
-\begin{figure}[ht]
-%\includegraphics[width=.5\textwidth]{./figures/fluBSP.pdf}  
-\subfigure[Bayesian Skyline]{\includegraphics[width = .22\textwidth]{./figures/fluBSP.pdf}}
-\subfigure[Lineages Through Time]{\includegraphics[width = .22\textwidth]{./figures/fluLTT.pdf}}
-\caption{The plot (a) performs Bayesian Skyline reconstruction to estimate population sizes over time and (b) Lineages Through Time , from ``Flu'' data, where time starts from year 1997 at the 0 position and the unit is year.}
-\label{fig:flu}
-\end{figure}
-
 Tracer recently offers a fairly general solution of looking at conditional posterior distributions.
@@
 
-Even users of \textsc{MrBayes} and \textsc{RevBayes} find Tracer handy.
+%Even users of \textsc{MrBayes} and \textsc{RevBayes} find Tracer handy.
+
+\subsection*{Divergence Dating}
+
+\begin{figure}[ht]
+\subfigure[Estimates]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/frequency.pdf}}
+\subfigure[Density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/multiKDE.pdf}}\\
+\subfigure[Joint-Marginal]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/joint-marginal.pdf}}
+\subfigure[Trace]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/trace.pdf}}
+\caption{Image exported from four analysis tabs in Tracer which is used in the ``Divergence Dating'' tutorial.}
+\label{fig:4tabs}
+\end{figure}
+
+\subsection*{Virus Practical}
+
+\begin{figure}[ht]
+%\includegraphics[width=.5\textwidth]{./figures/fluBSP.pdf}  
+\subfigure[Bayesian Skyline]{\includegraphics[width = .23\textwidth]{./figures/fluBSP.pdf}}
+\subfigure[Lineages Through Time]{\includegraphics[width = .23\textwidth]{./figures/fluLTT.pdf}}
+\caption{The plot (a) performs Bayesian Skyline reconstruction to estimate population sizes over time and (b) Lineages Through Time , from ``Flu'' data, where time starts from year 1997 at the 0 position and the unit is year.}
+\label{fig:flu}
+\end{figure}
 
@@
 We make the Tracer package available in both executable and source code forms.
-Tracer requires Java version 1.6 or greater and executables for Windows, Mac OS and Linux platforms are located at \url{http://beast.bio.ed.ac.uk} which serves as the main page for the package.
-This page also links to a sizable list of self-contained, step-by-step tutorials covering basic to advance usage of Tracer to summarize posterior distributions of a large set of phylogenetic models simulated using BEAST.
+Tracer requires Java version 1.6 or greater and executables for Windows, Mac OS and Linux platforms are located at \url{http://beast.bio.ed.ac.uk/Tracer} % \url{https://github.com/beast-dev/tracer} ?
+which serves as the main page for the package.
+This page also links to a sizeable list of self-contained, step-by-step tutorials covering basic to advance usage of Tracer to summarise posterior distributions of a large set of phylogenetic models simulated using BEAST.
 For example, popular tutorials describe how to use Tracer to generate marginal parameter summaries and infer population dynamics trajectories over time.
 
-Github houses the Tracer's version-controlled source code within \url{https://github.com/beast-dev/tracer} and links to two GoogleGroup discussion groups related to Tracer.  
-The first is the ``beast-users" group (\url{http://groups.google.com/group/beast-users}) with over 1,500 members. 
+Github houses the Tracer's source code within \url{https://github.com/beast-dev/tracer} ,and links to a GoogleGroup discussion group related to Tracer,  
+which is the ``beast-users" group (\url{http://groups.google.com/group/beast-users}) with over 1,500 members. 
 %At the time of writing, forty-seven developers belong to the ``beast-dev" group that facilitates BEAST development across three continents.

### 2017-07-25 manuscript/tracer16.tex
add example 1

@@
 \documentclass{bioinfo}
-\copyrightyear{2012}
-\pubyear{2012}
+\copyrightyear{2017}
+\pubyear{2017}
 
@@
 % Title must be 150 characters or less
-\title[Tracer 1.6]{Bayesian MCMC diagnostics and summarization using Tracer 1.6}
+\title[Tracer 1.6]{Bayesian MCMC diagnostics and summarisation using Tracer 1.6}
 
-\author[Rambaut \textit{et~al}]{ Andrew Rambaut\,$^{1}$, Alexei J.~Drummond\,$^{2,3}$, Dong Xie\,$^{2,3}$, Marc A.~Suchard\,$^{4,5,6}$}
+\author[Rambaut \textit{et~al}]{ Andrew Rambaut\,$^{1}$, Alexei J.~Drummond\,$^{2,3}$, Dong Xie\,$^{2,3}$, Marc A.~Suchard\,$^{4,5}$}
 
@@
 $^{3}$Centre for Computational Evolution, University of Auckland, Auckland, NZ\\
-$^{4,5}$Departments of Biomathematics and Human Genetics, David Geffen School of Medicine at UCLA, and \\
-$^{6}$Department of Biostatistics, UCLA Fielding School of Public Health, University of California, Los Angeles, USA \\
+$^{4}$Departments of Biomathematics and Human Genetics, David Geffen School of Medicine, UCLA, and \\
+$^{5}$Department of Biostatistics, UCLA Fielding School of Public Health, University of California, Los Angeles, USA \\
 }
@@
 \section{Motivation:}
-Computational evolutionary biology, statistical phylogenetics and coalescent-based population genetics are becoming increasingly central to the analysis and understanding of molecular sequence data. 
+%Computational evolutionary biology, statistical phylogenetics and coalescent-based population genetics are becoming increasingly central to the analysis and understanding of molecular sequence data. 
+Bayesian inference of phylogeny using Markov chain Monte Carlo (MCMC) plays a central role for understanding the story of evolutions from molecular sequence data.  
+The comprehension of posterior distributions from an implementation of this framework is the significant matter.
 \section{Results:}
 %We describe here a software package, Tracer version 1.6, which ... 
-
+We here present a software package \emph{Tracer} (version 1.6) for visualising and analysing the trace files generated by Bayesian MCMC runs. 
+It also provides kernel density estimation, cheap marginal likelihood estimator, 
+multivariate analyses (is this reluctant?), %TODO
+demographic trajectory reconstruction, conditional posterior distribution summary, etc. 
 \section{Availability:}
@@
 Bayesian inference of phylogeny \citep{huelsenbeck2001bayesian, felsenstein2004inferring,yang2006computational} has been a popular approach to reconstruct the evolutionary history and discover the evolutionary relationships among organisms that could come from genes, genomes, individuals or species.
-The statistical framework is based on the posterior probability of phylogenetic trees given the data, and currently takes the advantage of Markov chain Monte Carlo (MCMC) to promote the computational speed of Bayesian estimation to reach the approximated optimisation.
+The statistical framework is based on the posterior probability of phylogenetic trees given the data, and currently takes the advantage of Markov chain Monte Carlo (MCMC) to promote the computational speed of Bayesian estimation to reach an approximated optimisation.
 The posterior summaries, as well as as scientifically relevant marginalisation, %TODO
-are therefore critical important to understand a Bayesian phylogenetic analysis.
+are therefore critically important for understanding a Bayesian phylogenetic analysis.
 %However, the full posterior distribution is high-dimensional and often difficult to visualise. 
-We then develop a software package Tracer (version 1.6) for visualising and analysing the trace files generated by Bayesian MCMC runs that is the parameter values sampled from the chain. 
+We then develop a software package Tracer (version 1.6) for this purpose, which analyses Bayesian MCMC log files containing the list of parameters whose values sampled from the chain. 
 It can be used to analyse runs of BEAST \citep{drummond2007beast,drummond2012bayesian}, BEAST2 \citep{bouckaert2014beast2}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes}, LAMARC \citep{kuhner2006lamarc}, Migrate \citep{beerli2006comparison},  and possibly other MCMC programs.
-In addition, Tracer will be able to reconstruct the distribution of effective population sizes over time if the population genetic model was involved in the analysis.
+In addition, Tracer will be able to redraw the distribution of effective population sizes over time if a demographic model was involved in the analysis.
 
@@
 
-Tracer can analyse either a single MCMC log or combine the shared traces from multiple logs as designed in Figure~\ref{fig:tracer}. 
+\begin{figure}[ht]
+\includegraphics[width=.38\textwidth]{./figures/tracer.pdf}  
+\caption{The system design of Tracer 1.6}
+\label{fig:tracer}
+\end{figure}
+
+Tracer is able to analyse either a single MCMC log or combine the shared traces from multiple logs as designed in Figure~\ref{fig:tracer}. 
 The list of traces in each log are created from the sampling parameters logged during a MCMC process, which include their statistical summaries from their values at all states logged. 
@@
 
-\begin{figure}[ht]
-\includegraphics[width=.38\textwidth]{./figures/tracer.pdf}  
-\caption{The design of Tracer}
-\label{fig:tracer}
-\end{figure}
-
-Tracer can handle 3 trace types, also known as variable types \citep{mendenhall2012introduction}: real, integer, and categorical.
+Tracer handles three trace types, also known as variable types \citep{mendenhall2012introduction}: real, integer, and categorical.
 The real trace is used to handle quantitative continuous variables, which is the most common type in logged parameters.
 The integer is dealing with quantitative discrete variables, and the categorical trace is for qualitative categorical variables. 
-% check about defination 
+% check about definition 
 These three types can be automatically assigned when the log file is imported. But they can also be changed later.
 
-Multiple traces can be selected as shown in Figure~\ref{fig:multitrace}, which will overlay the plots for the different traces allowing comparisons to be made.
-Similarly multiple trace files can be selected as well to compare different runs.
+The multiple traces can be selected as shown in Figure~\ref{fig:multitrace}, which will overlay the plots for the different traces allowing comparisons to be made.
+The multiple trace files can be similarly selected as well to compare different runs.
 If multiple trace files have the same trace names then a ``Combined'' trace will automatically appear, which can be selected as well as the individual trace files.
@@
 
+%TODO simplify, some duplication to 1st example
 \begin{itemize}
 \item Estimates - this shows the mean, standard deviation, confidence intervals, ESS, and other statistics about the selected trace(s).
-A frequency distribution will also be plotted for a single selected trace, such as Figure~\ref{fig:4tabs}.a.
+A histogram presenting the frequency distribution over values will also be plotted for a single selected trace, such as Figure~\ref{fig:4tabs}.a.
 Interval bars for quantitative traces (real or integer) or violin plots for categorical traces will be drawn for a comparison, such as Figure~\ref{fig:multitrace}, if multiple traces are selected.
@@
 
-\item Joint-Marginal - this only appears if exactly 2 traces are chosen. It then plots one against the other to look at their joint-marginal distribution, for example, in Figure~\ref{fig:4tabs}.c.
+\item Joint-Marginal - this only appears if exactly two traces are chosen. It then plots one against the other to look at their joint-marginal distribution, for example, in Figure~\ref{fig:4tabs}.c.
 
@@
 Tracer also provides demographic reconstruction resulting in a graphical plot to estimate the distribution of effective population sizes over time, given a MCMC log from a Bayesian framework or additionally with the tree log depending on the model.
-For example, ``Demographic Reconstruction" performs reconstructions of the distribution for a number of models, such as constant size, exponential growth \citep{drummond2002estimating}, and logistic growth.
+For example, ``Demographic Reconstruction" performs reconstructions of the distribution for a number of models, such as constant size \citep{drummond2002estimating}, exponential growth \citep{drummond2002estimating}, and logistic growth (?).
 %citation?
@@
 ``GMRF Skyride Reconstruction'' and ``SkyGrid Reconstruction'' are respectively available for a temporal smoothing method using the Gaussian Markov random field (GMRF) model  \citep{minin2008smooth} and a generalisation of the GMRF model \citep{gill2012improving}.
-``Lineages Through Time'' plots quantiles of the number of lineages against time, which describes the change of diversification of all organisms over time. %citation?
+``Lineages Through Time'' (?) plots quantiles of the number of lineages against time, which describes the change of diversification of all organisms over time. %citation?
 All these models are available in BEAST \citep{drummond2007beast}.
 
+%TODO better to re-write
 Tracer recently offers a fairly general solution of looking at conditional posterior distributions.
@@
 
-\subsection*{Divergence Dating}
+\subsection*{Calibrating the tree to estimate TMRCA}
+
+Figure~\ref{fig:4tabs} demonstrates a snapshot of posterior summaries in Tracer for a BEAST \citep{drummond2007beast} run created from a 898 bp alignment of sequences of 12 species of primates (http://beast.bio.ed.ac.uk/tutorials?).
+%data citation
+%using HKY + $\Gamma$ model, uncorrelated lognormal relaxed molecular clock, and Yule prior on the tree.
+In this analysis, we furthermore specify three calibrations on the phylogenetic tree to estimate time to the most recent common ancestor (TMRCA) based on our prior fossil knowledge.
+
+The first TMRCA prior, called ``ingroup'', includes all species except the \textit{Lemur} which will form the outgroup in the tree. %monophyletic
+The second group is ``Human-Chimp'' that contains only \textit{Homo sapiens} and \textit{Pan}. 
+The last consists of everything except \textit{Lemur}, \textit{Saimiri} and \textit{Tarsius} to create the hominoid-cercopithecoid split, which is named as ``HomiCerco''. 
+Since we know the most recent common ancestor of humans and chimps may live on about $5-7$ million years ago, a normal distribution centered at 6 million years with a standard deviation of 0.5 million years is assigned to the ``Human-Chimp'' prior.
+Another normal distribution of $24 \pm 0.5$ (million years) is likewise defined for ``HomiCerco''.
 
 \begin{figure}[ht]
-\subfigure[Estimates]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/frequency.pdf}}
+\subfigure[Estimates]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/treeRootHeight.pdf}}
 \subfigure[Density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/multiKDE.pdf}}\\
-\subfigure[Joint-Marginal]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/joint-marginal.pdf}}
+\subfigure[Joint-Marginal]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/joint-marginal-th-ingroup.pdf}}
 \subfigure[Trace]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/trace.pdf}}
-\caption{Image exported from four analysis tabs in Tracer which is used in the ``Divergence Dating'' tutorial.}
+\caption{Images exported from four analysis tabs in Tracer for calibrating the tree to estimate TMRCA, where the black is ``treeModel.rootHeight'' parameter, blue is TMRCA parameter of ``Human-Chimp'', red is ``ingroup'', and yellow is ``HomiCerco''.}
 \label{fig:4tabs}
@@
 
-\subsection*{Virus Practical}
+Selecting the ``treeModel.rootHeight'' parameter gives the marginal posterior distribution of the age of the root of entire tree.
+In the example in Figure~\ref{fig:4tabs}.a, we can conclude the root of the tree is most likely between about 39.01 and 63.67 million years old with the mean of 50.48 million years.
+
+According to the kernel density estimation of the two TMRCA parameters (``Human-Chimp'' in blue and ``HomiCerco'' in yellow) in Figure~\ref{fig:4tabs}.b, these two parameters that we used to calibrate the tree explicitly have posterior distributions very similar to the prior distributions that we specified.
+
+The marginal posterior densities of ``treeModel.rootHeight'' and ``ingroup'' are nearly overlapped each other.
+If we switch to Figure~\ref{fig:4tabs}.c, their joint-marginal distribution plot is almost lined up together, which indicates ??.
+But this isn't an issue here, because the ``ingroup'' only excludes one outgroup species from everyone else.
+%TODO
+
+Figure~\ref{fig:4tabs}.d illustrates the traces of these four parameters against their logged states from MCMC chain.
+``Human-Chimp'' and ``HomiCerco'' parameters are immediately reaching equilibrium from the starting point, but ``treeModel.rootHeight'' and ``ingroup'' clearly show a burn-in period.
+%TODO ?
+
+\subsection*{Demographic reconstruction}
+
+The second analysis uses an 1698bp alignment of 21 sequences from the Influenza A virus (?). %data citation
+The piecewise-constant Bayesian skyline is selected for the tree prior, where the number of groups is reduced to 5 to avoid over-parameterize the model.
 

### 2017-07-27 manuscript/tracer16.tex
add example 2

@@
 %citation?
-Besides ``Bayesian Skyline Reconstruction'' reconstructs a generalised Bayesian skyline plot (e.g.
+Besides ``Bayesian Skyline Reconstruction'' reconstructs a generalised Bayesian coalescent skyline plot (e.g.
 Figure~\ref{fig:flu}.a) describing demographic history under canonical scenarios \citep{drummond2005bayesian},
@@
 %data citation
-%using HKY + $\Gamma$ model, uncorrelated lognormal relaxed molecular clock, and Yule prior on the tree.
+%using HKY + $\Gamma_4$ model, uncorrelated lognormal relaxed molecular clock, and Yule prior on the tree.
 In this analysis, we furthermore specify three calibrations on the phylogenetic tree to estimate time to the most recent common ancestor (TMRCA) based on our prior fossil knowledge.
@@
 The last consists of everything except \textit{Lemur}, \textit{Saimiri} and \textit{Tarsius} to create the hominoid-cercopithecoid split, which is named as ``HomiCerco''. 
-Since we know the most recent common ancestor of humans and chimps may live on about $5-7$ million years ago, a normal distribution centered at 6 million years with a standard deviation of 0.5 million years is assigned to the ``Human-Chimp'' prior.
+Since we know the most recent common ancestor of humans and chimps may live about $5-7$ million years ago, a normal distribution centered at 6 million years with a standard deviation of 0.5 million years is assigned to the ``Human-Chimp'' prior.
 Another normal distribution of $24 \pm 0.5$ (million years) is likewise defined for ``HomiCerco''.
@@
 
-Selecting the ``treeModel.rootHeight'' parameter gives the marginal posterior distribution of the age of the root of entire tree.
-In the example in Figure~\ref{fig:4tabs}.a, we can conclude the root of the tree is most likely between about 39.01 and 63.67 million years old with the mean of 50.48 million years.
+Selecting the ``treeModel.rootHeight'' parameter gives the marginal posterior distribution of the age of the root of entire tree in the example in Figure~\ref{fig:4tabs}.a.
+We can conclude the root of the tree is most likely between about 39.01 and 63.67 million years old expressed by the 95\% HPD interval with the mean of 50.48 million years.
 
-According to the kernel density estimation of the two TMRCA parameters (``Human-Chimp'' in blue and ``HomiCerco'' in yellow) in Figure~\ref{fig:4tabs}.b, these two parameters that we used to calibrate the tree explicitly have posterior distributions very similar to the prior distributions that we specified.
+Figure~\ref{fig:4tabs}.b shows the kernel density estimation of the two TMRCA parameters that ``Human-Chimp'' is blue and ``HomiCerco'' is yellow.
+These two parameters that we used to calibrate the tree explicitly have posterior distributions very similar to the prior distributions that we specified.
 
-The marginal posterior densities of ``treeModel.rootHeight'' and ``ingroup'' are nearly overlapped each other.
-If we switch to Figure~\ref{fig:4tabs}.c, their joint-marginal distribution plot is almost lined up together, which indicates ??.
+The marginal posterior densities of ``treeModel.rootHeight'' and ``ingroup'' are nearly overlapped each other, when we switch to Figure~\ref{fig:4tabs}.c.
+Their joint-marginal distribution plot is almost lined up together, which indicates ??.
 But this isn't an issue here, because the ``ingroup'' only excludes one outgroup species from everyone else.
-%TODO
+%TODO how to explain this problem further?
 
@@
 
-\subsection*{Demographic reconstruction}
+\subsection*{Demographic reconstruction and temporal analysis}
 
-The second analysis uses an 1698bp alignment of 21 sequences from the Influenza A virus (?). %data citation
-The piecewise-constant Bayesian skyline is selected for the tree prior, where the number of groups is reduced to 5 to avoid over-parameterize the model.
+In the second example, there are two analyses conducted using an 1698bp alignment of 21 sequences from the Influenza A virus (H5N1?) collected between 1997 and 2005 (?). %data citation
+To perform Bayesian coalescent skyline plot in Figure~\ref{fig:flu}.a, we exploit Tracer to exact past population dynamics from both MCMC log and tree log produced by a BEAST \citep{drummond2007beast} run.
+%using piecewise-constant Bayesian skyline tree prior with 5 groups in order to avoid over-parameterize the model.
+The plot reveals the growth of the effective population sizes of Influenza infections during 8 years, which obviously involves an epidemic break-out period between 2002 and 2004.
 
@@
 %\includegraphics[width=.5\textwidth]{./figures/fluBSP.pdf}  
-\subfigure[Bayesian Skyline]{\includegraphics[width = .23\textwidth]{./figures/fluBSP.pdf}}
+\subfigure[Bayesian Coalescent Skyline]{\includegraphics[width = .23\textwidth]{./figures/fluBSP.pdf}}
 \subfigure[Lineages Through Time]{\includegraphics[width = .23\textwidth]{./figures/fluLTT.pdf}}
-\caption{The plot (a) performs Bayesian Skyline reconstruction to estimate population sizes over time and (b) Lineages Through Time , from ``Flu'' data, where time starts from year 1997 at the 0 position and the unit is year.}
+\caption{(a) Bayesian Coalescent Skyline plot and (b) Lineages Through Time plot for the Influenza A virus data.}
 \label{fig:flu}
 \end{figure}
+
+Lineages Through Time in Figure~\ref{fig:flu}.b, similar curve
+
+diversification has been gradually developed before 2003's break-out, but it also looks stopped after 2004.
 

### 2017-07-28 manuscript/tracer16.tex

--- 2017-07-25 manuscript/tracer16.tex
@@
 It can be used to analyse runs of BEAST \citep{drummond2007beast,drummond2012bayesian}, BEAST2 \citep{bouckaert2014beast2}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes}, LAMARC \citep{kuhner2006lamarc}, Migrate \citep{beerli2006comparison},  and possibly other MCMC programs.
-In addition, Tracer will be able to redraw the distribution of effective population sizes over time if a demographic model was involved in the analysis.
+In addition, Tracer's demographic reconstruction functions are utilised to redraw the distribution of effective population sizes over time if an available demographic model was involved in the analysis.
 
@@
 The list of traces in each log are created from the sampling parameters logged during a MCMC process, which include their statistical summaries from their values at all states logged. 
-The effective sample size (ESS) is one of important measurements of a sampled continuous parameter, which tells the number of effectively independent draws from the posterior distribution that the Markov chain is equivalent to.
+The effective sample size (ESS) is an important measurement designed for a sampled parameter having continuous values, which tells the number of effectively independent draws from the posterior distribution that the Markov chain is equivalent to.
 
@@
 The integer is dealing with quantitative discrete variables, and the categorical trace is for qualitative categorical variables. 
-% check about definition 
-These three types can be automatically assigned when the log file is imported. But they can also be changed later.
+% check above definition 
+%These three types can be automatically assigned when the log file is imported. But they can also be changed later.
 
@@
 The multiple trace files can be similarly selected as well to compare different runs.
-If multiple trace files have the same trace names then a ``Combined'' trace will automatically appear, which can be selected as well as the individual trace files.
+If multiple trace files have the same trace names,  then a ``Combined'' trace will automatically appear, which can be selected as well as the individual trace files.
 
@@
 
-\item Density - this shows the Bayesian posterior density plot for the selected trace(s), if they are continuous, the kernel density estimates will be available, for example, in Figure~\ref{fig:4tabs}.b.
+\item Density - this shows the Bayesian posterior density plot for the selected trace(s), if they are continuous, the kernel density estimates will be available for continuous traces, for example, in Figure~\ref{fig:4tabs}.b.
 
@@
 
-Tracer also provides demographic reconstruction resulting in a graphical plot to estimate the distribution of effective population sizes over time, given a MCMC log from a Bayesian framework or additionally with the tree log depending on the model.
-For example, ``Demographic Reconstruction" performs reconstructions of the distribution for a number of models, such as constant size \citep{drummond2002estimating}, exponential growth \citep{drummond2002estimating}, and logistic growth (?).
+Tracer also provides demographic reconstruction resulting in a graphical plot, often applyed to reconstruct epidemic dynamics, given a MCMC log from a Bayesian framework or additionally with the tree log depending on the model.
+The currently available models are constant size \citep{drummond2002estimating}, exponential growth \citep{drummond2002estimating}, logistic growth (?), Bayesian coalescent skyline \citep{drummond2005bayesian}, GMRF Bayesian skyride \citep{minin2008smooth}, and Bayesian coalescent skygrid \citep{gill2012improving}.
+%For example, ``Demographic Reconstruction" performs reconstructions of the distribution for a number of models, such as constant size \citep{drummond2002estimating}, exponential growth \citep{drummond2002estimating}, and logistic growth (?).
 %citation?
-Besides ``Bayesian Skyline Reconstruction'' reconstructs a generalised Bayesian skyline plot (e.g.
+%Besides ``Bayesian Skyline Reconstruction'' reconstructs a generalised Bayesian coalescent skyline plot (e.g.
 Figure~\ref{fig:flu}.a) describing demographic history under canonical scenarios \citep{drummond2005bayesian},
-``GMRF Skyride Reconstruction'' and ``SkyGrid Reconstruction'' are respectively available for a temporal smoothing method using the Gaussian Markov random field (GMRF) model  \citep{minin2008smooth} and a generalisation of the GMRF model \citep{gill2012improving}.
-``Lineages Through Time'' (?) plots quantiles of the number of lineages against time, which describes the change of diversification of all organisms over time. %citation?
-All these models are available in BEAST \citep{drummond2007beast}.
+%``GMRF Skyride Reconstruction'' and ``SkyGrid Reconstruction'' are respectively available for a temporal smoothing method using the Gaussian Markov random field (GMRF) model  \citep{minin2008smooth} and a generalisation of the GMRF model \citep{gill2012improving}.
+Besides these, ``Lineages Through Time'' (?) is also available to plot the quantiles of the number of lineages against time, which describes the change of diversification of all organisms over time.
+%citation?
+%All these models are available in BEAST \citep{drummond2007beast}.
 
-%TODO better to re-write
+%TODO is it OK ?
 Tracer recently offers a fairly general solution of looking at conditional posterior distributions.
-This, specifically, supports for BSSVS forms of model averaging, in which some parameters are only in the likelihood when their sub-model is ``indicated" by some indicator function that is usually a discrete, integer or boolean variable.
-In that case the posterior of the parameter should not include the states when it was sampled only in the prior because the sub-model it belonged to was ``turned off".
-This is relevant for EBSP, Random Local Clocks model, microsatellite model averaging and relaxed clock model averaging methods all from my group in last couple of years.
+This, specifically, supports for Bayesian stochastic search variable selection (BSSVS) forms of model averaging, in which some parameters are only in the likelihood when their sub-model is indicated by an indicator function usually assigned to a discrete, integer or boolean variable.
+In that case, the posterior of the parameter should not include the states when it was sampled only in the prior, because the sub-model it belonged to was ``turned off".
+This is relevant for extended Bayesian skyline plot, random local clock model, microsatellite model averaging, and relaxed clock model averaging methods, etc.
 It is also relevant for BSSVS in phylogeography as well depending on how the state is logged.
@@
 %data citation
-%using HKY + $\Gamma$ model, uncorrelated lognormal relaxed molecular clock, and Yule prior on the tree.
-In this analysis, we furthermore specify three calibrations on the phylogenetic tree to estimate time to the most recent common ancestor (TMRCA) based on our prior fossil knowledge.
+
+%using HKY + $\Gamma_4$ model, uncorrelated lognormal relaxed molecular clock, and Yule prior on the tree.
+In this analysis, we furthermore specified three calibrations on the phylogenetic tree to estimate time to the most recent common ancestor (TMRCA) based on our prior fossil knowledge.
 
@@
 The last consists of everything except \textit{Lemur}, \textit{Saimiri} and \textit{Tarsius} to create the hominoid-cercopithecoid split, which is named as ``HomiCerco''. 
-Since we know the most recent common ancestor of humans and chimps may live on about $5-7$ million years ago, a normal distribution centered at 6 million years with a standard deviation of 0.5 million years is assigned to the ``Human-Chimp'' prior.
+We know the most recent common ancestor of humans and chimps may live about $5-7$ million years ago, a normal distribution centered at 6 million years with a standard deviation of 0.5 million years is assigned to the ``Human-Chimp'' prior.
 Another normal distribution of $24 \pm 0.5$ (million years) is likewise defined for ``HomiCerco''.
@@
 
-Selecting the ``treeModel.rootHeight'' parameter gives the marginal posterior distribution of the age of the root of entire tree.
-In the example in Figure~\ref{fig:4tabs}.a, we can conclude the root of the tree is most likely between about 39.01 and 63.67 million years old with the mean of 50.48 million years.
+After loading the log, selecting the ``treeModel.rootHeight'' parameter gives the marginal posterior distribution of the age of the root of entire tree.
+The blue bars in Figure~\ref{fig:4tabs}.a present the most compact interval on the root height of the tree that contains 95\% of the posterior probability, referring to the period between about 39.01 and 63.67 million years old with the mean of 50.48 million years.
+%TODO rate is faster than Yang's book r_ape = 5.4 * 10^-10 per year, r_monkey = 8.0, r_root = 6.6 
+Correspondingly selecting ``meanRate'' provides the rate of evolution averaged over the whole tree, which has the mean of 0.0129 substitutions per site per million years with 95\% HPD Interval between 0.0101 and 0.0159.
 
-According to the kernel density estimation of the two TMRCA parameters (``Human-Chimp'' in blue and ``HomiCerco'' in yellow) in Figure~\ref{fig:4tabs}.b, these two parameters that we used to calibrate the tree explicitly have posterior distributions very similar to the prior distributions that we specified.
+Figure~\ref{fig:4tabs}.b shows the kernel density estimations of these four parameters.
+The two TMRCA parameters, ``Human-Chimp'' in blue and ``HomiCerco'' in yellow, that we used to calibrate the tree explicitly have posterior distributions very similar to the prior distributions that we specified.
 
 The marginal posterior densities of ``treeModel.rootHeight'' and ``ingroup'' are nearly overlapped each other.
-If we switch to Figure~\ref{fig:4tabs}.c, their joint-marginal distribution plot is almost lined up together, which indicates ??.
-But this isn't an issue here, because the ``ingroup'' only excludes one outgroup species from everyone else.
-%TODO
+When we switch to their joint-marginal distribution plot in Figure~\ref{fig:4tabs}.c, the samples are almost lined up together, which indicates they are associated.
+But this isn't an issue here, because the ``ingroup'' only excludes one outgroup species from all taxa. 
+%TODO how to explain this problem further?
 
@@
 
-\subsection*{Demographic reconstruction}
+\subsection*{Non-parametric coalescent estimator}
 
-The second analysis uses an 1698bp alignment of 21 sequences from the Influenza A virus (?). %data citation
-The piecewise-constant Bayesian skyline is selected for the tree prior, where the number of groups is reduced to 5 to avoid over-parameterize the model.
+In the second example, there are two analyses conducted using a H3N2 data set that is the subset of a comprehensive influenza A virus data set spanning several epidemic seasons in the New York state.
+The data contains a  ?bp alignment of 165 Hemagglutinin genes collected between about 1999 and 2004 \citep{rambaut2008genomic}, explicating H3N2 spreading during three epidemic seasons.
+
+To be able to reconstruct Bayesian coalescent skygrid \citep{minin2008smooth} plot, we exploit Tracer to exact past population dynamics from both MCMC log and tree log produced by a BEAST \citep{drummond2007beast} run.
+The HKY + $\Gamma_4$ model, strict molecular clock, and Bayesian skygrid tree prior was respectively setup in the run, where a grid of 50 intervals over 5 years estimating 10 population sizes by year was constructed in the skygrid prior.
+
+%What type of dynamics does the H3N2 skyride plot suggest?
+According to Tracer's skygrid plot in Figure~\ref{fig:flu}.a, the changing patterns of genetic diversity in viral isolates from New York state clearly reveal the seasonal dynamics of influenza A in individual temperate populations.
+The absolute amount of genetic diversity, even at seasonal peaks, is small compared to other rapidly evolving viruses that infect far fewer people, suggesting that strong natural selection, in addition to periodic bottlenecks, reduces the level of diversity that co-circulates at any time.
+Simulations demonstrate that this reconstruction of genetic diversity is robust to the sampling protocol.
+%TODO copied from (Rambaut et al., 2008)
 
@@
 %\includegraphics[width=.5\textwidth]{./figures/fluBSP.pdf}  
-\subfigure[Bayesian Skyline]{\includegraphics[width = .23\textwidth]{./figures/fluBSP.pdf}}
-\subfigure[Lineages Through Time]{\includegraphics[width = .23\textwidth]{./figures/fluLTT.pdf}}
-\caption{The plot (a) performs Bayesian Skyline reconstruction to estimate population sizes over time and (b) Lineages Through Time , from ``Flu'' data, where time starts from year 1997 at the 0 position and the unit is year.}
+\subfigure[Bayesian coalescent skygrid]{\includegraphics[width = .23\textwidth]{./figures/fig21.png}}
+\subfigure[Lineages through time]{\includegraphics[width = .23\textwidth]{./figures/fluLTT.pdf}}
+\caption{(a) Bayesian coalescent skygrid plot and (b) lineages through time for the Influenza A/H3N2 virus data.}
 \label{fig:flu}
 \end{figure}
+
+Lineages Through Time in Figure~\ref{fig:flu}.b, similar curve
+
+diversification has been gradually developed before 2003's break-out, but it also looks stopped after 2004.
 
@@
 
-\begin{figure}[H]
-\includegraphics[width=.5\textwidth]{./figures/comp-95HPD.pdf}  
-\caption{The comparison of  95\% HPD intervals from multi-trace}
-\label{fig:comp95HPD}
-\end{figure}
+%\begin{figure}[H]
+%\includegraphics[width=.5\textwidth]{./figures/comp-95HPD.pdf}  
+%\caption{The comparison of  95\% HPD intervals from multi-trace}
+%\label{fig:comp95HPD}
+%\end{figure}
 
-\begin{figure}[H]
-\includegraphics[width=.5\textwidth]{./figures/jointPrInt.png}  
-\caption{The Joint probability table of two integer traces}
-\label{fig:int:jointpr}
-\end{figure}
+%\begin{figure}[H]
+%\includegraphics[width=.5\textwidth]{./figures/jointPrInt.png}  
+%\caption{The Joint probability table of two integer traces}
+%\label{fig:int:jointpr}
+%\end{figure}
 

### 2017-07-28 src/tracer/analysis/DemographicPlotPanel.java
Making the default colours a bit more attractive.

@@ -32,6 +32,8 @@ import javax.swing.*;
 import java.awt.*;
 import java.awt.event.ActionEvent;
 
+import static tracer.analysis.TemporalAnalysisPlotPanel.AREA_COLOUR;
+
 
 /**
  * A panel that displays demographic plot
@@ -167,15 +169,15 @@ public class DemographicPlotPanel extends JPanel {
         if (solidIntervalCheckBox.isSelected()) {
 
             AreaPlot areaPlot = new AreaPlot(xData, yDataUpper, xData, yDataLower);
-            areaPlot.setLineColor(new Color(0x9999FF));
+            areaPlot.setLineColor(AREA_COLOUR);
             demoChart.addPlot(areaPlot);
         } else {
             LinePlot plot = new LinePlot(xData, yDataLower);
-            plot.setLineStyle(new BasicStroke(1.0F), new Color(0x9999FF));
+            plot.setLineStyle(new BasicStroke(1.0F), AREA_COLOUR);
             demoChart.addPlot(plot);
 
             plot = new LinePlot(xData, yDataUpper);
-            plot.setLineStyle(new BasicStroke(1.0F), new Color(0x9999FF));
+            plot.setLineStyle(new BasicStroke(1.0F), AREA_COLOUR);
             demoChart.addPlot(plot);
 
         }

### 2017-07-28 src/tracer/analysis/TemporalAnalysisPlotPanel.java

@@ -41,7 +41,7 @@ import java.util.ArrayList;
  * @version $Id: TemporalAnalysisPlotPanel.java,v 1.2 2006/12/01 09:36:17 rambaut Exp $
  */
 public class TemporalAnalysisPlotPanel extends JPanel {
-
+    public static final Color AREA_COLOUR = new Color(0x1f78b4);
 
     private JChart demoChart = new JChart(new LinearAxis(Axis.AT_DATA, Axis.AT_DATA), new LogAxis());
     private JChartPanel chartPanel = new JChartPanel(demoChart, null, "", "");
@@ -173,15 +173,15 @@ public class TemporalAnalysisPlotPanel extends JPanel {
             if (solidIntervalCheckBox.isSelected()) {
 
                 AreaPlot areaPlot = new AreaPlot(analysis.xData, analysis.yDataUpper, analysis.xData, analysis.yDataLower);
-                areaPlot.setLineColor(new Color(0x9999FF));
+                areaPlot.setLineColor(AREA_COLOUR);
                 demoChart.addPlot(areaPlot);
             } else {
                 LinePlot plot = new LinePlot(analysis.xData, analysis.yDataLower);
-                plot.setLineStyle(new BasicStroke(1.0F), new Color(0x9999FF));
+                plot.setLineStyle(new BasicStroke(1.0F), AREA_COLOUR);
                 demoChart.addPlot(plot);
 
                 plot = new LinePlot(analysis.xData, analysis.yDataUpper);
-                plot.setLineStyle(new BasicStroke(1.0F), new Color(0x9999FF));
+                plot.setLineStyle(new BasicStroke(1.0F), AREA_COLOUR);
                 demoChart.addPlot(plot);
             }
         }

### 2017-07-28 src/tracer/traces/DensityPanel.java

@@ -368,9 +368,9 @@ public class DensityPanel extends NTracesChartPanel {
                             Plot plot2 = setupKDEPlot(values, td);
                             plot2.setName(name + " KDE");
                             if (tl instanceof CombinedTraces) {
-                                plot2.setLineStyle(new BasicStroke(2.0f), paints[i]);
+                                plot2.setLineStyle(new BasicStroke(2.0f), PAINTS[i]);
                             } else {
-                                plot2.setLineStyle(new BasicStroke(1.0f), paints[i]);
+                                plot2.setLineStyle(new BasicStroke(1.0f), PAINTS[i]);
                             }
                             traceChart.addPlot(plot2);
                         }
@@ -392,9 +392,9 @@ public class DensityPanel extends NTracesChartPanel {
                     if (plot != null) {
                         plot.setName(name);
                         if (tl instanceof CombinedTraces) {
-                            plot.setLineStyle(new BasicStroke(2.0f), paints[i]);
+                            plot.setLineStyle(new BasicStroke(2.0f), PAINTS[i]);
                         } else {
-                            plot.setLineStyle(new BasicStroke(1.0f), paints[i]);
+                            plot.setLineStyle(new BasicStroke(1.0f), PAINTS[i]);
                         }
 
                         traceChart.addPlot(plot);
@@ -406,7 +406,7 @@ public class DensityPanel extends NTracesChartPanel {
                     if (currentSettings.colourBy == ColourByOptions.COLOUR_BY_TRACE || currentSettings.colourBy == ColourByOptions.COLOUR_BY_ALL) {
                         i++;
                     }
-                    if (i == paints.length) i = 0;
+                    if (i == PAINTS.length) i = 0;
                 }
             }
             if (currentSettings.colourBy == ColourByOptions.COLOUR_BY_FILE) {
@@ -414,7 +414,7 @@ public class DensityPanel extends NTracesChartPanel {
             } else if (currentSettings.colourBy == ColourByOptions.COLOUR_BY_TRACE) {
                 i = 0;
             }
-            if (i == paints.length) i = 0;
+            if (i == PAINTS.length) i = 0;
         }
 
         setXLabMultiTraces();

### 2017-07-28 src/tracer/traces/RawTracePanel.java

@@ -263,7 +263,7 @@ public class RawTracePanel extends NTracesChartPanel {
                                 getTraceChart().getYAxis().setRange(0.0, 1.0);
                             }
                         }
-                        minMax = getTraceChart().addTrace(name, stateStart, stateStep, values, burninValues, paints[i]);
+                        minMax = getTraceChart().addTrace(name, stateStart, stateStep, values, burninValues, PAINTS[i]);
 
                     } else if (trace.getTraceType() == TraceType.CATEGORICAL) {
 
@@ -286,7 +286,7 @@ public class RawTracePanel extends NTracesChartPanel {
                         }
 
                         getTraceChart().setYAxis(false, categoryDataMap);
-                        minMax = getTraceChart().addTrace(name, stateStart, stateStep, doubleData, doubleBurninData, paints[i]);
+                        minMax = getTraceChart().addTrace(name, stateStart, stateStep, doubleData, doubleBurninData, PAINTS[i]);
 
                     } else {
                         throw new RuntimeException("Trace type is not recognized: " + trace.getTraceType());
@@ -299,7 +299,7 @@ public class RawTracePanel extends NTracesChartPanel {
                     if (currentSettings.colourBy == ColourByOptions.COLOUR_BY_TRACE || currentSettings.colourBy == ColourByOptions.COLOUR_BY_ALL) {
                         i++;
                     }
-                    if (i == paints.length) i = 0;
+                    if (i == PAINTS.length) i = 0;
                 }
             }
             if (currentSettings.colourBy == ColourByOptions.COLOUR_BY_FILE) {
@@ -307,7 +307,7 @@ public class RawTracePanel extends NTracesChartPanel {
             } else if (currentSettings.colourBy == ColourByOptions.COLOUR_BY_TRACE) {
                 i = 0;
             }
-            if (i == paints.length) i = 0;
+            if (i == PAINTS.length) i = 0;
         }// for (TraceList tl : traceLists)
         if (traceLists.length > 1 || traceNames.size() > 1) {
             Variate.D xV = new Variate.D(valuesX);

### 2017-07-28 src/tracer/traces/TraceChartPanel.java

@@ -57,7 +57,7 @@ public abstract class TraceChartPanel extends JPanel implements Exportable {
         COLOUR_BY_ALL
     };
 
-    protected static final Paint[] paints = new Paint[]{
+    protected static final Paint[] LEGACY_PAINTS = new Paint[]{
             Color.BLACK,
             new Color(64, 35, 225),
             new Color(229, 35, 60),
@@ -69,8 +69,59 @@ public abstract class TraceChartPanel extends JPanel implements Exportable {
             new Color(239, 255, 34),
             Color.DARK_GRAY
     };
+    
+    protected static final Paint[] COLORBREWER_PAIRED = new Paint[]{
+            new Color(0xa6cee3),
+            new Color(0x1f78b4),
+            new Color(0xb2df8a),
+            new Color(0x33a02c),
+            new Color(0xfb9a99),
+            new Color(0xe31a1c),
+            new Color(0xfdbf6f),
+            new Color(0xff7f00),
+            new Color(0xcab2d6),
+            new Color(0x6a3d9a),
+            new Color(0xffff99),
+            new Color(0xb15928)
+    };
 
-    protected class Settings {
+    protected static final Paint[] COLORBREWER_DARK2 = new Paint[]{
+            new Color(0x1b9e77),
+            new Color(0xd95f02),
+            new Color(0x7570b3),
+            new Color(0xe7298a),
+            new Color(0x66a61e),
+            new Color(0xe6ab02),
+            new Color(0xa6761d),
+            new Color(0x666666)
+    };
+
+    protected static final Paint[] RAINBOW = new Paint[]{
+            new Color(0x46a4d5),
+            new Color(0x3cbeb7),
+            new Color(0xb0be36),
+            new Color(0xfdcb42),
+            new Color(0xf7a139),
+            new Color(0xf4622e),
+            new Color(0xd93a3d),
+            new Color(0xec667c),
+            new Color(0xa267d6)
+    };
+
+    protected static final Paint[] RAINBOW_2 = new Paint[]{
+            new Color(0x394958),
+            new Color(0x5b9586),
+            new Color(0x6f955b),
+            new Color(0xdfbd5d),
+            new Color(0xd17e53),
+            new Color(0xc25553),
+            new Color(0x955b6e)
+    };
+
+    // paint pallet being used
+    protected static final Paint[] PAINTS = COLORBREWER_DARK2;
+
+   protected class Settings {
         // shared settings
         ChartSetupDialog chartSetupDialog = null;
         int legendAlignment = 0;

### 2017-07-28 src/tracer/traces/FrequencyPanel.java
Making the default colours a bit more attractive.

@@ -32,6 +32,7 @@ import dr.inference.trace.TraceList;
 import dr.inference.trace.TraceType;
 
 import javax.swing.*;
+import java.awt.*;
 import java.awt.event.ActionEvent;
 import java.util.List;
 
@@ -44,6 +45,9 @@ import java.util.List;
  * @version $Id: FrequencyPanel.java,v 1.1.1.2 2006/04/25 23:00:09 rambaut Exp $
  */
 public class FrequencyPanel extends OneTraceChartPanel {
+
+    public final static Paint BAR_PAINT = new Color(0x2f8aa3);
+    public final static  Paint QUANTILE_PAINT = new Color(0xd6bd58);
 
     private Settings currentSettings = new Settings();
 //    private Map<String, Settings> settingsMap = new HashMap<String, Settings>();
@@ -132,6 +136,8 @@ public class FrequencyPanel extends OneTraceChartPanel {
             TraceType traceType = trace.getTraceType();
             if (traceType.isContinuous()) {
                 plot = new FrequencyPlot(values, currentSettings.minimumBins, td);
+
+                plot.setPaints(BAR_PAINT, QUANTILE_PAINT);
 
                 if (td != null) {
                     plot.setIntervals(td.getUpperHPD(), td.getLowerHPD());

### 2017-08-01 manuscript/tracer16.tex
abstract

@@
 %Computational evolutionary biology, statistical phylogenetics and coalescent-based population genetics are becoming increasingly central to the analysis and understanding of molecular sequence data. 
-Bayesian inference of phylogeny using Markov chain Monte Carlo (MCMC) plays a central role for understanding the story of evolutions from molecular sequence data.  
-The comprehension of posterior distributions from an implementation of this framework is the significant matter.
+Bayesian inference of phylogeny using Markov chain Monte Carlo (MCMC) plays a central role in understanding evolutionary history and processes from molecular sequence data.  
+Visualising and analysing the samples from the posterior distribution is a key step in any non-trivial Bayesian inference, and it is especially important for methods that employ MCMC.
 \section{Results:}
-%We describe here a software package, Tracer version 1.6, which ... 
-We here present a software package \emph{Tracer} (version 1.6) for visualising and analysing the trace files generated by Bayesian MCMC runs. 
-It also provides kernel density estimation, cheap marginal likelihood estimator, 
-multivariate analyses (is this reluctant?), %TODO
-demographic trajectory reconstruction, conditional posterior distribution summary, etc. 
+Here we present a software package \emph{Tracer} (version 1.6) for visualising and analysing the trace files generated by Bayesian MCMC runs. 
+The \emph{Tracer} software also provides kernel density estimation, cheap marginal likelihood estimation, 
+multivariate visualisation, demographic trajectory reconstruction, conditional posterior distribution summary and more. 
 \section{Availability:}

### 2017-08-01 manuscript/tracer16.tex
intro

@@
 
-Bayesian inference of phylogeny \citep{huelsenbeck2001bayesian, felsenstein2004inferring,yang2006computational} has been a popular approach to reconstruct the evolutionary history and discover the evolutionary relationships among organisms that could come from genes, genomes, individuals or species.
-The statistical framework is based on the posterior probability of phylogenetic trees given the data, and currently takes the advantage of Markov chain Monte Carlo (MCMC) to promote the computational speed of Bayesian estimation to reach an approximated optimisation.
-The posterior summaries, as well as as scientifically relevant marginalisation, %TODO
-are therefore critically important for understanding a Bayesian phylogenetic analysis.
+Bayesian inference of phylogeny \citep{huelsenbeck2001bayesian, yang2006computational, DrummondBouckaert2015} has been a popular approach to reconstruct the evolutionary history and discover the evolutionary relationships among taxa, whether genes, genomes, individuals or species.
+Bayesian inference can be used to produce the posterior probability distribution over the space of phylogenetic trees given the data.
+Most major software packages currently used employ Markov chain Monte Carlo (MCMC) to produce a sample from the posterior distribution.
+Visualising, summarizing and marginalising these posterior samples is necessary for evaluating and reporting a Bayesian phylogenetic analysis.
 %However, the full posterior distribution is high-dimensional and often difficult to visualise. 

### 2017-08-03 manuscript/tracer16.tex
update acknowledgments

@@
 
-We thank the National Evolutionary Synthesis Center (NESCent) for sponsoring a working group (Software for Bayesian Evolutionary Analysis) that facilitated the development of Tracer version 1.6. 
-This work was supported in part by funding from the Marsden Trust, NSF (DMS 0856099), NIH (R01 GM086887, R01 HG006139), The Royal Society of London, BBSRC (BB/H011285/1) and the Wellcome Trust (WT092807MA).
+This work was supported in part by
+the Wellcome Trust through project 206298/Z/17/Z (Artic Network),
+the European Union Seventh Framework Programme under grant agreement no. 725422-RESERVOIRDOCS,
+the National Science Foundation through grant DMS 1264153,
+the National Institutes of Health under grant R01 AI107034,
+and
+the Marsden Trust.
 

### 2017-08-03 manuscript/tracer16.tex
use most recent references, unless describing historical developments

@@
 We then develop a software package Tracer (version 1.6) for this purpose, which analyses Bayesian MCMC log files containing the list of parameters whose values sampled from the chain.
-It can be used to analyse runs of BEAST \citep{drummond2007beast,drummond2012bayesian}, BEAST2 \citep{bouckaert2014beast2}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes}, LAMARC \citep{kuhner2006lamarc}, Migrate \citep{beerli2006comparison},  and possibly other MCMC programs.
+It can be used to analyse runs of BEAST \citep{drummond2012bayesian}, BEAST2 \citep{bouckaert2014beast2}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes}, LAMARC \citep{kuhner2006lamarc}, Migrate \citep{beerli2006comparison},  and possibly other MCMC programs.
 In addition, Tracer's demographic reconstruction functions are utilised to redraw the distribution of effective population sizes over time if an available demographic model was involved in the analysis.
@@
 \begin{figure}[ht]
-\includegraphics[width=.38\textwidth]{./figures/tracer.pdf}
+\centerline{
+\includegraphics[width=0.45\textwidth]{./figures/tracer.pdf}
+}
 \caption{The system design of Tracer 1.6}
@@
 
-Figure~\ref{fig:4tabs} demonstrates a snapshot of posterior summaries in Tracer for a BEAST \citep{drummond2007beast} run created from a 898 bp alignment of sequences of 12 species of primates (http://beast.bio.ed.ac.uk/tutorials?).
+Figure~\ref{fig:4tabs} demonstrates a snapshot of posterior summaries in Tracer for a BEAST \citep{drummond2012bayesian} run created from a 898 bp alignment of sequences of 12 species of primates (http://beast.bio.ed.ac.uk/tutorials?).
 %data citation
@@
 
-To be able to reconstruct Bayesian coalescent skygrid \citep{minin2008smooth} plot, we exploit Tracer to exact past population dynamics from both MCMC log and tree log produced by a BEAST \citep{drummond2007beast} run.
+To be able to reconstruct Bayesian coalescent skygrid \citep{minin2008smooth} plot, we exploit Tracer to exact past population dynamics from both MCMC log and tree log produced by a BEAST \citep{drummond2012bayesian} run.
 The HKY + $\Gamma_4$ model, strict molecular clock, and Bayesian skygrid tree prior was respectively setup in the run, where a grid of 50 intervals over 5 years estimating 10 population sizes by year was constructed in the skygrid prior.
@@
 
-\section*{Old Text}
-
-%You can also select the "Demographic Analysis" from the Analysis menu - This plots the distribution of demographic population sizes over time for a number of models (constant size, exponential growth \& logistic growth) that are available in BEAST.
+%\section*{Old Text}
+%
+%%You can also select the "Demographic Analysis" from the Analysis menu - This plots the distribution of demographic population sizes over time for a number of models (constant size, exponential growth \& logistic growth) that are available in BEAST.
 This involves you selecting the traces for each parameter of the model.
 You should only select the model that was actually run under BEAST (e.g., if you ran an exponential growth model, you shouldn't plot the constant population size model).
-
-%The "Analysis" menu also contains options for performing Bayesian Skyline reconstructions and for calculating Bayes Factors between runs.
-
-Feature list:
-\begin{itemize}
-\item KDEs
-\item cheap marginal likelihood estimators
-\item demographic trajectory reconstruction
-\end{itemize}
-
-%Model selection HME and AICM have been removed
-New features:
-\begin{itemize}
-\item Three trace types: real, integer, string
-\item Kernel density estimates
-\item Conditional posterior distribution
-%\item HME
-%\item AICM
-\end{itemize}
-
-%\begin{figure}[H]
-%\includegraphics[width=.5\textwidth]{./figures/comp-95HPD.pdf}
-%\caption{The comparison of  95\% HPD intervals from multi-trace}
-%\label{fig:comp95HPD}
-%\end{figure}
-
-%\begin{figure}[H]
-%\includegraphics[width=.5\textwidth]{./figures/jointPrInt.png}
-%\caption{The Joint probability table of two integer traces}
-%\label{fig:int:jointpr}
-%\end{figure}
-
-%\begin{figure}[ht]
-%\includegraphics[width=.5\textwidth]{./figures/frequency.pdf}
-%\caption{A frequency distribution for a continuous trace}
-%\label{fig:freq}
-%\end{figure}
-%\begin{figure}[ht]
-%\includegraphics[width=.5\textwidth]{./figures/multiKDE.pdf}
-%\caption{KDE of multiple traces}
-%\label{fig:multiKDE}
-%\end{figure}
-%\begin{figure}[ht]
-%\includegraphics[width=.5\textwidth]{./figures/joint-marginal.pdf}
-%\caption{The joint-marginal distribution of two selected traces}
-%\label{fig:trace}
-%\end{figure}
-%\begin{figure}[ht]
-%\includegraphics[width=.5\textwidth]{./figures/trace.pdf}
-%\caption{The selected traces against state}
-%\label{fig:trace}
-%\end{figure}
-
-\subsubsection*{Diagnostics:} Blah blah
-
-\subsubsection*{Demographic reconstruction:}
-
-%\subsubsection*{Model selection:}
-
-\subsubsection*{Conditional posterior distribution:}
+%
+%%The "Analysis" menu also contains options for performing Bayesian Skyline reconstructions and for calculating Bayes Factors between runs.
+%
+%Feature list:
+%\begin{itemize}
+%\item KDEs
+%\item cheap marginal likelihood estimators
+%\item demographic trajectory reconstruction
+%\end{itemize}
+%
+%%Model selection HME and AICM have been removed
+%New features:
+%\begin{itemize}
+%\item Three trace types: real, integer, string
+%\item Kernel density estimates
+%\item Conditional posterior distribution
+%%\item HME
+%%\item AICM
+%\end{itemize}
+%
+%%\begin{figure}[H]
+%%\includegraphics[width=.5\textwidth]{./figures/comp-95HPD.pdf}
+%%\caption{The comparison of  95\% HPD intervals from multi-trace}
+%%\label{fig:comp95HPD}
+%%\end{figure}
+%
+%%\begin{figure}[H]
+%%\includegraphics[width=.5\textwidth]{./figures/jointPrInt.png}
+%%\caption{The Joint probability table of two integer traces}
+%%\label{fig:int:jointpr}
+%%\end{figure}
+%
+%%\begin{figure}[ht]
+%%\includegraphics[width=.5\textwidth]{./figures/frequency.pdf}
+%%\caption{A frequency distribution for a continuous trace}
+%%\label{fig:freq}
+%%\end{figure}
+%%\begin{figure}[ht]
+%%\includegraphics[width=.5\textwidth]{./figures/multiKDE.pdf}
+%%\caption{KDE of multiple traces}
+%%\label{fig:multiKDE}
+%%\end{figure}
+%%\begin{figure}[ht]
+%%\includegraphics[width=.5\textwidth]{./figures/joint-marginal.pdf}
+%%\caption{The joint-marginal distribution of two selected traces}
+%%\label{fig:trace}
+%%\end{figure}
+%%\begin{figure}[ht]
+%%\includegraphics[width=.5\textwidth]{./figures/trace.pdf}
+%%\caption{The selected traces against state}
+%%\label{fig:trace}
+%%\end{figure}
+%
+%
+%\subsubsection*{Diagnostics:} Blah blah
+%
+%\subsubsection*{Demographic reconstruction:}
+%
+%%\subsubsection*{Model selection:}
+%
+%\subsubsection*{Conditional posterior distribution:}
 
@@
 %Tracer is a program for analysing the trace files generated by Bayesian MCMC runs (that is, the continuous parameter values sampled from the chain).
-It can be used to analyse runs of BEAST \citep{drummond2007beast,drummond2012bayesian}, BEAST2 \citep{bouckaert2014beast2}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes}, LAMARC \citep{kuhner2006lamarc}, Migrate \citep{beerli2006comparison} and possibly other MCMC programs.
+It can be used to analyse runs of BEAST \citep{drummond2012bayesian,drummond2012bayesian}, BEAST2 \citep{bouckaert2014beast2}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes}, LAMARC \citep{kuhner2006lamarc}, Migrate \citep{beerli2006comparison} and possibly other MCMC programs.
 

### 2017-08-03 manuscript/tracer16.tex
edit introduction

@@
 
-Bayesian inference of phylogeny \citep{huelsenbeck2001bayesian, yang2006computational, DrummondBouckaert2015} has been a popular approach to reconstruct the evolutionary history and discover the evolutionary relationships among taxa, whether genes, genomes, individuals or species.
-Bayesian inference can be used to produce the posterior probability distribution over the space of phylogenetic trees given the data.
-Most major software packages currently used employ Markov chain Monte Carlo (MCMC) to produce a sample from the posterior distribution.
-Visualising, summarizing and marginalising these posterior samples is necessary for evaluating and reporting a Bayesian phylogenetic analysis.
+Bayesian inference of phylogeny using Markov chain Monte Carlo (MCMC) \citep{rannala1996probability,mau1999bayesian} flourishes as a popular approach to uncover the evolutionary relationships among taxa, whether genes, genomes, individuals or species.
+MCMC approaches generate samples of model parameter values including the phylogenetic tree drawn from their posterior distribution given molecular sequence data.
+Visualising, tabulating and marginalising these samples is critical for approximating the posterior quantities of interest that one reports in a Bayesian phylogenetic analysis.
 %However, the full posterior distribution is high-dimensional and often difficult to visualise.
-We then develop a software package Tracer (version 1.6) for this purpose, which analyses Bayesian MCMC log files containing the list of parameters whose values sampled from the chain.
-It can be used to analyse runs of BEAST \citep{drummond2012bayesian}, BEAST2 \citep{bouckaert2014beast2}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes}, LAMARC \citep{kuhner2006lamarc}, Migrate \citep{beerli2006comparison},  and possibly other MCMC programs.
-In addition, Tracer's demographic reconstruction functions are utilised to redraw the distribution of effective population sizes over time if an available demographic model was involved in the analysis.
+To facilitate this task, we develop the Tracer (version 1.6) software package that parses MCMC log files of samples to interactively explore the high-dimensional posterior distribution.
+%
+%containing the list of parameters whose values sampled from the chain.
+Tracer works automatically with sample output from
+%It can be used to analyse runs of
+BEAST 1 \citep{drummond2012bayesian}, BEAST 2 \citep{bouckaert2014beast2}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes}, LAMARC \citep{kuhner2006lamarc}, Migrate \citep{beerli2006comparison} and possibly other MCMC programs from other domains.
+
+(move to somewhere else) In addition, Tracer's demographic reconstruction functions are utilised to redraw the distribution of effective population sizes over time if an available demographic model was involved in the analysis.
 

### 2017-08-03 manuscript/tracer16.tex
smoother intro for demo reconstruction

@@
 \end{figure}
+
+A key feature of Tracer lies in its ability to link together samples from quantitative random variable log files and tree log files.
+This linkage is not available other popular post-processing MCMC software, such as coda \cite{citeref}, and enables users to approximate the posterior distribution of scientifically relevant quantities that are functions of both the tree and other quantitative measures.
+Demographic reconstruction through time is a prime example.
 

### 2017-08-03 manuscript/tracer16.tex
add other similar software

--- 2017-08-03 manuscript/tracer16.tex
@@
 \end{figure}
+
+A key feature of Tracer lies in its ability to link together samples from quantitative random variable log files and tree log files.
+This linkage is not available in other popular post-processing MCMC software, such as coda \citep{plummer2006coda} and AWTY \citep{nylander2007awty}, and enables users to approximate the posterior distribution of scientifically relevant quantities that are functions of both the tree and other quantitative measures.
+Demographic reconstruction through time is a prime example.
 

### 2017-08-05 manuscript/tracer16.tex
better title
merge and remove figures; we have too many

@@
 % Title must be 150 characters or less
-\title[Tracer 1.6]{Bayesian MCMC diagnostics and summarisation using Tracer 1.6}
+\title[Tracer 1.6]{Posterior summarisation in Bayesian phylogenetics using Tracer 1.6}
 
@@
 
-\begin{figure}[ht]
-\centerline{
-\includegraphics[width=0.45\textwidth]{./figures/tracer.pdf}
-}
-\caption{The system design of Tracer 1.6}
-\label{fig:tracer}
-\end{figure}
+% MAS: We need to limit the number of figures
+%\begin{figure}[ht]
+%\centerline{
+%\includegraphics[width=0.45\textwidth]{./figures/tracer.pdf}
+%}
+%\caption{The system design of Tracer 1.6}
+%\label{fig:tracer}
+%\end{figure}
 
@@
 \begin{figure}[ht]
-\includegraphics[width=.5\textwidth]{./figures/multitrace.png}
+%\includegraphics[width=.5\textwidth]{./figures/multitrace.png}
+\subfigure[Multiple traces]{\includegraphics[width=0.46\textwidth]{./figures/multitrace.png}}
+\subfigure[TangHuLu chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/tanghulu.pdf}}
+\subfigure[boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/boxplot.pdf}}
 \caption{A comparison of multiple traces}
@@
 
-\begin{figure}[ht]
-\subfigure[TangHuLu chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/tanghulu.pdf}}
-\subfigure[boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/boxplot.pdf}}
-\caption{Visualisation of joint-marginal distribution (a) between two integer or categorical traces using TangHuLu chart, and (b) between one continuous trace and one integer or categorical trace using box-and-whisker plot.}
-\label{fig:tanghulu}
-\end{figure}
+%\begin{figure}[ht]
+%\subfigure[TangHuLu chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/tanghulu.pdf}}
+%\subfigure[boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/boxplot.pdf}}
+%\caption{Visualisation of joint-marginal distribution (a) between two integer or categorical traces using TangHuLu chart, and (b) between one continuous trace and one integer or categorical trace using box-and-whisker plot.}
+%\label{fig:tanghulu}
+%\end{figure}
 
@@
 %\includegraphics[width=.5\textwidth]{./figures/fluBSP.pdf}
-\subfigure[Bayesian coalescent skygrid]{\includegraphics[width = .23\textwidth]{./figures/fig21.png}}
-\subfigure[Lineages through time]{\includegraphics[width = .23\textwidth]{./figures/fluLTT.pdf}}
+%\subfigure[Bayesian coalescent skygrid]{\includegraphics[width = .23\textwidth]{./figures/fig21.png}}
+\centerline{
+\includegraphics[width=0.4\textwidth]{./figures/fig21.png}
+}
+% MAS: Please use vector-graphic format, like PDF
+%\subfigure[Lineages through time]{\includegraphics[width = .23\textwidth]{./figures/fluLTT.pdf}}
 \caption{(a) Bayesian coalescent skygrid plot and (b) lineages through time for the Influenza A/H3N2 virus data.}

### 2017-09-16 manuscript/tracer17.tex
bump version to 1.7

--- 2017-08-05 manuscript/tracer16.tex
@@
 % Title must be 150 characters or less
-\title[Tracer 1.6]{Posterior summarisation in Bayesian phylogenetics using Tracer 1.6}
+\title[Tracer 1.7]{Posterior summarisation in Bayesian phylogenetics using Tracer 1.7}
 
@@
 \section{Results:}
-Here we present a software package \emph{Tracer} (version 1.6) for visualising and analysing the trace files generated by Bayesian MCMC runs.
+Here we present a software package \emph{Tracer} (version 1.7) for visualising and analysing the trace files generated by Bayesian MCMC runs.
 The \emph{Tracer} software also provides kernel density estimation, cheap marginal likelihood estimation,
@@
 %However, the full posterior distribution is high-dimensional and often difficult to visualise.
-To facilitate this task, we develop the Tracer (version 1.6) software package that parses MCMC log files of samples to interactively explore the high-dimensional posterior distribution.
+To facilitate this task, we develop the Tracer (version 1.7) software package that parses MCMC log files of samples to interactively explore the high-dimensional posterior distribution.
 %
@@
 %}
-%\caption{The system design of Tracer 1.6}
+%\caption{The system design of Tracer 1.7}
 %\label{fig:tracer}
@@
 If multiple trace files have the same trace names,  then a ``Combined'' trace will automatically appear, which can be selected as well as the individual trace files.
-
-\begin{figure}[ht]
-%\includegraphics[width=.5\textwidth]{./figures/multitrace.png}
-\subfigure[Multiple traces]{\includegraphics[width=0.46\textwidth]{./figures/multitrace.png}}
-\subfigure[TangHuLu chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/tanghulu.pdf}}
-\subfigure[boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/boxplot.pdf}}
-\caption{A comparison of multiple traces}
-\label{fig:multitrace}
-\end{figure}
 
@@
 Furthermore, box-and-whisker plots are used to display the joint-marginal distribution between one continuous trace and one integer or categorical trace as shown in Figure~\ref{fig:tanghulu}.b.
+
+%
+\begin{figure}[ht]
+%\includegraphics[width=.5\textwidth]{./figures/multitrace.png}
+\subfigure[Multiple traces]{\includegraphics[width=0.46\textwidth]{./figures/multitrace.png}}
+\subfigure[TangHuLu chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/tanghulu.pdf}}
+\subfigure[boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/boxplot.pdf}}
+\caption{A comparison of multiple traces}
+\label{fig:multitrace}
+\end{figure}
+%
 
@@
 \bibliographystyle{natbib}
-\bibliography{tracer16}
+\bibliography{tracer17}
 

### 2017-09-16 src/tracer/traces/JointDensityPanel.java
reverting back to previous constructor until problem has been solved

@@ -424,6 +424,7 @@ public class JointDensityPanel extends TraceChartPanel {
                     for (String two : covarianceData.getTraceNames()) {
                         //System.out.println("adding CovariancePlot: (" + one + "," + two + ")");
                         Plot plot = new CovariancePlot(two, covarianceData.getDataForKey(one), covarianceData.getDataForKey(two));
+                        //Plot plot = new CovariancePlot(two, covarianceData.getDataForKey(one), covarianceData.getDataForKey(two), pointsCheckBox.isSelected(), false);
                         //plot.setLineStyle(new BasicStroke(2.0f), currentSettings.palette[0]);
                         getChartPanel().getChart().addPlot(plot);
                     }

### 2017-09-16 src/tracer/traces/JointDensityPanel.java
switching between box plot and covariance plot is problematic

@@ -165,7 +165,6 @@ public class JointDensityPanel extends TraceChartPanel {
         }
     }
 
-
     protected JChart getChart() {
         if (currentType == Type.BOXPLOT) {
             return traceChart;
@@ -222,7 +221,7 @@ public class JointDensityPanel extends TraceChartPanel {
         JToolBar toolBar = super.createToolBar();
 
         // toolbar empty at the moment...
-        
+
         return toolBar;
     }
 
@@ -262,15 +261,15 @@ public class JointDensityPanel extends TraceChartPanel {
     // it was private void setupChartOrTable()
     protected void setupTraces() {
 
-        //System.out.println("setupTraces(): " + traceNames.size());
-        //System.out.println("covariance entries = " + covarianceData.numberOfEntries());
+        System.out.println("setupTraces(): " + traceNames.size());
+        System.out.println("covariance entries = " + covarianceData.numberOfEntries());
 
         if (traceNames != null && traceNames.size() <= 2) {
 
-            getChartPanel().removeAll();
-//            if (currentType == Type.COVARIANCE) {
-//                if (!removeAllPlots(false)) return;
-//            }
+            //getChartPanel().removeAll();
+            if (currentType == Type.COVARIANCE) {
+                if (!removeAllPlots(false)) return;
+            }
 
             currentType = Type.BOXPLOT;
 
@@ -374,9 +373,9 @@ public class JointDensityPanel extends TraceChartPanel {
 
         } else {
 
-//            if (currentType == Type.BOXPLOT) {
+            if (currentType == Type.BOXPLOT) {
                 getChartPanel().removeAll();
-//            }
+            }
 
             currentType = Type.COVARIANCE;
 
@@ -395,7 +394,7 @@ public class JointDensityPanel extends TraceChartPanel {
                     int traceIndex = tl.getTraceIndex(traceName);
                     Trace trace = tl.getTrace(traceIndex);
                     TraceCorrelation td = tl.getCorrelationStatistics(traceIndex);
-                    Plot plot = null;
+                    //Plot plot = null;
 
                     if (trace != null) {
                         String name = tl.getTraceName(traceIndex);

### 2017-12-15 manuscript/tracer17.tex
rewrote ms. switched to a single example data set. remade figures

--- 2017-08-03 manuscript/tracer16.tex
@@
 % Title must be 150 characters or less
-\title[Tracer 1.6]{Bayesian MCMC diagnostics and summarisation using Tracer 1.6}
+\title[Tracer 1.7]{Posterior summarisation in Bayesian phylogenetics using Tracer 1.7}
 
-\author[Rambaut \textit{et~al}]{ Andrew Rambaut\,$^{1}$, Alexei J.~Drummond\,$^{2,3}$, Dong Xie\,$^{2,3}$, Marc A.~Suchard\,$^{4,5}$}
+\author[Rambaut \textit{et~al.}]{ Andrew Rambaut\,$^{1}$, Guy Baele$^{2}$,  Dong Xie\,$^{3,4}$, Marc A.~Suchard\,$^{5,6,7}$, Alexei J.~Drummond\,$^{3,4}$}
 
@@
 $^{1}$Institute of Evolutionary Biology, University of Edinburgh, Edinburgh, UK\\
-$^{2}$Department of Computer Science, University of Auckland, Auckland, NZ\\
-$^{3}$Centre for Computational Evolution, University of Auckland, Auckland, NZ\\
-$^{4}$Departments of Biomathematics and Human Genetics, David Geffen School of Medicine, UCLA, and \\
-$^{5}$Department of Biostatistics, UCLA Fielding School of Public Health, University of California, Los Angeles, USA \\
+$^{2}$Department of Microbiology and Immunology, Rega Institute, KU Leuven, Leuven, Belgium\\
+$^{3}$Department of Computer Science, University of Auckland, Auckland, NZ\\
+$^{4}$Centre for Computational Evolution, University of Auckland, Auckland, NZ\\
+$^{5}$Departments of Biomathematics and Human Genetics, David Geffen School of Medicine, University of California, Los Angeles, USA \\
+$^{6}$Department of Biostatistics, UCLA Fielding School of Public Health, University of California, Los Angeles, USA \\
+$^{7}$Department of Human Genetics, David Geffen School of Medicine at UCLA, University of California, Los Angeles, USA\\
 }
@@
 \section{Motivation:}
-%Computational evolutionary biology, statistical phylogenetics and coalescent-based population genetics are becoming increasingly central to the analysis and understanding of molecular sequence data.
 Bayesian inference of phylogeny using Markov chain Monte Carlo (MCMC) plays a central role in understanding evolutionary history and processes from molecular sequence data.
@@
 \section{Results:}
-Here we present a software package \emph{Tracer} (version 1.6) for visualising and analysing the trace files generated by Bayesian MCMC runs.
-The \emph{Tracer} software also provides kernel density estimation, cheap marginal likelihood estimation,
-multivariate visualisation, demographic trajectory reconstruction, conditional posterior distribution summary and more.
+Here we present a software package \emph{Tracer} (version 1.7) for visualising and analysing the trace files generated through Bayesian inference.
+The \emph{Tracer} software also provides kernel density estimation, multivariate visualisation, demographic trajectory reconstruction, conditional posterior distribution summary and more.
 \section{Availability:}
-Tracer is open-source under the GNU lesser general public license and available at
-\url{http://beast.bio.ed.ac.uk/Tracer} and \url{https://github.com/beast-dev/tracer}.
+Tracer is open-source under the GNU lesser general public license and available at \url{http://http://beast.community/tracer} and \url{https://github.com/beast-dev/tracer}.
 
@@
 
-Bayesian inference of phylogeny using Markov chain Monte Carlo (MCMC) \citep{rannala1996probability,mau1999bayesian} flourishes as a popular approach to uncover the evolutionary relationships among taxa, whether genes, genomes, individuals or species.
-MCMC approaches generate samples of model parameter values including the phylogenetic tree drawn from their posterior distribution given molecular sequence data.
-Visualising, tabulating and marginalising these samples is critical for approximating the posterior quantities of interest that one reports in a Bayesian phylogenetic analysis.
-%However, the full posterior distribution is high-dimensional and often difficult to visualise.
-To facilitate this task, we develop the Tracer (version 1.6) software package that parses MCMC log files of samples to interactively explore the high-dimensional posterior distribution.
-%
-%containing the list of parameters whose values sampled from the chain.
-Tracer works automatically with sample output from
-%It can be used to analyse runs of
-BEAST 1 \citep{drummond2012bayesian}, BEAST 2 \citep{bouckaert2014beast2}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes}, LAMARC \citep{kuhner2006lamarc}, Migrate \citep{beerli2006comparison} and possibly other MCMC programs from other domains.
-
-(move to somewhere else) In addition, Tracer's demographic reconstruction functions are utilised to redraw the distribution of effective population sizes over time if an available demographic model was involved in the analysis.
-
-%\begin{itemize}
-%\item Importance of posterior summaries (as scientifically relevant marginalization of the full posterior distribution)
-%\item Full posterior distribution is high-dimensional and (often) difficult to visualize
-%\item Posterior simulator agnostic -- highly used with \textsc{MrBayes} as well
-%\end{itemize}
+Bayesian inference of phylogeny using Markov chain Monte Carlo (MCMC) \citep{rannala1996probability, mau1999bayesian} flourishes as a popular approach to uncover the evolutionary relationships among operational taxonomic units (OTUs), such as genes, genomes, individuals or species.
+MCMC approaches generate samples of model parameter values - including the phylogenetic tree - drawn from their posterior distribution given molecular sequence data and a selection of evolutionary models.
+Visualising, tabulating and marginalising these samples is critical for approximating the posterior quantities of interest that one reports as the outcome of a Bayesian phylogenetic analysis.
+To facilitate this task, we have developed the Tracer (version 1.7) software package, which allows users to process MCMC log files containing parameter samples and to interactively explore the high-dimensional posterior distribution.
+Tracer works automatically with sample output from BEAST 1 \citep{drummond2012bayesian}, BEAST 2 \citep{bouckaert2014beast2}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes}, LAMARC \citep{kuhner2006lamarc}, Migrate \citep{beerli2006comparison} and possibly other MCMC programs from other domains.
 
@@
 
-\begin{figure}[ht]
-\centerline{
-\includegraphics[width=0.45\textwidth]{./figures/tracer.pdf}
-}
-\caption{The system design of Tracer 1.6}
-\label{fig:tracer}
-\end{figure}
+Tracer is able to analyse a single MCMC log and can also combine posterior samples from multiple trace files stemming from an independent evaluation of the same dataset with an identical model combination.
+For each trace file loaded, Tracer analyses the posterior samples from all the available parameters, regardless of whether these are discrete or continuous, and present statistical summaries as well as visualisation options for all of these parameters.
+Importantly, the effective sample size (ESS) - an important measurement designed for a sampled parameter having continuous values - is immediately apparent in the default view of Tracer and allows users to assess the number of effectively independent draws from the posterior distribution for that parameter (see Figure \ref{fig:overview}:a).
+Colour coding is provided for the ESS values to assist the user in determining potential mixing problems in the Bayesian inference performed, with cut-off values at 100 and 200.
 
-Tracer is able to analyse either a single MCMC log or combine the shared traces from multiple logs as designed in Figure~\ref{fig:tracer}.
-The list of traces in each log are created from the sampling parameters logged during a MCMC process, which include their statistical summaries from their values at all states logged.
-The effective sample size (ESS) is an important measurement designed for a sampled parameter having continuous values, which tells the number of effectively independent draws from the posterior distribution that the Markov chain is equivalent to.
+Tracer is able to handle three trace types, also known as variable types \citep{mendenhall2012introduction}: real, integer, and categorical.
+The real trace type is used to handle quantitative continuous variables, which commonly appear in phylodynamic analyses, whereas the integer trace type allows to deal with quantitative discrete variables and the categorical trace is meant for qualitative categorical variables.
+The three trace types are automatically assigned when the trace file is imported but can also manually be set.
 
-Tracer handles three trace types, also known as variable types \citep{mendenhall2012introduction}: real, integer, and categorical.
-The real trace is used to handle quantitative continuous variables, which is the most common type in logged parameters.
-The integer is dealing with quantitative discrete variables, and the categorical trace is for qualitative categorical variables.
-% check above definition
-%These three types can be automatically assigned when the log file is imported. But they can also be changed later.
+Multiple parameter traces can be selected from the traces panel, which will result in a side-by-side comparison or overlay of the selected parameters' visualisations for the different traces (see Figure \ref{fig:overview}:b-e).
+Multiple trace files can be selected in a similar fashion to compare posterior samples between different replicates of an analysis. 
+If multiple trace files have the same file name,  then a ``Combined'' trace will automatically appear, which can be selected as well as the individual trace files.
+Tracer consists of four panels that allow to present both individual and joint visualisation options for either a single or multiple selected trace files:
 
-The multiple traces can be selected as shown in Figure~\ref{fig:multitrace}, which will overlay the plots for the different traces allowing comparisons to be made.
-The multiple trace files can be similarly selected as well to compare different runs.
-If multiple trace files have the same trace names,  then a ``Combined'' trace will automatically appear, which can be selected as well as the individual trace files.
+\begin{itemize}
 
-\begin{figure}[ht]
-\includegraphics[width=.5\textwidth]{./figures/multitrace.png}
-\caption{A comparison of multiple traces}
-\label{fig:multitrace}
-\end{figure}
+\item Estimates: this panel reports common summary statistics such as the mean, standard deviation, highest posterior density interval, ESS, and other statistics concerning the selected parameter(s). 
+A histogram presenting the frequency distribution over values will be shown for a single selected parameter of any trace type (see Figure \ref{fig:overview}:a), whereas side-by-side boxplots will be shown when multiple continuous parameters are selected (see Figure \ref{fig:overview}:b).
 
-There are four analysis tabs to choose from either a selected trace (sampling parameter) or multiple traces:
+\item Marginal Density: this panel shows a density plot for the posterior sample of the selected parameter(s).
+Kernel density estimation (see Figure \ref{fig:overview}:c), histograms and violin plots (see Figure \ref{fig:overview}:d) are available for continuous traces, whereas frequency plots are employed for integer traces.
 
-%TODO simplify, some duplication to 1st example
-\begin{itemize}
-\item Estimates - this shows the mean, standard deviation, confidence intervals, ESS, and other statistics about the selected trace(s).
-A histogram presenting the frequency distribution over values will also be plotted for a single selected trace, such as Figure~\ref{fig:4tabs}.a.
-Interval bars for quantitative traces (real or integer) or violin plots for categorical traces will be drawn for a comparison, such as Figure~\ref{fig:multitrace}, if multiple traces are selected.
+\item Joint-Marginal: visualisation in this panel only appears when two or more parameters are selected.
+The actual visualisation depends on the trace type of the selected parameters.
+We show several examples in the next section of the paper.
 
-\item Density - this shows the Bayesian posterior density plot for the selected trace(s), if they are continuous, the kernel density estimates will be available for continuous traces, for example, in Figure~\ref{fig:4tabs}.b.
-
-\item Joint-Marginal - this only appears if exactly two traces are chosen. It then plots one against the other to look at their joint-marginal distribution, for example, in Figure~\ref{fig:4tabs}.c.
-
-% trace is duplicate to trace, use trajectory?
-\item Trace - this shows the trajectory of the selected trace(s) against state or generation number as it can be seen in Figure~\ref{fig:4tabs}.d.
-It can be used to check mixing, choose a suitable burn-in and look for trends that might suggest problems with convergence.
+\item Trace: this panel shows a line plot connecting the posterior samples of one or more selected parameters against state or generation number (see Figure \ref{fig:overview}:e).
+Such a trace plot is typically used to assess mixing, select a suitable burn-in and identify trends that suggest convergence issues.
 
@@
 
-In addition to illustrating joint-marginal distribution between two continuous traces, we newly invented a TangHuLu chart to visualise the joint probability between two integer or categorical traces using coloured bubbles, as shown in Figure~\ref{fig:tanghulu}.a.
-The size of circle is proportional to the joint probability, the blue coloured circle is in the credible set of given a probability threshold default to 0.95, the red is in the set not in the credible set.
-The tile background is coloured if there is any circle, in case the very small circle wouldn't be missed out.
-The coloured background can also help to show the area of the credible set and non-credible set.
+Tracer offers an elegant solution of visualising conditional posterior distributions. 
+A typical use case involves support for Bayesian stochastic search variable selection (BSSVS), a form of model averaging, in which parameters appear in the likelihood function only when a specific model is indicated by an indicator function.
+When BSSVS is used, the posterior distribution of the parameter should not include those states that were sampled from the prior when its model was not part of the likelihood.
+Discrete phylogeographic analyses frequently employ such an approach due to the potentially large amount of transition rates that need to be estimated \citep{Lemey2009}, but this is also relevant when employing model averaging approaches such as for relaxed molecular clocks for example \citep{Li2012}.
 
-Furthermore, box-and-whisker plots are used to display the joint-marginal distribution between one continuous trace and one integer or categorical trace as shown in Figure~\ref{fig:tanghulu}.b.
+Additionally, Tracer provides demographic reconstruction resulting in a graphical plot, often applied to reconstruct epidemic dynamics.
+Depending on the actual demographic model used in the analysis, an additional file containing samples from the posterior tree distribution will be required.
+The currently available models are constant size, exponential and logistic growth \citep{drummond2002estimating}, Bayesian coalescent skyline \citep{drummond2005bayesian}, Gaussian Markov Random Field (GMRF) Bayesian skyride \citep{minin2008smooth}, and Bayesian coalescent skygrid \citep{gill2012improving}.
+%Not sure if the statement below is correct ...
+All of these demographic models are available in BEAST 1 \citep{drummond2012bayesian} and BEAST 2 \citep{bouckaert2014beast2}.
+%Besides these, ``Lineages Through Time'' (?) is also available to plot the quantiles of the number of lineages against time, which describes the change of diversification of all organisms over time.
+%citation?
 
 \begin{figure}[ht]
-\subfigure[TangHuLu chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/tanghulu.pdf}}
-\subfigure[boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/boxplot.pdf}}
-\caption{Visualisation of joint-marginal distribution (a) between two integer or categorical traces using TangHuLu chart, and (b) between one continuous trace and one integer or categorical trace using box-and-whisker plot.}
-\label{fig:tanghulu}
+\subfigure[Tracer overview]{\includegraphics[width=0.47\textwidth]{./figures/rabv-overview.pdf}}
+\subfigure[Boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-boxplots.pdf}}
+\subfigure[Density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-density.pdf}}\\
+\subfigure[Violin]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-violin.pdf}}
+\subfigure[Trace]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-traces.pdf}}
+\caption{Overview of Tracer functionality and individual parameter visualisations: (a) Main Tracer panel upon loading a single trace file; (b) boxplot representation of two continuous parameters; (c) kernel density estimates; (d) violin plots; (e) the actual traces connecting the parameter values visited by the Markov chain.}
+\label{fig:overview}
 \end{figure}
-
-A key feature of Tracer lies in its ability to link together samples from quantitative random variable log files and tree log files.
-This linkage is not available in other popular post-processing MCMC software, such as coda \citep{plummer2006coda} and AWTY \citep{nylander2007awty}, and enables users to approximate the posterior distribution of scientifically relevant quantities that are functions of both the tree and other quantitative measures.
-Demographic reconstruction through time is a prime example.
-
-Tracer also provides demographic reconstruction resulting in a graphical plot, often applyed to reconstruct epidemic dynamics, given a MCMC log from a Bayesian framework or additionally with the tree log depending on the model.
-The currently available models are constant size \citep{drummond2002estimating}, exponential growth \citep{drummond2002estimating}, logistic growth (?), Bayesian coalescent skyline \citep{drummond2005bayesian}, GMRF Bayesian skyride \citep{minin2008smooth}, and Bayesian coalescent skygrid \citep{gill2012improving}.
-%For example, ``Demographic Reconstruction" performs reconstructions of the distribution for a number of models, such as constant size \citep{drummond2002estimating}, exponential growth \citep{drummond2002estimating}, and logistic growth (?).
-%citation?
-%Besides ``Bayesian Skyline Reconstruction'' reconstructs a generalised Bayesian coalescent skyline plot (e.g.
-Figure~\ref{fig:flu}.a) describing demographic history under canonical scenarios \citep{drummond2005bayesian},
-%``GMRF Skyride Reconstruction'' and ``SkyGrid Reconstruction'' are respectively available for a temporal smoothing method using the Gaussian Markov random field (GMRF) model  \citep{minin2008smooth} and a generalisation of the GMRF model \citep{gill2012improving}.
-Besides these, ``Lineages Through Time'' (?) is also available to plot the quantiles of the number of lineages against time, which describes the change of diversification of all organisms over time.
-%citation?
-%All these models are available in BEAST \citep{drummond2007beast}.
-
-%TODO is it OK ?
-Tracer recently offers a fairly general solution of looking at conditional posterior distributions.
-This, specifically, supports for Bayesian stochastic search variable selection (BSSVS) forms of model averaging, in which some parameters are only in the likelihood when their sub-model is indicated by an indicator function usually assigned to a discrete, integer or boolean variable.
-In that case, the posterior of the parameter should not include the states when it was sampled only in the prior, because the sub-model it belonged to was ``turned off".
-This is relevant for extended Bayesian skyline plot, random local clock model, microsatellite model averaging, and relaxed clock model averaging methods, etc.
-It is also relevant for BSSVS in phylogeography as well depending on how the state is logged.
-
-% TODO more ?
 
@@
 
-%Even users of \textsc{MrBayes} and \textsc{RevBayes} find Tracer handy.
+\subsection*{Cross-species dynamics of rabies virus in North American bats}
 
-\subsection*{Calibrating the tree to estimate TMRCA}
-
-Figure~\ref{fig:4tabs} demonstrates a snapshot of posterior summaries in Tracer for a BEAST \citep{drummond2012bayesian} run created from a 898 bp alignment of sequences of 12 species of primates (http://beast.bio.ed.ac.uk/tutorials?).
-%data citation
-
-%using HKY + $\Gamma_4$ model, uncorrelated lognormal relaxed molecular clock, and Yule prior on the tree.
-In this analysis, we furthermore specified three calibrations on the phylogenetic tree to estimate time to the most recent common ancestor (TMRCA) based on our prior fossil knowledge.
-
-The first TMRCA prior, called ``ingroup'', includes all species except the \textit{Lemur} which will form the outgroup in the tree. %monophyletic
-The second group is ``Human-Chimp'' that contains only \textit{Homo sapiens} and \textit{Pan}.
-The last consists of everything except \textit{Lemur}, \textit{Saimiri} and \textit{Tarsius} to create the hominoid-cercopithecoid split, which is named as ``HomiCerco''.
-We know the most recent common ancestor of humans and chimps may live about $5-7$ million years ago, a normal distribution centered at 6 million years with a standard deviation of 0.5 million years is assigned to the ``Human-Chimp'' prior.
-Another normal distribution of $24 \pm 0.5$ (million years) is likewise defined for ``HomiCerco''.
+We provide on example on reconstructing the spatial dispersal and cross-species dynamics of rabies virus (RABV) in North American bat populations based on a set of 372 nucleoprotein gene sequences (nucleotide positions: 594-1353).
+The data set comprises a total of 17 bat species sampled between 1997 and 2006 across 14 states in the United States \citep{Streicker}, including two additional species that had been excluded from the original analysis owing to a limited amount of available sequences \citep{Faria2013}.
+%We also include a viral sequence with an unknown sampling date (accession no. TX5275, sampled in Texas from Lasiurus borealis), which will be adequately accommodated in our inference.
+We estimate the ancestral locations of RABV using a Bayesian discrete phylogeographic approach and, at the same time, infer the history of host jumping using the same model approach \citep{Lemey2009}.
+We employ the Bayesian skygrid coalescent model \citep{gill2012improving} to estimate the effective population sizes over time.
 
 \begin{figure}[ht]
-\subfigure[Estimates]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/treeRootHeight.pdf}}
-\subfigure[Density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/multiKDE.pdf}}\\
-\subfigure[Joint-Marginal]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/joint-marginal-th-ingroup.pdf}}
-\subfigure[Trace]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/trace.pdf}}
-\caption{Images exported from four analysis tabs in Tracer for calibrating the tree to estimate TMRCA, where the black is ``treeModel.rootHeight'' parameter, blue is TMRCA parameter of ``Human-Chimp'', red is ``ingroup'', and yellow is ``HomiCerco''.}
+\subfigure[TangHuLu chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integerjointmarginal.pdf}}
+\subfigure[Marginal density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integermarginaldensity.pdf}} \\
+\subfigure[Scatter plot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-correlation.pdf}}
+\subfigure[Correlation plot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-joint-real.pdf}} \\
+\subfigure[Box-and-whisker plot]{\includegraphics[width = .465\textwidth, height = 3cm]{./figures/rabv-catreal.pdf}}
+\caption{Joint marginal and marginal density plots of: (a) two integer variables through a TangHuLu chart; (b) frequency plots the marginal density of multiple integer or categorical variables; (c) classic scatter or correlation plot for two continuous variables; (d) large correlation matrices for multiple ($> 2$) continuous variables; (e) box-and-whisker plots for the joint marginal distribution of a continuous and an integer or categorical variable.}
 \label{fig:4tabs}
@@
 
-After loading the log, selecting the ``treeModel.rootHeight'' parameter gives the marginal posterior distribution of the age of the root of entire tree.
-The blue bars in Figure~\ref{fig:4tabs}.a present the most compact interval on the root height of the tree that contains 95\% of the posterior probability, referring to the period between about 39.01 and 63.67 million years old with the mean of 50.48 million years.
-%TODO rate is faster than Yang's book r_ape = 5.4 * 10^-10 per year, r_monkey = 8.0, r_root = 6.6
-Correspondingly selecting ``meanRate'' provides the rate of evolution averaged over the whole tree, which has the mean of 0.0129 substitutions per site per million years with 95\% HPD Interval between 0.0101 and 0.0159.
+Performing BSSVS in a discrete phylogeographic approach provides us with parameters of the integer and categorical trace type.
+While a single parameter of such types can easily be visualised using a frequency plot, we have developed a TangHuLu chart to visualise the joint probability between two integer or categorical traces using coloured bubbles (see Figure \ref{fig:4tabs}:a).
+The size of circle drawn is proportional to the joint probability, %the blue coloured circle is in the credible set of given a probability threshold default to 0.95, the red is in the set not in the credible set.
+with the tile background being coloured if there is in fact a circle, to enhance visibility. 
+%The coloured background can also help to show the area of the credible set and non-credible set.
+Multiple integer parameters can also be visualised using marginal density plots (see Figure \ref{fig:4tabs}:b), with a different colour scale being assigned to each parameter.
+Both figures show a clear overlap in the number of non-zero transition rates in the discrete trait models for host and location, as identified using a BSSVS procedure.
+The visualisation options for integer or categorical parameters are complemented with popular plots for continuous variables, such as a scatter or correlation plot (see Figure \ref{fig:4tabs}:c) for two continuous variables, which indicates a strong positive correlation between the clock rates of both discrete trait models.
+We also provide an extension of such a scatter plot for multiple (i.e.
+more than two) continuous variables (see Figure \ref{fig:4tabs}:d; \citet{Murdoch}), which makes use of coloured ellipses to indicate the strength of correlation between each pair of continuous variables.
+Colour gradients indicate the strength and direction of the correlation, with a colour gradient from white to dark blue indicating increasingly strong correlation and a colour gradient from white to dark red indicating increasingly strong anti-correlation.
+Additionally, the strength of correlation is reflected in the shape of the ellipse drawn, with no correlation being drawn as a circle and perfect (anti-)correlation being drawn as a line.
+Finally, in order to display the joint marginal distribution between a continuous variable and an integer or categorical variable, we make use of box-and-whisker plots.
+Using such a plot (Figure \ref{fig:4tabs}:e), we can infer that the range of the non-zero transition rates can depend on the ancestral location state inferred at the root of the underlying phylogeny.
 
-Figure~\ref{fig:4tabs}.b shows the kernel density estimations of these four parameters.
-The two TMRCA parameters, ``Human-Chimp'' in blue and ``HomiCerco'' in yellow, that we used to calibrate the tree explicitly have posterior distributions very similar to the prior distributions that we specified.
-
-The marginal posterior densities of ``treeModel.rootHeight'' and ``ingroup'' are nearly overlapped each other.
-When we switch to their joint-marginal distribution plot in Figure~\ref{fig:4tabs}.c, the samples are almost lined up together, which indicates they are associated.
-But this isn't an issue here, because the ``ingroup'' only excludes one outgroup species from all taxa.
-%TODO how to explain this problem further?
-
-Figure~\ref{fig:4tabs}.d illustrates the traces of these four parameters against their logged states from MCMC chain.
-``Human-Chimp'' and ``HomiCerco'' parameters are immediately reaching equilibrium from the starting point, but ``treeModel.rootHeight'' and ``ingroup'' clearly show a burn-in period.
-%TODO ?
-
-\subsection*{Non-parametric coalescent estimator}
-
-In the second example, there are two analyses conducted using a H3N2 data set that is the subset of a comprehensive influenza A virus data set spanning several epidemic seasons in the New York state.
-The data contains a  ?bp alignment of 165 Hemagglutinin genes collected between about 1999 and 2004 \citep{rambaut2008genomic}, explicating H3N2 spreading during three epidemic seasons.
-
-To be able to reconstruct Bayesian coalescent skygrid \citep{minin2008smooth} plot, we exploit Tracer to exact past population dynamics from both MCMC log and tree log produced by a BEAST \citep{drummond2012bayesian} run.
-The HKY + $\Gamma_4$ model, strict molecular clock, and Bayesian skygrid tree prior was respectively setup in the run, where a grid of 50 intervals over 5 years estimating 10 population sizes by year was constructed in the skygrid prior.
-
-%What type of dynamics does the H3N2 skyride plot suggest?
-According to Tracer's skygrid plot in Figure~\ref{fig:flu}.a, the changing patterns of genetic diversity in viral isolates from New York state clearly reveal the seasonal dynamics of influenza A in individual temperate populations.
-The absolute amount of genetic diversity, even at seasonal peaks, is small compared to other rapidly evolving viruses that infect far fewer people, suggesting that strong natural selection, in addition to periodic bottlenecks, reduces the level of diversity that co-circulates at any time.
-Simulations demonstrate that this reconstruction of genetic diversity is robust to the sampling protocol.
-%TODO copied from (Rambaut et al., 2008)
+The Bayesian skygrid coalescent model \citep{gill2012improving} allows us to use Tracer to reconstruct the demographic history of RABV in North America by drawing the effective population sizes over time (see Figure \ref{fig:rabv}).
+Given the properties of this coalescent prior, which requires to specify a grid of intervals over the timeframe of the epidemic, Tracer is able to infer this history without requiring a posterior sample of trees.
+According to Tracer's skygrid plot, it is clear that rabies virus has successfully established itself in North American bat species, with its numbers rising steadily throughout recent centuries.
+Following a decline in numbers at the end of last century, we observe a recent sharp increase in population size of RABV in North American bats.
 
 \begin{figure}[ht]
-%\includegraphics[width=.5\textwidth]{./figures/fluBSP.pdf}
-\subfigure[Bayesian coalescent skygrid]{\includegraphics[width = .23\textwidth]{./figures/fig21.png}}
-\subfigure[Lineages through time]{\includegraphics[width = .23\textwidth]{./figures/fluLTT.pdf}}
-\caption{(a) Bayesian coalescent skygrid plot and (b) lineages through time for the Influenza A/H3N2 virus data.}
-\label{fig:flu}
+\centerline{
+\includegraphics[width=0.46\textwidth]{./figures/rabv-skygrid.pdf}
+}
+\caption{Estimating the effective population sizes over time using a Bayesian skygrid demographic reconstruction for rabies virus in North America.}
+\label{fig:rabv}
 \end{figure}
 
-Lineages Through Time in Figure~\ref{fig:flu}.b, similar curve
+\section*{Availability}
 
-diversification has been gradually developed before 2003's break-out, but it also looks stopped after 2004.
-
-\section*{Availability and Future Directions}
-
-We make the Tracer package available in both executable and source code forms.
-Tracer requires Java version 1.6 or greater and executables for Windows, Mac OS and Linux platforms are located at \url{http://beast.bio.ed.ac.uk/Tracer} % \url{https://github.com/beast-dev/tracer} ?
-which serves as the main page for the package.
+Tracer 1.7 is open-source under the GNU lesser general public license and available in both executable and source code forms at \url{http://beast.community/tracer}.
+It requires Java version 1.6 or greater.
+Github houses the Tracer's source code within \url{https://github.com/beast-dev/tracer}, and links to a GoogleGroup discussion group related to Tracer, i.e.
+the ``beast-users" group (\url{http://groups.google.com/group/beast-users}).
 This page also links to a sizeable list of self-contained, step-by-step tutorials covering basic to advance usage of Tracer to summarise posterior distributions of a large set of phylogenetic models simulated using BEAST.
 For example, popular tutorials describe how to use Tracer to generate marginal parameter summaries and infer population dynamics trajectories over time.
-
-Github houses the Tracer's source code within \url{https://github.com/beast-dev/tracer} ,and links to a GoogleGroup discussion group related to Tracer,
-which is the ``beast-users" group (\url{http://groups.google.com/group/beast-users}) with over 1,500 members.
-%At the time of writing, forty-seven developers belong to the ``beast-dev" group that facilitates BEAST development across three continents.
-
-Future development directions for Tracer focus on \ldots
 
@@
 
-\section*{Acknowledgments}
+\section*{Acknowledgements}
 
-This work was supported in part by
-the Wellcome Trust through project 206298/Z/17/Z (Artic Network),
-the European Union Seventh Framework Programme under grant agreement no. 725422-RESERVOIRDOCS,
-the National Science Foundation through grant DMS 1264153,
-the National Institutes of Health under grant R01 AI107034,
-and
-the Marsden Trust.
+This work was supported in part by the Wellcome Trust through project 206298/Z/17/Z (Artic Network), the European Union Seventh Framework Programme under grant agreement no.
+725422-RESERVOIRDOCS, the National Science Foundation through grant DMS 1264153, the National Institutes of Health under grant R01 AI107034, and the Marsden Trust.
+GB acknowledges support from the Interne Fondsen KU Leuven / Internal Funds KU Leuven.
 

### 2017-12-15 manuscript/tracer17.tex
remove white-spaces

@@
 
-\author[Rambaut \textit{et~al.}]{ Andrew Rambaut\,$^{1}$, Guy Baele$^{2}$,  Dong Xie\,$^{3,4}$, Marc A.~Suchard\,$^{5,6,7}$, Alexei J.~Drummond\,$^{3,4}$}
+\author[Rambaut \textit{et~al.}]{
+Andrew Rambaut\,$^{1}$,
+Guy Baele$^{2}$,
+Dong Xie\,$^{3,4}$,
+Alexei J.~Drummond\,$^{3,4}$,
+Marc A.~Suchard\,$^{5,6,7}$
+}
 
@@
 %%\includegraphics[width=.5\textwidth]{./figures/joint-marginal.pdf}
-%%\caption{The joint-marginal distribution of two selected traces}
+%%\caption{The joint-marginal distribution of two selected tes}
 %%\label{fig:trace}

### 2017-12-15 manuscript/tracer17.tex
comment-out large box-and-wisker plot

@@
 \subfigure[Correlation plot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-joint-real.pdf}} \\
-\subfigure[Box-and-whisker plot]{\includegraphics[width = .465\textwidth, height = 3cm]{./figures/rabv-catreal.pdf}}
-\caption{Joint marginal and marginal density plots of: (a) two integer variables through a TangHuLu chart; (b) frequency plots the marginal density of multiple integer or categorical variables; (c) classic scatter or correlation plot for two continuous variables; (d) large correlation matrices for multiple ($> 2$) continuous variables; (e) box-and-whisker plots for the joint marginal distribution of a continuous and an integer or categorical variable.}
+%\subfigure[Box-and-whisker plot]{\includegraphics[width = .465\textwidth, height = 3cm]{./figures/rabv-catreal.pdf}}
+\caption{Joint marginal and marginal density plots of: (a) two integer variables through a TangHuLu chart; (b) frequency plots the marginal density of multiple integer or categorical variables; (c) classic scatter or correlation plot for two continuous variables; (d) large correlation matrices for multiple ($> 2$) continuous variables.
+%; (e) box-and-whisker plots for the joint marginal distribution of a continuous and an integer or categorical variable.
+}
 \label{fig:4tabs}
@@
 Additionally, the strength of correlation is reflected in the shape of the ellipse drawn, with no correlation being drawn as a circle and perfect (anti-)correlation being drawn as a line.
-Finally, in order to display the joint marginal distribution between a continuous variable and an integer or categorical variable, we make use of box-and-whisker plots.
-Using such a plot (Figure \ref{fig:4tabs}:e), we can infer that the range of the non-zero transition rates can depend on the ancestral location state inferred at the root of the underlying phylogeny.
+%Finally, in order to display the joint marginal distribution between a continuous variable and an integer or categorical variable, we make use of box-and-whisker plots.
+%Using such a plot (Figure \ref{fig:4tabs}:e), we can infer that the range of the non-zero transition rates can depend on the ancestral location state inferred at the root of the underlying phylogeny.
 

### 2017-12-15 manuscript/tracer17.tex
capitalize words in references

@@
 Alexei J.~Drummond\,$^{3,4}$,
-Marc A.~Suchard\,$^{5,6,7}$
+Marc A.~Suchard\,$^{5,6}$
 }
@@
 $^{6}$Department of Biostatistics, UCLA Fielding School of Public Health, University of California, Los Angeles, USA \\
-$^{7}$Department of Human Genetics, David Geffen School of Medicine at UCLA, University of California, Los Angeles, USA\\
+%$^{7}$Department of Human Genetics, David Geffen School of Medicine at UCLA, University of California, Los Angeles, USA\\
 }

### 2017-12-15 manuscript/tracer17.tex
edits to abstract

@@
 \section{Motivation:}
-Bayesian inference of phylogeny using Markov chain Monte Carlo (MCMC) plays a central role in understanding evolutionary history and processes from molecular sequence data.
-Visualising and analysing the samples from the posterior distribution is a key step in any non-trivial Bayesian inference, and it is especially important for methods that employ MCMC.
+Bayesian inference of phylogeny using Markov chain Monte Carlo (MCMC) plays a central role in understanding evolutionary history from molecular sequence data.
+Visualising and analysing the MCMC-generated samples from the posterior distribution is a key step in any non-trivial Bayesian inference.
+%, and it is especially important for methods that employ MCMC.
 \section{Results:}
-Here we present a software package \emph{Tracer} (version 1.7) for visualising and analysing the trace files generated through Bayesian inference.
-The \emph{Tracer} software also provides kernel density estimation, multivariate visualisation, demographic trajectory reconstruction, conditional posterior distribution summary and more.
+We present the software package \emph{Tracer} (version 1.7) for visualising and analysing the MCMC trace files generated through Bayesian phylogenetic inference.
+\emph{Tracer} provides kernel density estimation, multivariate visualisation, demographic trajectory reconstruction, conditional posterior distribution summary and more.
 \section{Availability:}
-Tracer is open-source under the GNU lesser general public license and available at \url{http://http://beast.community/tracer} and \url{https://github.com/beast-dev/tracer}.
+\emph{Tracer} is open-source and available at \url{http://http://beast.community/tracer} and \url{https://github.com/beast-dev/tracer}.
 
@@
 This work was supported in part by the Wellcome Trust through project 206298/Z/17/Z (Artic Network), the European Union Seventh Framework Programme under grant agreement no.
-725422-RESERVOIRDOCS, the National Science Foundation through grant DMS 1264153, the National Institutes of Health under grant R01 AI107034, and the Marsden Trust.
+725422-RESERVOIRDOCS, the National Science Foundation through grant DMS 1264153, the National Institutes of Health under grants R01 AI107034 and U19 AI135995, and the Marsden Trust.
 GB acknowledges support from the Interne Fondsen KU Leuven / Internal Funds KU Leuven.

### 2017-12-15 manuscript/tracer17.tex
edits to availability section

@@
 \usepackage{url}
+
+\newcommand{\tracer}{\emph{Tracer}}
 
@@
 $^{4}$Centre for Computational Evolution, University of Auckland, Auckland, NZ\\
-$^{5}$Departments of Biomathematics and Human Genetics, David Geffen School of Medicine, University of California, Los Angeles, USA \\
-$^{6}$Department of Biostatistics, UCLA Fielding School of Public Health, University of California, Los Angeles, USA \\
+$^{5}$Department of Human Genetics, David Geffen School of Medicine, University of California, Los Angeles, USA \\
+$^{6}$Department of Biostatistics, Fielding School of Public Health, University of California, Los Angeles, USA \\
 %$^{7}$Department of Human Genetics, David Geffen School of Medicine at UCLA, University of California, Los Angeles, USA\\
@@
 \section{Results:}
-We present the software package \emph{Tracer} (version 1.7) for visualising and analysing the MCMC trace files generated through Bayesian phylogenetic inference.
-\emph{Tracer} provides kernel density estimation, multivariate visualisation, demographic trajectory reconstruction, conditional posterior distribution summary and more.
+We present the software package \tracer\ (version 1.7) for visualising and analysing the MCMC trace files generated through Bayesian phylogenetic inference.
+\tracer\ provides kernel density estimation, multivariate visualisation, demographic trajectory reconstruction, conditional posterior distribution summary and more.
 \section{Availability:}
-\emph{Tracer} is open-source and available at \url{http://http://beast.community/tracer} and \url{https://github.com/beast-dev/tracer}.
+\tracer\ is open-source and available at \url{http://http://beast.community/tracer}.
 
@@
 
-Bayesian inference of phylogeny using Markov chain Monte Carlo (MCMC) \citep{rannala1996probability, mau1999bayesian} flourishes as a popular approach to uncover the evolutionary relationships among operational taxonomic units (OTUs), such as genes, genomes, individuals or species.
+Bayesian inference of phylogeny using Markov chain Monte Carlo (MCMC) \citep{rannala1996probability, mau1999bayesian} flourishes as a popular approach to uncover the evolutionary relationships among taxa, such as genes, genomes, individuals or species.
 MCMC approaches generate samples of model parameter values - including the phylogenetic tree - drawn from their posterior distribution given molecular sequence data and a selection of evolutionary models.
 Visualising, tabulating and marginalising these samples is critical for approximating the posterior quantities of interest that one reports as the outcome of a Bayesian phylogenetic analysis.
-To facilitate this task, we have developed the Tracer (version 1.7) software package, which allows users to process MCMC log files containing parameter samples and to interactively explore the high-dimensional posterior distribution.
-Tracer works automatically with sample output from BEAST 1 \citep{drummond2012bayesian}, BEAST 2 \citep{bouckaert2014beast2}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes}, LAMARC \citep{kuhner2006lamarc}, Migrate \citep{beerli2006comparison} and possibly other MCMC programs from other domains.
+To facilitate this task, we have developed the \tracer\ (version 1.7) software package to
+process MCMC trace files containing parameter samples and to interactively explore the high-dimensional posterior distribution.
+\tracer\ works automatically with sample output from BEAST 1 \citep{drummond2012bayesian}, BEAST 2 \citep{bouckaert2014beast2},  LAMARC \citep{kuhner2006lamarc},  Migrate \citep{beerli2006comparison}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes} and possibly other MCMC programs from other domains.
 
@@
 
-Tracer is able to analyse a single MCMC log and can also combine posterior samples from multiple trace files stemming from an independent evaluation of the same dataset with an identical model combination.
-For each trace file loaded, Tracer analyses the posterior samples from all the available parameters, regardless of whether these are discrete or continuous, and present statistical summaries as well as visualisation options for all of these parameters.
-Importantly, the effective sample size (ESS) - an important measurement designed for a sampled parameter having continuous values - is immediately apparent in the default view of Tracer and allows users to assess the number of effectively independent draws from the posterior distribution for that parameter (see Figure \ref{fig:overview}:a).
+\tracer\ is able to analyse a single MCMC log and can also combine posterior samples from multiple trace files stemming from an independent evaluation of the same dataset with an identical model combination.
+For each trace file loaded, \tracer\ analyses the posterior samples from all the available parameters, regardless of whether these are discrete or continuous, and present statistical summaries as well as visualisation options for all of these parameters.
+Importantly, the effective sample size (ESS) - an important measurement designed for a sampled parameter having continuous values - is immediately apparent in the default view of \tracer\ and allows users to assess the number of effectively independent draws from the posterior distribution for that parameter (see Figure \ref{fig:overview}:a).
 Colour coding is provided for the ESS values to assist the user in determining potential mixing problems in the Bayesian inference performed, with cut-off values at 100 and 200.
 
-Tracer is able to handle three trace types, also known as variable types \citep{mendenhall2012introduction}: real, integer, and categorical.
+\tracer\ is able to handle three trace types, also known as variable types \citep{mendenhall2012introduction}: real, integer, and categorical.
 The real trace type is used to handle quantitative continuous variables, which commonly appear in phylodynamic analyses, whereas the integer trace type allows to deal with quantitative discrete variables and the categorical trace is meant for qualitative categorical variables.
@@
 If multiple trace files have the same file name,  then a ``Combined'' trace will automatically appear, which can be selected as well as the individual trace files.
-Tracer consists of four panels that allow to present both individual and joint visualisation options for either a single or multiple selected trace files:
+\tracer\ consists of four panels that allow to present both individual and joint visualisation options for either a single or multiple selected trace files:
 
@@
 
-Tracer offers an elegant solution of visualising conditional posterior distributions.
+\tracer\ offers an elegant solution of visualising conditional posterior distributions.
 A typical use case involves support for Bayesian stochastic search variable selection (BSSVS), a form of model averaging, in which parameters appear in the likelihood function only when a specific model is indicated by an indicator function.
@@
 
-Additionally, Tracer provides demographic reconstruction resulting in a graphical plot, often applied to reconstruct epidemic dynamics.
+Additionally, \tracer\ provides demographic reconstruction resulting in a graphical plot, often applied to reconstruct epidemic dynamics.
 Depending on the actual demographic model used in the analysis, an additional file containing samples from the posterior tree distribution will be required.
@@
 \subfigure[Trace]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-traces.pdf}}
-\caption{Overview of Tracer functionality and individual parameter visualisations: (a) Main Tracer panel upon loading a single trace file; (b) boxplot representation of two continuous parameters; (c) kernel density estimates; (d) violin plots; (e) the actual traces connecting the parameter values visited by the Markov chain.}
+\caption{Overview of \tracer\ functionality and individual parameter visualisations: (a) Main \tracer\ panel upon loading a single trace file; (b) boxplot representation of two continuous parameters; (c) kernel density estimates; (d) violin plots; (e) the actual traces connecting the parameter values visited by the Markov chain.}
 \label{fig:overview}
@@
 
-The Bayesian skygrid coalescent model \citep{gill2012improving} allows us to use Tracer to reconstruct the demographic history of RABV in North America by drawing the effective population sizes over time (see Figure \ref{fig:rabv}).
-Given the properties of this coalescent prior, which requires to specify a grid of intervals over the timeframe of the epidemic, Tracer is able to infer this history without requiring a posterior sample of trees.
-According to Tracer's skygrid plot, it is clear that rabies virus has successfully established itself in North American bat species, with its numbers rising steadily throughout recent centuries.
+The Bayesian skygrid coalescent model \citep{gill2012improving} allows us to use \tracer\ to reconstruct the demographic history of RABV in North America by drawing the effective population sizes over time (see Figure \ref{fig:rabv}).
+Given the properties of this coalescent prior, which requires to specify a grid of intervals over the timeframe of the epidemic, \tracer\ is able to infer this history without requiring a posterior sample of trees.
+According to \tracer's skygrid plot, it is clear that rabies virus has successfully established itself in North American bat species, with its numbers rising steadily throughout recent centuries.
 Following a decline in numbers at the end of last century, we observe a recent sharp increase in population size of RABV in North American bats.
@@
 
-Tracer 1.7 is open-source under the GNU lesser general public license and available in both executable and source code forms at \url{http://beast.community/tracer}.
-It requires Java version 1.6 or greater.
-Github houses the Tracer's source code within \url{https://github.com/beast-dev/tracer}, and links to a GoogleGroup discussion group related to Tracer, i.e.
-the ``beast-users" group (\url{http://groups.google.com/group/beast-users}).
-This page also links to a sizeable list of self-contained, step-by-step tutorials covering basic to advance usage of Tracer to summarise posterior distributions of a large set of phylogenetic models simulated using BEAST.
-For example, popular tutorials describe how to use Tracer to generate marginal parameter summaries and infer population dynamics trajectories over time.
+\tracer\ 1.7 is open-source under the GNU lesser general public license and available in both source code (\url{https://github.com/beast-dev/tracer}) and executable (\url{http://beast.community/tracer}) forms.
+This latter page also serves up self-contained, step-by-step tutorials covering basic to advance usage of \tracer\ to summarise posteriors under a variety of phylogenetic models using BEAST.
+Popular tutorials employ \tracer\ to generate marginal parameter summaries and ito nfer population dynamics trajectories over time.
+\tracer\ requires Java version 1.6 or greater.
 

### 2017-12-15 manuscript/tracer17.tex
edits for example section

@@
 % Title must be 150 characters or less
-\title[Tracer 1.7]{Posterior summarisation in Bayesian phylogenetics using Tracer 1.7}
+\title[Tracer]{Posterior summarisation in Bayesian phylogenetics using Tracer 1.7}
 
@@
 
-\subsection*{Cross-species dynamics of rabies virus in North American bats}
+\subsection*{Cross-species dynamics of rabies in North American bats}
 
-We provide on example on reconstructing the spatial dispersal and cross-species dynamics of rabies virus (RABV) in North American bat populations based on a set of 372 nucleoprotein gene sequences (nucleotide positions: 594-1353).
-The data set comprises a total of 17 bat species sampled between 1997 and 2006 across 14 states in the United States \citep{Streicker}, including two additional species that had been excluded from the original analysis owing to a limited amount of available sequences \citep{Faria2013}.
+We use \tracer\ to infer the spatial dispersal and cross-species dynamics of rabies virus (RABV) in North American bat populations.
+The data set comprises 372 \textit{nucleoprotein} gene sequences (nucleotide positions: 594-1353) from
+17 bat species, sampled between 1997 and 2006 across 14 states in the United States \citep{Streicker,Faria2013}.
+%, including two additional species that had been excluded from the original analysis owing to a limited amount of available sequences \citep{Faria2013}.
 %We also include a viral sequence with an unknown sampling date (accession no. TX5275, sampled in Texas from Lasiurus borealis), which will be adequately accommodated in our inference.
-We estimate the ancestral locations of RABV using a Bayesian discrete phylogeographic approach and, at the same time, infer the history of host jumping using the same model approach \citep{Lemey2009}.
-We employ the Bayesian skygrid coalescent model \citep{gill2012improving} to estimate the effective population sizes over time.
+We estimate the ancestral locations of RABV and their host-jumping history using a Bayesian discrete phylogeographic approach with BSSVS \citep{Lemey2009}, while simultaneously estimating RABV effective population sizes over time through a Bayesian skygrid coalescent model \citep{gill2012improving}.
 
@@
 
-Performing BSSVS in a discrete phylogeographic approach provides us with parameters of the integer and categorical trace type.
-While a single parameter of such types can easily be visualised using a frequency plot, we have developed a TangHuLu chart to visualise the joint probability between two integer or categorical traces using coloured bubbles (see Figure \ref{fig:4tabs}:a).
-The size of circle drawn is proportional to the joint probability, %the blue coloured circle is in the credible set of given a probability threshold default to 0.95, the red is in the set not in the credible set.
-with the tile background being coloured if there is in fact a circle, to enhance visibility.
+Phylogeographic BSSVS inference includes parameters of both integer (number of non-zero transition rates) and categorical (host or location-state) trace type.
+%While a single parameter of such types can easily be visualised using a frequency plot,
+In \tracer, we have developed a TangHuLu chart to visualise the joint probability distribution between two integer or categorical traces using coloured bubbles (see Figure \ref{fig:4tabs}:a).
+Here, circle size is proportional to the joint probability, %the blue coloured circle is in the credible set of given a probability threshold default to 0.95, the red is in the set not in the credible set.
+with a coloured tile background if this probability reaches a nominal threshold to enhance visibility.
 %The coloured background can also help to show the area of the credible set and non-credible set.
-Multiple integer parameters can also be visualised using marginal density plots (see Figure \ref{fig:4tabs}:b), with a different colour scale being assigned to each parameter.
-Both figures show a clear overlap in the number of non-zero transition rates in the discrete trait models for host and location, as identified using a BSSVS procedure.
-The visualisation options for integer or categorical parameters are complemented with popular plots for continuous variables, such as a scatter or correlation plot (see Figure \ref{fig:4tabs}:c) for two continuous variables, which indicates a strong positive correlation between the clock rates of both discrete trait models.
-We also provide an extension of such a scatter plot for multiple (i.e.
+%Multiple integer parameters can also be visualised using
+Marginal density plots (see Figure \ref{fig:4tabs}:b) can also display multiple integer parameters, each with unique colour scales.
+With approximately equal numbers of transition rates, both figures suggest a similar host and location trait model complexity.
+%Both figures show a clear overlap in the number of non-zero transition rates in the discrete trait models for host and location, as identified using a BSSVS procedure.
+\tracer\ also provides
+%The visualisation options for integer or categorical parameters are complemented with
+ popular visualisations for continuous parameters, including scatter plots (see Figure \ref{fig:4tabs}:c) for two parameters,
+ and recent extensions for correlations between $\ge\hspace{-0.32em}2$ continuous parameters (Figure \ref{fig:4tabs};d; \citet{Murdoch}).
+ %, which indicates a strong positive correlation between the clock rates of both discrete trait models.
+%We also provide an extension of such a scatter plot for multiple (i.e.
 more than two) continuous variables (see Figure \ref{fig:4tabs}:d; \citet{Murdoch}), which makes use of coloured ellipses to indicate the strength of correlation between each pair of continuous variables.
-Colour gradients indicate the strength and direction of the correlation, with a colour gradient from white to dark blue indicating increasingly strong correlation and a colour gradient from white to dark red indicating increasingly strong anti-correlation.
-Additionally, the strength of correlation is reflected in the shape of the ellipse drawn, with no correlation being drawn as a circle and perfect (anti-)correlation being drawn as a line.
+Colour gradients indicate the strength and direction of the correlation, from red (strong negative-correlation) to blue (strong positive-correlation).
+%with a colour gradient from white to dark blue indicating increasingly strong correlation and a colour gradient from white to dark red indicating increasingly strong anti-correlation.
+The shape of ellipse drawn re-enforces the strength of correlation, with no correlation appearing as a circle and perfect (anti-)correlation as a line.
 %Finally, in order to display the joint marginal distribution between a continuous variable and an integer or categorical variable, we make use of box-and-whisker plots.
@@
 
-The Bayesian skygrid coalescent model \citep{gill2012improving} allows us to use \tracer\ to reconstruct the demographic history of RABV in North America by drawing the effective population sizes over time (see Figure \ref{fig:rabv}).
-Given the properties of this coalescent prior, which requires to specify a grid of intervals over the timeframe of the epidemic, \tracer\ is able to infer this history without requiring a posterior sample of trees.
-According to \tracer's skygrid plot, it is clear that rabies virus has successfully established itself in North American bat species, with its numbers rising steadily throughout recent centuries.
-Following a decline in numbers at the end of last century, we observe a recent sharp increase in population size of RABV in North American bats.
+\tracer\ reconstructs the demographic history of RABV by drawing the effective population sizes over time (Figure \ref{fig:rabv}).
+%Given the properties of this coalescent prior, which requires to specify a grid of intervals over the timeframe of the epidemic, \tracer\ is able to infer this history without requiring a posterior sample of trees.
+%According to \tracer's skygrid plot, it is clear that
+RABV has successfully established itself in North American bat species, with its effective population size rising steadily throughout recent centuries.
+Following a rapid decline at the end of last century, we observe a recent sharp increase in size, reflecting [INSERT SOMETHING].
 
-\begin{figure}[ht]
+\begin{figure}[b]
 \centerline{
@@
 
-\tracer\ 1.7 is open-source under the GNU lesser general public license and available in both source code (\url{https://github.com/beast-dev/tracer}) and executable (\url{http://beast.community/tracer}) forms.
+\tracer\ is open-source under the GNU lesser general public license and available in both source code (\url{https://github.com/beast-dev/tracer}) and executable (\url{http://beast.community/tracer}) forms.
 This latter page also serves up self-contained, step-by-step tutorials covering basic to advance usage of \tracer\ to summarise posteriors under a variety of phylogenetic models using BEAST.
-Popular tutorials employ \tracer\ to generate marginal parameter summaries and ito nfer population dynamics trajectories over time.
+Popular tutorials employ \tracer\ to generate marginal parameter summaries and to infer population dynamics trajectories over time.
 \tracer\ requires Java version 1.6 or greater.

### 2017-12-15 manuscript/tracer17.tex
heavily edited design section

@@
 $^{4}$Centre for Computational Evolution, University of Auckland, Auckland, NZ\\
-$^{5}$Department of Human Genetics, David Geffen School of Medicine, University of California, Los Angeles, USA \\
-$^{6}$Department of Biostatistics, Fielding School of Public Health, University of California, Los Angeles, USA \\
+$^{5}$Department of Human Genetics, University of California, Los Angeles, USA \\
+$^{6}$Department of Biostatistics, University of California, Los Angeles, USA \\
 %$^{7}$Department of Human Genetics, David Geffen School of Medicine at UCLA, University of California, Los Angeles, USA\\
@@
 
-\tracer\ is able to analyse a single MCMC log and can also combine posterior samples from multiple trace files stemming from an independent evaluation of the same dataset with an identical model combination.
-For each trace file loaded, \tracer\ analyses the posterior samples from all the available parameters, regardless of whether these are discrete or continuous, and present statistical summaries as well as visualisation options for all of these parameters.
-Importantly, the effective sample size (ESS) - an important measurement designed for a sampled parameter having continuous values - is immediately apparent in the default view of \tracer\ and allows users to assess the number of effectively independent draws from the posterior distribution for that parameter (see Figure \ref{fig:overview}:a).
-Colour coding is provided for the ESS values to assist the user in determining potential mixing problems in the Bayesian inference performed, with cut-off values at 100 and 200.
+A key feature of \tracer\ is its ability to link together samples from quantitative model parameter trace files with tree trace files.
+This linkage is not available in other popular post-processing MCMC software, such as coda \citep{plummer2006coda} and AWTY \citep{nylander2007awty}, and enables users to approximate the posterior distribution of scientifically relevant quantities that are functions of both the tree and other quantitative measures.
+Further, \tracer\ can analyse a single trace or combine samples from multiple traces.
+For each trace, \tracer\ examines the posterior samples from all the available parameters, regardless of whether these are continuous, integer or categorical, and presents statistical summaries and visualizations.
+Immediately apparent in the default \tracer\ view, the effective sample size (ESS) is one such statistic that
+allows users to assess the number of effectively independent draws from the posterior distribution the trace represents (Figure \ref{fig:overview}:a).
+Colour coding assists the user in determining potential MCMC mixing problems, with cut-off values at 100 and 200.
 
-\tracer\ is able to handle three trace types, also known as variable types \citep{mendenhall2012introduction}: real, integer, and categorical.
-The real trace type is used to handle quantitative continuous variables, which commonly appear in phylodynamic analyses, whereas the integer trace type allows to deal with quantitative discrete variables and the categorical trace is meant for qualitative categorical variables.
-The three trace types are automatically assigned when the trace file is imported but can also manually be set.
+%\tracer\ is able to handle three trace types, also known as variable types \citep{mendenhall2012introduction}: real, integer, and categorical.
+%The real trace type is used to handle quantitative continuous variables, which commonly appear in phylodynamic analyses, whereas the integer trace type allows to deal with quantitative discrete variables and the categorical trace is meant for qualitative categorical variables.
+%The three trace types are automatically assigned when the trace file is imported but can also manually be set.
 
-Multiple parameter traces can be selected from the traces panel, which will result in a side-by-side comparison or overlay of the selected parameters' visualisations for the different traces (see Figure \ref{fig:overview}:b-e).
+Selecting multiple parameters from the ``Traces" panel on the left generates a side-by-side comparison or an overlay of the selected parameters' visualisations (Figure \ref{fig:overview}:b-e).
 Multiple trace files can be selected in a similar fashion to compare posterior samples between different replicates of an analysis.
-If multiple trace files have the same file name,  then a ``Combined'' trace will automatically appear, which can be selected as well as the individual trace files.
-\tracer\ consists of four panels that allow to present both individual and joint visualisation options for either a single or multiple selected trace files:
+If multiple trace files have the same file name,  then a ``Combined'' trace appears automatically.
+% , which can be selected as well as the individual trace files.
+\tracer\ generates four display panels for the selected parameters:
 
@@
 
-\item Estimates: this panel reports common summary statistics such as the mean, standard deviation, highest posterior density interval, ESS, and other statistics concerning the selected parameter(s).
-A histogram presenting the frequency distribution over values will be shown for a single selected parameter of any trace type (see Figure \ref{fig:overview}:a), whereas side-by-side boxplots will be shown when multiple continuous parameters are selected (see Figure \ref{fig:overview}:b).
+\item Estimates: Reports common summary statistics such as the sample mean, standard deviation, highest posterior density interval and ESS.
+Also presents a histogram of sample values for a single selected parameter (Figure \ref{fig:overview}:a) or  side-by-side boxplots for multiple continuous parameters (Figure \ref{fig:overview}:b).
 
-\item Marginal Density: this panel shows a density plot for the posterior sample of the selected parameter(s).
-Kernel density estimation (see Figure \ref{fig:overview}:c), histograms and violin plots (see Figure \ref{fig:overview}:d) are available for continuous traces, whereas frequency plots are employed for integer traces.
+\item Marginal Density: Draws density plots for the selected parameter(s), including  kernel density estimates (Figure \ref{fig:overview}:c), histograms and violin plots (Figure \ref{fig:overview}:d) for continuous parameters and frequency plots for categorical or integer parameters.
 
-\item Joint-Marginal: visualisation in this panel only appears when two or more parameters are selected.
-The actual visualisation depends on the trace type of the selected parameters.
+\item Joint-Marginal: Visualisation in this panel appears after selecting two or more parameters and the plot form depends on the parameter types.
 We show several examples in the next section of the paper.
 
-\item Trace: this panel shows a line plot connecting the posterior samples of one or more selected parameters against state or generation number (see Figure \ref{fig:overview}:e).
-Such a trace plot is typically used to assess mixing, select a suitable burn-in and identify trends that suggest convergence issues.
+\item Trace: Constructs line plots connecting the sequential samples of one or more selected parameters against state or generation number (Figure \ref{fig:overview}:e).
+Users typically use this plot to assess mixing, select a suitable burn-in and identify trends that suggest convergence issues.
 
@@
 
-\tracer\ offers an elegant solution of visualising conditional posterior distributions.
-A typical use case involves support for Bayesian stochastic search variable selection (BSSVS), a form of model averaging, in which parameters appear in the likelihood function only when a specific model is indicated by an indicator function.
-When BSSVS is used, the posterior distribution of the parameter should not include those states that were sampled from the prior when its model was not part of the likelihood.
-Discrete phylogeographic analyses frequently employ such an approach due to the potentially large amount of transition rates that need to be estimated \citep{Lemey2009}, but this is also relevant when employing model averaging approaches such as for relaxed molecular clocks for example \citep{Li2012}.
-
-Additionally, \tracer\ provides demographic reconstruction resulting in a graphical plot, often applied to reconstruct epidemic dynamics.
-Depending on the actual demographic model used in the analysis, an additional file containing samples from the posterior tree distribution will be required.
-The currently available models are constant size, exponential and logistic growth \citep{drummond2002estimating}, Bayesian coalescent skyline \citep{drummond2005bayesian}, Gaussian Markov Random Field (GMRF) Bayesian skyride \citep{minin2008smooth}, and Bayesian coalescent skygrid \citep{gill2012improving}.
+\tracer\ offers a solution of visualising conditional posterior distributions as well.
+A typical use case involves Bayesian stochastic search variable selection (BSSVS), a form of model averaging, in which parameters influence the likelihood function only when a specific model is selected by a random indicator function.
+Under BSSVS, a posterior estimate of the parameter should only sample values from states where the indicator equals one.
+Discrete phylogeographic analyses frequently employ BSSVS due to the potentially large amount of transition rates that need to be estimated \citep{Lemey2009}, but this is also relevant when employing model averaging approaches such as for relaxed molecular clocks for example \citep{Li2012}.
+%
+Finally, \tracer\ provides demographic reconstruction resulting in a graphical plot, often applied to reconstruct epidemic dynamics.
+%Depending on the actual demographic model used in the analysis, an additional file containing samples from the posterior tree distribution will be required.
+The currently available models are constant size, exponential and logistic growth \citep{drummond2002estimating}, Bayesian  skyline \citep{drummond2005bayesian}, Bayesian skyride \citep{minin2008smooth} and Bayesian skygrid \citep{gill2012improving}.
 %Not sure if the statement below is correct ...
-All of these demographic models are available in BEAST 1 \citep{drummond2012bayesian} and BEAST 2 \citep{bouckaert2014beast2}.
+%All of these demographic models are available in BEAST 1 \citep{drummond2012bayesian} and BEAST 2 \citep{bouckaert2014beast2}.
 %Besides these, ``Lineages Through Time'' (?) is also available to plot the quantiles of the number of lineages against time, which describes the change of diversification of all organisms over time.

### 2017-12-15 manuscript/tracer17.tex
MAS done with edits this round

@@
 Users typically use this plot to assess mixing, select a suitable burn-in and identify trends that suggest convergence issues.
-
 \end{itemize}
+%
+\begin{figure}
+\subfigure[Tracer overview]{\includegraphics[width=0.47\textwidth]{./figures/rabv-overview.pdf}}
+\subfigure[Boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-boxplots.pdf}}
+\subfigure[Density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-density.pdf}}\\
+\subfigure[Violin]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-violin.pdf}}
+\subfigure[Trace]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-traces.pdf}}
+\caption{Overview of \tracer\ functionality and individual parameter visualisations: (a) Main \tracer\ panel upon loading a single trace file; (b) boxplot representation of two continuous parameters; (c) kernel density estimates; (d) violin plots; (e) the actual traces connecting the parameter values visited by the Markov chain.
+}
+\label{fig:overview}
+\end{figure}
 
@@
 
-\begin{figure}[ht]
-\subfigure[Tracer overview]{\includegraphics[width=0.47\textwidth]{./figures/rabv-overview.pdf}}
-\subfigure[Boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-boxplots.pdf}}
-\subfigure[Density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-density.pdf}}\\
-\subfigure[Violin]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-violin.pdf}}
-\subfigure[Trace]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-traces.pdf}}
-\caption{Overview of \tracer\ functionality and individual parameter visualisations: (a) Main \tracer\ panel upon loading a single trace file; (b) boxplot representation of two continuous parameters; (c) kernel density estimates; (d) violin plots; (e) the actual traces connecting the parameter values visited by the Markov chain.}
-\label{fig:overview}
-\end{figure}
-
 \section*{Example}
@@
 
-We use \tracer\ to infer the spatial dispersal and cross-species dynamics of rabies virus (RABV) in North American bat populations.
-The data set comprises 372 \textit{nucleoprotein} gene sequences (nucleotide positions: 594-1353) from
-17 bat species, sampled between 1997 and 2006 across 14 states in the United States \citep{Streicker,Faria2013}.
-%, including two additional species that had been excluded from the original analysis owing to a limited amount of available sequences \citep{Faria2013}.
-%We also include a viral sequence with an unknown sampling date (accession no. TX5275, sampled in Texas from Lasiurus borealis), which will be adequately accommodated in our inference.
-We estimate the ancestral locations of RABV and their host-jumping history using a Bayesian discrete phylogeographic approach with BSSVS \citep{Lemey2009}, while simultaneously estimating RABV effective population sizes over time through a Bayesian skygrid coalescent model \citep{gill2012improving}.
-
-\begin{figure}[ht]
+\begin{figure}
 \subfigure[TangHuLu chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integerjointmarginal.pdf}}
@@
 \end{figure}
+
+We use \tracer\ to infer the spatial dispersal and cross-species dynamics of rabies virus (RABV) in North American bat populations.
+The data set comprises 372 \textit{nucleoprotein} gene sequences (nucleotide positions: 594-1353) from
+17 bat species, sampled between 1997 and 2006 across 14 states in the United States \citep{Streicker,Faria2013}.
+%, including two additional species that had been excluded from the original analysis owing to a limited amount of available sequences \citep{Faria2013}.
+%We also include a viral sequence with an unknown sampling date (accession no. TX5275, sampled in Texas from Lasiurus borealis), which will be adequately accommodated in our inference.
+We estimate the ancestral locations of RABV and their host-jumping history using a Bayesian discrete phylogeographic approach with BSSVS \citep{Lemey2009}, while simultaneously estimating RABV effective population sizes over time through a Bayesian skygrid coalescent model \citep{gill2012improving}.
 
@@
 RABV has successfully established itself in North American bat species, with its effective population size rising steadily throughout recent centuries.
-Following a rapid decline at the end of last century, we observe a recent sharp increase in size, reflecting [INSERT SOMETHING].
+Following a rapid decline at the end of last century, we observe a recent sharp increase in size.
+%, reflecting [INSERT SOMETHING].
 
-\begin{figure}[b]
+\begin{figure}[t]
 \centerline{

### 2017-12-15 manuscript/tracer17.tex
shorter description of coalescent model list

@@
 %Depending on the actual demographic model used in the analysis, an additional file containing samples from the posterior tree distribution will be required.
-The currently available models are constant size, exponential and logistic growth \citep{drummond2002estimating}, Bayesian  skyline \citep{drummond2005bayesian}, Bayesian skyride \citep{minin2008smooth} and Bayesian skygrid \citep{gill2012improving}.
+%The currently
+Available models are constant size, exponential and logistic growth,
+%\citep{drummond2002estimating},
+%Bayesian  skyline \citep{drummond2005bayesian},
+%Bayesian skyride \citep{minin2008smooth} and
+%Bayesian skygrid \citep{gill2012improving}.
+and the non-parametric Bayesian skline, skyride and skygrid.
 %Not sure if the statement below is correct ...

### 2017-12-19 manuscript/tracer17.tex
Checks whether statistics are still being calculated.

@@
 process MCMC trace files containing parameter samples and to interactively explore the high-dimensional posterior distribution.
-\tracer\ works automatically with sample output from BEAST 1 \citep{drummond2012bayesian}, BEAST 2 \citep{bouckaert2014beast2},  LAMARC \citep{kuhner2006lamarc},  Migrate \citep{beerli2006comparison}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes} and possibly other MCMC programs from other domains.
+\tracer\ works automatically with sample output from BEAST \citep{drummond2012bayesian}, BEAST2 \citep{bouckaert2014beast2},  LAMARC \citep{kuhner2006lamarc},  Migrate \citep{beerli2006comparison}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes} and possibly other MCMC programs from other domains.
 
@@
 %Not sure if the statement below is correct ...
-%All of these demographic models are available in BEAST 1 \citep{drummond2012bayesian} and BEAST 2 \citep{bouckaert2014beast2}.
+%All of these demographic models are available in BEAST \citep{drummond2012bayesian} and BEAST2 \citep{bouckaert2014beast2}.
 %Besides these, ``Lineages Through Time'' (?) is also available to plot the quantiles of the number of lineages against time, which describes the change of diversification of all organisms over time.
@@
 \begin{figure}
-\subfigure[TangHuLu chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integerjointmarginal.pdf}}
+\subfigure[Bubble chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integerjointmarginal.pdf}}
 \subfigure[Marginal density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integermarginaldensity.pdf}} \\
@@
 %\subfigure[Box-and-whisker plot]{\includegraphics[width = .465\textwidth, height = 3cm]{./figures/rabv-catreal.pdf}}
-\caption{Joint marginal and marginal density plots of: (a) two integer variables through a TangHuLu chart; (b) frequency plots the marginal density of multiple integer or categorical variables; (c) classic scatter or correlation plot for two continuous variables; (d) large correlation matrices for multiple ($> 2$) continuous variables.
+\caption{Joint marginal and marginal density plots of: (a) two integer variables through a bubble chart; (b) frequency plots the marginal density of multiple integer or categorical variables; (c) classic scatter or correlation plot for two continuous variables; (d) large correlation matrices for multiple ($> 2$) continuous variables.
 %; (e) box-and-whisker plots for the joint marginal distribution of a continuous and an integer or categorical variable.
@@
 %While a single parameter of such types can easily be visualised using a frequency plot,
-In \tracer, we have developed a TangHuLu chart to visualise the joint probability distribution between two integer or categorical traces using coloured bubbles (see Figure \ref{fig:4tabs}:a).
+In \tracer, we have developed a bubble chart to visualise the joint probability distribution between two integer or categorical traces using coloured bubbles (see Figure \ref{fig:4tabs}:a).
 Here, circle size is proportional to the joint probability, %the blue coloured circle is in the credible set of given a probability threshold default to 0.95, the red is in the set not in the credible set.

### 2017-12-19 src/tracer/traces/IntervalsPanel.java

@@ -138,6 +138,11 @@ public class IntervalsPanel extends TraceChartPanel {
                     setMessage("Traces must be of the same type to visualize here.");
                     return;
                 }
+
+                if (trace.getTraceStatistics() == null) {
+                    setMessage("Trace statistics are still being calculated.");
+                    return;
+                }
             }
         }
 

### 2017-12-20 manuscript/tracer17.tex
Minor tweak

@@
 %While a single parameter of such types can easily be visualised using a frequency plot,
-In \tracer, we have developed a bubble chart to visualise the joint probability distribution between two integer or categorical traces using coloured bubbles (see Figure \ref{fig:4tabs}:a).
+In \tracer, we use a bubble chart to visualise the joint probability distribution between two integer or categorical traces (see Figure \ref{fig:4tabs}:a).
 Here, circle size is proportional to the joint probability, %the blue coloured circle is in the credible set of given a probability threshold default to 0.95, the red is in the set not in the credible set.

### 2018-01-23 manuscript/tracer17.tex
move figures to make room

@@
-\documentclass{bioinfo}
+\documentclass{bioinfo_tracer}
 \copyrightyear{2017}
@@
 We show several examples in the next section of the paper.
-
-\item Trace: Constructs line plots connecting the sequential samples of one or more selected parameters against state or generation number (Figure \ref{fig:overview}:e).
-Users typically use this plot to assess mixing, select a suitable burn-in and identify trends that suggest convergence issues.
-\end{itemize}
 %
-\begin{figure}
+\begin{figure}[t]
 \subfigure[Tracer overview]{\includegraphics[width=0.47\textwidth]{./figures/rabv-overview.pdf}}
@@
 \subfigure[Trace]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-traces.pdf}}
-\caption{Overview of \tracer\ functionality and individual parameter visualisations: (a) Main \tracer\ panel upon loading a single trace file; (b) boxplot representation of two continuous parameters; (c) kernel density estimates; (d) violin plots; (e) the actual traces connecting the parameter values visited by the Markov chain.
-}
+\caption{Overview of \tracer\ functionality and individual parameter visualisations: (a) Main \tracer\ panel upon loading a single trace file; (b) boxplot representation of two continuous parameters; (c) kernel density estimates; (d) violin plots; (e) the actual traces connecting the parameter values visited by the Markov chain.}
 \label{fig:overview}
 \end{figure}
+%
+\item Trace: Constructs line plots connecting the sequential samples of one or more selected parameters against state or generation number (Figure \ref{fig:overview}:e).
+Users typically use this plot to assess mixing, select a suitable burn-in and identify trends that suggest convergence issues.
+\end{itemize}
+%
 
@@
 
-\begin{figure}
+We use \tracer\ to infer the spatial dispersal and cross-species dynamics of rabies virus (RABV) in North American bat populations.
+%
+\begin{figure}[t]
 \subfigure[Bubble chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integerjointmarginal.pdf}}
@@
 \end{figure}
-
-We use \tracer\ to infer the spatial dispersal and cross-species dynamics of rabies virus (RABV) in North American bat populations.
+%
 The data set comprises 372 \textit{nucleoprotein} gene sequences (nucleotide positions: 594-1353) from

### 2018-01-24 manuscript/tracer17.tex
tiny edit

@@
  popular visualisations for continuous parameters, including scatter plots (see Figure \ref{fig:4tabs}:c) for two parameters,
- and recent extensions for correlations between $\ge\hspace{-0.32em}2$ continuous parameters (Figure \ref{fig:4tabs};d; \citet{Murdoch}).
+ and recent extensions for correlations between $\ge\hspace{-0.32em}2$ continuous parameters (Figure \ref{fig:4tabs}:d; \citet{Murdoch}).
  %, which indicates a strong positive correlation between the clock rates of both discrete trait models.

### 2018-01-28 manuscript/tracer17.tex

@@
 \tracer\ is open-source under the GNU lesser general public license and available in both source code (\url{https://github.com/beast-dev/tracer}) and executable (\url{http://beast.community/tracer}) forms.
-This latter page also serves up self-contained, step-by-step tutorials covering basic to advance usage of \tracer\ to summarise posteriors under a variety of phylogenetic models using BEAST.
+This latter page also serves up self-contained, step-by-step tutorials covering basic to advanced usage of \tracer\ to summarise posteriors under a variety of phylogenetic models using BEAST.
 Popular tutorials employ \tracer\ to generate marginal parameter summaries and to infer population dynamics trajectories over time.

### 2018-01-28 src/tracer/analysis/TemporalAnalysisFrame.java
Added menu commands ('Export Data Table' and 'Export Raw Data'). The idea is the first writes a table representing the visualization being displayed, the latter writes the underlying traces selected.

@@ -106,6 +106,11 @@ public class TemporalAnalysisFrame extends AuxilaryFrame implements TracerFileMe
 
         temporalAnalysisPlotPanel.addDemographicPlot(title, xData, yDataMean, yDataMedian, yDataUpper, yDataLower,
                 timeMean, timeMedian, timeUpper, timeLower);
+
+        // unless there was a meaningful data table to write
+        exportRawDataAction.setEnabled(false);
+        exportDataTableAction.setEnabled(true);
+
         setVisible(true);
     }
 
@@ -291,22 +296,30 @@ public class TemporalAnalysisFrame extends AuxilaryFrame implements TracerFileMe
         return null;
     }
 
-    public Action getExportDataAction() {
-        return exportDataAction;
-    }
-
-    public Action getExportPDFAction() {
-        return exportPDFAction;
+    @Override
+    public Action getExportRawDataAction() {
+        return exportRawDataAction;
     }
 
     @Override
-    public Action getExportFullStatisticsAction() {
-        return null;
+    public Action getExportDataTableAction() {
+        return exportDataTableAction;
     }
 
-    private AbstractAction exportDataAction = new AbstractAction("Export Data...") {
+    @Override
+    public Action getExportPDFAction() {
+        return exportPDFAction;
+    }
+    
+    private AbstractAction exportDataTableAction = new AbstractAction("Export Data Table...") {
         public void actionPerformed(ActionEvent ae) {
             doExportData();
+        }
+    };
+
+    private AbstractAction exportRawDataAction = new AbstractAction("Export Raw Data...") {
+        public void actionPerformed(ActionEvent ae) {
+            // do nothing
         }
     };
 

### 2018-01-28 src/tracer/application/TracerDefaultFileMenuFactory.java

@@ -79,13 +79,13 @@ public class TracerDefaultFileMenuFactory implements MenuFactory {
 
             menu.addSeparator();
 
-            item = new JMenuItem(((TracerFileMenuHandler)frame).getExportDataAction());
+            item = new JMenuItem(((TracerFileMenuHandler)frame).getExportDataTableAction());
             item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK));
             menu.add(item);
 
-//            item = new JMenuItem(((TracerFileMenuHandler)frame).getExportFullStatisticsAction());
-//            item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK + ActionEvent.SHIFT_MASK));
-//            menu.add(item);
+            item = new JMenuItem(((TracerFileMenuHandler)frame).getExportRawDataAction());
+            item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK + KeyEvent.SHIFT_MASK));
+            menu.add(item);
 
             item = new JMenuItem(((TracerFileMenuHandler)frame).getExportPDFAction());
             item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK + KeyEvent.ALT_MASK));

### 2018-01-28 src/tracer/application/TracerFileMenuHandler.java

@@ -36,9 +36,11 @@ public interface TracerFileMenuHandler {
 
 	Action getReloadAction();
 
-	Action getExportDataAction();
+	Action getExportRawDataAction();
+
+	Action getExportDataTableAction();
 
 	Action getExportPDFAction();
 
-	Action getExportFullStatisticsAction();
+//	Action getExportFullStatisticsAction();
 }

### 2018-01-28 src/tracer/application/TracerFrame.java

@@ -139,7 +139,7 @@ public class TracerFrame extends DocumentFrame implements TracerFileMenuHandler,
             }
         };
         setImportAction(importAction);
-        setExportAction(exportDataAction);
+        setExportAction(exportDataTableAction);
 
         setAnalysesEnabled(false);
     }
@@ -409,8 +409,9 @@ public class TracerFrame extends DocumentFrame implements TracerFileMenuHandler,
         getConditionalPosteriorDistAction().setEnabled(enabled);
 
         getExportAction().setEnabled(enabled);
-        getExportDataAction().setEnabled(enabled);
-        getExportFullStatisticsAction().setEnabled(enabled);
+        getExportRawDataAction().setEnabled(enabled);
+        getExportDataTableAction().setEnabled(enabled);
+//        getExportFullStatisticsAction().setEnabled(enabled);
         getExportPDFAction().setEnabled(enabled);
         getCopyAction().setEnabled(true);
     }
@@ -816,15 +817,7 @@ public class TracerFrame extends DocumentFrame implements TracerFileMenuHandler,
                 isIncomplete = true;
         }
 
-        java.util.List<String> selectedTraces = new ArrayList<String>();
-        for (int selRow : selRows) {
-            if (selRow < commonTraceNames.size()) {
-                selectedTraces.add(commonTraceNames.get(selRow));
-            }
-        }
-        if (selectedTraces.size() < 1) {
-            selectedTraces.add(commonTraceNames.get(0));
-        }
+        List<String> selectedTraces = getSelectedTraces();
 
         if (currentTraceLists.size() == 0 || isIncomplete) {
             tracePanel.setTraces(null, selectedTraces);
@@ -843,6 +836,23 @@ public class TracerFrame extends DocumentFrame implements TracerFileMenuHandler,
         realButton.setEnabled(selRows.length > 0);
         integerButton.setEnabled(selRows.length > 0);
         categoricalButton.setEnabled(selRows.length > 0);
+    }
+
+    public List<String> getSelectedTraces() {
+
+        int[] selRows = statisticTable.getSelectedRows();
+
+        List<String> selectedTraces = new ArrayList<String>();
+        for (int selRow : selRows) {
+            if (selRow < commonTraceNames.size()) {
+                selectedTraces.add(commonTraceNames.get(selRow));
+            }
+        }
+        if (selectedTraces.size() < 1) {
+            selectedTraces.add(commonTraceNames.get(0));
+        }
+
[…]

### 2018-01-28 src/tracer/application/TracerMacFileMenuFactory.java

@@ -87,13 +87,13 @@ public class TracerMacFileMenuFactory implements MenuFactory {
 
             menu.addSeparator();
 
-            item = new JMenuItem(((TracerFileMenuHandler)frame).getExportDataAction());
+            item = new JMenuItem(((TracerFileMenuHandler)frame).getExportDataTableAction());
             item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK));
             menu.add(item);
 
-//            item = new JMenuItem(((TracerFileMenuHandler)frame).getExportFullStatisticsAction());
-//            item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK + ActionEvent.SHIFT_MASK));
-//            menu.add(item);
+            item = new JMenuItem(((TracerFileMenuHandler)frame).getExportRawDataAction());
+            item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK + KeyEvent.SHIFT_MASK));
+            menu.add(item);
 
             item = new JMenuItem(((TracerFileMenuHandler)frame).getExportPDFAction());
             item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK + KeyEvent.ALT_MASK));
@@ -115,13 +115,13 @@ public class TracerMacFileMenuFactory implements MenuFactory {
 
             menu.addSeparator();
 
-            item = new JMenuItem("Export Data...");
+            item = new JMenuItem("Export Raw Data...");
             item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK));
             item.setEnabled(false);
             menu.add(item);
 
-            item = new JMenuItem("Export Statistic Summary...");
-            item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK + ActionEvent.SHIFT_MASK));
+            item = new JMenuItem("Export Data Table...");
+            item.setAccelerator(KeyStroke.getKeyStroke(KeyEvent.VK_E, MenuBarFactory.MENU_MASK + KeyEvent.SHIFT_MASK));
             item.setEnabled(false);
             menu.add(item);
 

### 2018-01-28 src/tracer/traces/ContinuousDensityPanel.java

@@ -364,6 +364,7 @@ public class ContinuousDensityPanel extends TraceChartPanel {
             getChartPanel().setYAxisTitle("Multiple Traces");
         }
     }
+    
     public String toString() {
         JChart chart = getChartPanel().getChart();
 

### 2018-01-28 src/tracer/traces/GridJointDensityPanel.java

@@ -238,7 +238,10 @@ public class GridJointDensityPanel extends TraceChartPanel {
 
     }
 
+    // Called by exportDataTable
     public String toString() {
+        // todo - make this a table of correlation coefficients
+
         if (getChart().getPlotCount() == 0) {
             return "no plot available";
         }

### 2018-01-28 src/tracer/traces/TracePanel.java

@@ -48,8 +48,6 @@ public class TracePanel extends javax.swing.JPanel implements Exportable {
     private final JointDensityPanel jointDensityPanel;
     private final RawTracePanel tracePanel;
 
-    private static final boolean USE_KDE = false;
-
     /**
      * Creates new form TracePanel
      */
@@ -96,14 +94,15 @@ public class TracePanel extends javax.swing.JPanel implements Exportable {
         java.awt.datatransfer.Clipboard clipboard =
                 Toolkit.getDefaultToolkit().getSystemClipboard();
 
+        // default for copy is the data table?
         java.awt.datatransfer.StringSelection selection =
-                new java.awt.datatransfer.StringSelection(getExportText());
+                new java.awt.datatransfer.StringSelection(getExportDataTableText());
 
         clipboard.setContents(selection, selection);
 
     }
 
-    public String getExportText() {
+    public String getExportDataTableText() {
         switch (tabbedPane.getSelectedIndex()) {
             case 0:
                 return summaryPanel.toString();
@@ -116,7 +115,6 @@ public class TracePanel extends javax.swing.JPanel implements Exportable {
         }
         return "";
     }
-
     public JComponent getExportableComponent() {
 
         JComponent exportable = null;

### 2018-01-30 manuscript/tracer17.tex
update figures

@@
 \begin{figure}[t]
-\subfigure[Tracer overview]{\includegraphics[width=0.47\textwidth]{./figures/rabv-overview.pdf}}
+\subfigure[Tracer overview]{\includegraphics[width=0.47\textwidth]{./figures/rabv-overview-2.pdf}}
 \subfigure[Boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-boxplots.pdf}}
@@
 \begin{figure}[t]
-\subfigure[Bubble chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integerjointmarginal.pdf}}
+\subfigure[Bubble chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integerjointmarginal-2.pdf}}
 \subfigure[Marginal density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integermarginaldensity.pdf}} \\

### 2018-01-31 src/tracer/traces/ContinuousDensityPanel.java
solve Colours for trace file, trace file and trace not working #129

@@ -27,7 +27,6 @@ package tracer.traces;
 
 import dr.app.gui.chart.*;
 import dr.inference.trace.Trace;
-import dr.inference.trace.TraceCorrelation;
 import dr.inference.trace.TraceList;
 import dr.inference.trace.TraceType;
 import dr.stats.Variate;
@@ -306,7 +305,7 @@ public class ContinuousDensityPanel extends TraceChartPanel {
 
                     if (plot != null) {
                         plot.setName(name);
-                        int selectedColour = currentSettings.cm.addTraceColour(tl.getName(), name, currentSettings.colourBy);
+                        int selectedColour = currentSettings.cm.addTraceColour(tl.getFullName(), name, currentSettings.colourBy);
                         if (tl instanceof CombinedTraces) {
                             plot.setLineStyle(new BasicStroke(2.0f), currentSettings.palette[selectedColour]);
                         } else {

### 2018-01-31 src/tracer/traces/DiscreteDensityPanel.java

@@ -31,8 +31,8 @@ import dr.stats.Variate;
 
 import javax.swing.*;
 import java.awt.*;
-import java.util.*;
-import java.util.List;
+import java.util.HashSet;
+import java.util.Set;
 
 /**
  * A panel that displays density plots of traces
@@ -192,7 +192,7 @@ public class DiscreteDensityPanel extends TraceChartPanel {
 
                 if (columnPlot != null) {
                     columnPlot.setName(name);
-                    int selectedColour = currentSettings.cm.addTraceColour(tl.getName(), name, currentSettings.colourBy);
+                    int selectedColour = currentSettings.cm.addTraceColour(tl.getFullName(), name, currentSettings.colourBy);
                     if (tl instanceof CombinedTraces) {
                         columnPlot.setLineStyle(new BasicStroke(2.0f), currentSettings.palette[selectedColour]);
                     } else {

### 2018-01-31 src/tracer/traces/RawTracePanel.java

@@ -275,13 +275,13 @@ public class RawTracePanel extends TraceChartPanel {
                             getChart().getYAxis().setRange(0.0, 1.0);
                         }
                     }
-                    int selectedColour = currentSettings.cm.addTraceColour(tl.getName(), name, currentSettings.colourBy);
+                    int selectedColour = currentSettings.cm.addTraceColour(tl.getFullName(), name, currentSettings.colourBy);
                     //System.out.println(tl.getName() + " ; " + name + " : " + selectedColour);
                     minMax = getChart().addTrace(name, stateStart, stateStep, values, burninValues, currentSettings.palette[selectedColour]);
                 } else if (trace.getTraceType() == TraceType.CATEGORICAL) {
                     //change Y axis to discrete
                     getChart().setYAxis(new DiscreteAxis(trace.getCategoryLabelMap(), true, true));
-                    int selectedColour = currentSettings.cm.addTraceColour(tl.getName(), name, currentSettings.colourBy);
+                    int selectedColour = currentSettings.cm.addTraceColour(tl.getFullName(), name, currentSettings.colourBy);
                     //System.out.println(tl.getName() + " ; " + name + " : " + selectedColour);
                     minMax = getChart().addTrace(name, stateStart, stateStep, values, burninValues, currentSettings.palette[selectedColour]);
                 } else {

### 2018-01-31 manuscript/tracer17.tex
revert to original author-order

@@
 Andrew Rambaut\,$^{1}$,
-Guy Baele$^{2}$,
-Dong Xie\,$^{3,4}$,
-Alexei J.~Drummond\,$^{3,4}$,
+Alexei J.~Drummond\,$^{2,3}$,
+Guy Baele$^{4}$,
+Dong Xie\,$^{2,3}$,
 Marc A.~Suchard\,$^{5,6}$
@@
 \address{
-$^{1}$Institute of Evolutionary Biology, University of Edinburgh, Edinburgh, UK\\
-$^{2}$Department of Microbiology and Immunology, Rega Institute, KU Leuven, Leuven, Belgium\\
-$^{3}$Department of Computer Science, University of Auckland, Auckland, NZ\\
-$^{4}$Centre for Computational Evolution, University of Auckland, Auckland, NZ\\
+$^{1}$Institute of Evolutionary Biology, University of Edinburgh, Edinburgh, UK \\
+$^{2}$Department of Computer Science, University of Auckland, Auckland, NZ \\
+$^{3}$Centre for Computational Evolution, University of Auckland, Auckland, NZ \\
+$^{4}$Department of Microbiology and Immunology, Rega Institute, KU Leuven, Leuven, Belgium\\
 $^{5}$Department of Human Genetics, University of California, Los Angeles, USA \\

### 2018-02-01 manuscript/tracer17.tex
small changes

@@
 Alexei J.~Drummond\,$^{2,3}$,
+Dong Xie\,$^{2,3}$,
 Guy Baele$^{4}$,
-Dong Xie\,$^{2,3}$,
 Marc A.~Suchard\,$^{5,6}$
@@
 \section{Availability:}
-\tracer\ is open-source and available at \url{http://http://beast.community/tracer}.
+\tracer\ is open-source and available at \url{http://beast.community/tracer}.
 
@@
 
-Bayesian inference of phylogeny using Markov chain Monte Carlo (MCMC) \citep{rannala1996probability, mau1999bayesian} flourishes as a popular approach to uncover the evolutionary relationships among taxa, such as genes, genomes, individuals or species.
+Bayesian inference of phylogeny using Markov chain Monte Carlo (MCMC) \citep{rannala1996probability, mau1999bayesian, drummond2002estimating} flourishes as a popular approach to uncover the evolutionary relationships among taxa, such as genes, genomes, individuals or species.
 MCMC approaches generate samples of model parameter values - including the phylogenetic tree - drawn from their posterior distribution given molecular sequence data and a selection of evolutionary models.
@@
 %The currently
-Available models are constant size, exponential and logistic growth,
-%\citep{drummond2002estimating},
-%Bayesian  skyline \citep{drummond2005bayesian},
-%Bayesian skyride \citep{minin2008smooth} and
-%Bayesian skygrid \citep{gill2012improving}.
-and the non-parametric Bayesian skline, skyride and skygrid.
+Available models are constant size, exponential and logistic growth \citep{drummond2002estimating},
+and the non-parametric Bayesian skyline \citep{drummond2005bayesian,heledDrummond2008}, skyride \citep{minin2008smooth} and skygrid \citep{gill2012improving}.
 %Not sure if the statement below is correct ...
@@
 In \tracer, we use a bubble chart to visualise the joint probability distribution between two integer or categorical traces (see Figure \ref{fig:4tabs}:a).
-Here, circle size is proportional to the joint probability, %the blue coloured circle is in the credible set of given a probability threshold default to 0.95, the red is in the set not in the credible set.
+Here, circle area is proportional to the joint probability, %the blue coloured circle is in the credible set of given a probability threshold default to 0.95, the red is in the set not in the credible set.
 with a coloured tile background if this probability reaches a nominal threshold to enhance visibility.
@@
 This work was supported in part by the Wellcome Trust through project 206298/Z/17/Z (Artic Network), the European Union Seventh Framework Programme under grant agreement no.
-725422-RESERVOIRDOCS, the National Science Foundation through grant DMS 1264153, the National Institutes of Health under grants R01 AI107034 and U19 AI135995, and the Marsden Trust.
+725422-RESERVOIRDOCS, the National Science Foundation through grant DMS 1264153, the National Institutes of Health under grants R01 AI107034 and U19 AI135995.
+AJD acknowledges support from Marsden grant contract UOA1611.
 GB acknowledges support from the Interne Fondsen KU Leuven / Internal Funds KU Leuven.

### 2018-02-02 manuscript/tracer17.tex
Removed the bit about the linking of trees and parameters as it didn't really make sense. Probably is worth adding a paragraph about other software for post-mcmc analysis (i.e., AWTY) but say how these are complementary and look at different things.

@@
 
-A key feature of \tracer\ is its ability to link together samples from quantitative model parameter trace files with tree trace files.
-This linkage is not available in other popular post-processing MCMC software, such as coda \citep{plummer2006coda} and AWTY \citep{nylander2007awty}, and enables users to approximate the posterior distribution of scientifically relevant quantities that are functions of both the tree and other quantitative measures.
-Further, \tracer\ can analyse a single trace or combine samples from multiple traces.
-For each trace, \tracer\ examines the posterior samples from all the available parameters, regardless of whether these are continuous, integer or categorical, and presents statistical summaries and visualizations.
+%A key feature of \tracer\ is its ability to link together samples from quantitative model parameter trace files with tree trace files. %% AR - only for the demographic reconstructions...
+%This linkage is not available in other popular post-processing MCMC software, such as coda \citep{plummer2006coda} and AWTY \citep{nylander2007awty}, and enables users to approximate the posterior distribution of scientifically relevant quantities that are functions of both the tree and other quantitative measures.
+%% AR - ...
+which are intended in these other software packages.
+For each trace, \tracer\ examines the posterior samples from all the available parameters, treating continuous, integer and categorical parameters appropriately, and presents statistical summaries and visualizations.
+Further, \tracer\ can analyse a single trace or combine samples from multiple files.
 Immediately apparent in the default \tracer\ view, the effective sample size (ESS) is one such statistic that

### 2018-02-02 manuscript/tracer17.tex
Added paragraph about other packages. Probably needs work.

@@
 
+Other packages are available for the post-processing of MCMC samples are available and are largely complementary to \tracer\.
+The R package `coda' \citep{plummer2006coda} provides some of the functionality of \tracer\ within the R programming environment.
+`AWTY' \citep{nylander2007awty} looks at the convergence of phylogenetic MCMC runs for the phylogenetic tree itself.
+%% AR - anything else?
+The MDS tree-space approaches of Hillis etc.?
+
 \section*{Availability}

### 2018-02-02 manuscript/tracer17.tex
edited new AR paragraph; please reformat to fit back on 3 pages, pretty please.

--- 2018-02-02 manuscript/tracer17.tex
@@
 
+Other packages are available for the post-processing of MCMC samples and are largely complementary to \tracer\.
+The R package `coda' \citep{plummer2006coda} provides some of the functionality of \tracer\ within the R programming environment.
+`AWTY' \citep{nylander2007awty} explores the convergence of the phylogenetic tree parameter itself across multiple MCMC runs.
+%% AR - anything else? The MDS tree-space approaches of Hillis etc.?
+
 \section*{Availability}

### 2018-02-02 manuscript/tracer17.tex
towards 3 pages - almost there

@@
 We show several examples in the next section of the paper.
+
+\item Trace: Constructs line plots connecting the sequential samples of one or more selected parameters against state or generation number (Figure \ref{fig:overview}:e).
+Users typically use this plot to assess mixing, select a suitable burn-in and identify trends that suggest convergence issues.
+\end{itemize}
+%
+
 %
@@
 \subfigure[Trace]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-traces.pdf}}
+\vspace{-0.25cm}
 \caption{Overview of \tracer\ functionality and individual parameter visualisations: (a) Main \tracer\ panel upon loading a single trace file; (b) boxplot representation of two continuous parameters; (c) kernel density estimates; (d) violin plots; (e) the actual traces connecting the parameter values visited by the Markov chain.}
@@
 %
-\item Trace: Constructs line plots connecting the sequential samples of one or more selected parameters against state or generation number (Figure \ref{fig:overview}:e).
-Users typically use this plot to assess mixing, select a suitable burn-in and identify trends that suggest convergence issues.
-\end{itemize}
+
 %
-
+\begin{figure}[t]
+\subfigure[Bubble chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integerjointmarginal-2.pdf}}
+\subfigure[Marginal density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integermarginaldensity.pdf}} \\
+\subfigure[Scatter plot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-correlation.pdf}}
+\subfigure[Correlation plot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-joint-real.pdf}}
+\vspace{-0.25cm}
+%\subfigure[Box-and-whisker plot]{\includegraphics[width = .465\textwidth, height = 3cm]{./figures/rabv-catreal.pdf}}
+\caption{Joint marginal and marginal density plots of: (a) two integer variables through a bubble chart; (b) frequency plots the marginal density of multiple integer or categorical variables; (c) classic scatter or correlation plot for two continuous variables; (d) large correlation matrices for multiple ($> 2$) continuous variables.
+%; (e) box-and-whisker plots for the joint marginal distribution of a continuous and an integer or categorical variable.
+}
+\label{fig:4tabs}
+\end{figure}
+%
 \tracer\ offers a solution of visualising conditional posterior distributions as well.
@@
 
+\vspace{-0.5cm}
+
 \section*{Example}
@@
 We use \tracer\ to infer the spatial dispersal and cross-species dynamics of rabies virus (RABV) in North American bat populations.
-%
-\begin{figure}[t]
-\subfigure[Bubble chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integerjointmarginal-2.pdf}}
-\subfigure[Marginal density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integermarginaldensity.pdf}} \\
-\subfigure[Scatter plot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-correlation.pdf}}
-\subfigure[Correlation plot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-joint-real.pdf}} \\
-%\subfigure[Box-and-whisker plot]{\includegraphics[width = .465\textwidth, height = 3cm]{./figures/rabv-catreal.pdf}}
-\caption{Joint marginal and marginal density plots of: (a) two integer variables through a bubble chart; (b) frequency plots the marginal density of multiple integer or categorical variables; (c) classic scatter or correlation plot for two continuous variables; (d) large correlation matrices for multiple ($> 2$) continuous variables.
-%; (e) box-and-whisker plots for the joint marginal distribution of a continuous and an integer or categorical variable.
-}
-\label{fig:4tabs}
-\end{figure}
-%
-The data set comprises 372 \textit{nucleoprotein} gene sequences (nucleotide positions: 594-1353) from
-17 bat species, sampled between 1997 and 2006 across 14 states in the United States \citep{Streicker,Faria2013}.
+The data set comprises 372 \textit{nucleoprotein} gene sequences from 17 bat species, sampled between 1997 and 2006 across 14 states in the United States \citep{Streicker,Faria2013}.
 %, including two additional species that had been excluded from the original analysis owing to a limited amount of available sequences \citep{Faria2013}.
@@
 }
+\vspace{-0.25cm}
 \caption{Estimating the effective population sizes over time using a Bayesian skygrid demographic reconstruction for rabies virus in North America.}
@@
 
-Other packages are available for the post-processing of MCMC samples and are largely complementary to \tracer\.
+Other packages are available for the post-processing of MCMC samples and are largely complementary to \tracer\ .
 The R package `coda' \citep{plummer2006coda} provides some of the functionality of \tracer\ within the R programming environment.
@@
 %% AR - anything else? The MDS tree-space approaches of Hillis etc.?
+
+\vspace{-0.25cm}
 
@@
 
+\vspace{-0.25cm}
+
 \section*{Acknowledgements}
@@
 
+\vspace{-0.25cm}
+
 %\section*{References}

### 2018-02-02 manuscript/tracer17.tex
saved some more space
minor edits

@@
 \begin{figure}[t]
-\subfigure[Tracer overview]{\includegraphics[width=0.47\textwidth]{./figures/rabv-overview-2.pdf}}
+\subfigure[Tracer overview]{\hspace{0.40cm}\includegraphics[width=0.432\textwidth]{./figures/rabv-overview-3.pdf}}
 \subfigure[Boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-boxplots.pdf}}
@@
 %\subfigure[Box-and-whisker plot]{\includegraphics[width = .465\textwidth, height = 3cm]{./figures/rabv-catreal.pdf}}
-\caption{Joint marginal and marginal density plots of: (a) two integer variables through a bubble chart; (b) frequency plots the marginal density of multiple integer or categorical variables; (c) classic scatter or correlation plot for two continuous variables; (d) large correlation matrices for multiple ($> 2$) continuous variables.
+\caption{Multi-parameter visualisations of: (a) the joint probability distribution of two integer variables through a bubble chart; (b) the marginal density of multiple integer or categorical variables through frequency plots; (c) two continuous variables through a classic scatter or correlation plot; (d) multiple ($> 2$) continuous variables using large correlation matrices.
 %; (e) box-and-whisker plots for the joint marginal distribution of a continuous and an integer or categorical variable.
@@
 %
+
 \tracer\ offers a solution of visualising conditional posterior distributions as well.
@@
 
-We use \tracer\ to infer the spatial dispersal and cross-species dynamics of rabies virus (RABV) in North American bat populations.
+We use \tracer\ to infer the spatial dispersal and cross-species dynamics of rabies virus (RABV) in North American bats.
 The data set comprises 372 \textit{nucleoprotein} gene sequences from 17 bat species, sampled between 1997 and 2006 across 14 states in the United States \citep{Streicker,Faria2013}.
-%, including two additional species that had been excluded from the original analysis owing to a limited amount of available sequences \citep{Faria2013}.
-%We also include a viral sequence with an unknown sampling date (accession no. TX5275, sampled in Texas from Lasiurus borealis), which will be adequately accommodated in our inference.
-We estimate the ancestral locations of RABV and their host-jumping history using a Bayesian discrete phylogeographic approach with BSSVS \citep{Lemey2009}, while simultaneously estimating RABV effective population sizes over time through a Bayesian skygrid coalescent model \citep{gill2012improving}.
+%We estimate the ancestral locations of RABV and their host-jumping history using a Bayesian discrete phylogeographic approach with BSSVS \citep{Lemey2009}, while simultaneously estimating RABV effective population sizes over time through a Bayesian skygrid coalescent model \citep{gill2012improving}.
+We estimate RABV ancestral locations and host-jumping history using a Bayesian discrete phylogeographic approach with BSSVS, while simultaneously estimating effective population sizes over time through a Bayesian skygrid coalescent model \citep{gill2012improving}.
 
-Phylogeographic BSSVS inference includes parameters of both integer (number of non-zero transition rates) and categorical (host or location-state) trace type.
+Phylogeographic BSSVS inference includes parameters of both integer (number of non-zero transition rates) and categorical (host or location-state) trace types.
 %While a single parameter of such types can easily be visualised using a frequency plot,
-In \tracer, we use a bubble chart to visualise the joint probability distribution between two integer or categorical traces (see Figure \ref{fig:4tabs}:a).
-Here, circle area is proportional to the joint probability, %the blue coloured circle is in the credible set of given a probability threshold default to 0.95, the red is in the set not in the credible set.
+In \tracer, a bubble chart visualises the joint probability distribution between two integer or categorical traces (see Figure \ref{fig:4tabs}:a).
+Circle area is proportional to the joint probability, %the blue coloured circle is in the credible set of given a probability threshold default to 0.95, the red is in the set not in the credible set.
 with a coloured tile background if this probability reaches a nominal threshold to enhance visibility.
@@
 %Multiple integer parameters can also be visualised using
-Marginal density plots (see Figure \ref{fig:4tabs}:b) can also display multiple integer parameters, each with unique colour scales.
-With approximately equal numbers of transition rates, both figures suggest a similar host and location trait model complexity.
+Marginal density plots can also display multiple integer parameters, each with unique colour scales (see Figure \ref{fig:4tabs}:b).
+With approximately equal numbers of transition rates, both figures suggest similar host and location trait model complexity.
 %Both figures show a clear overlap in the number of non-zero transition rates in the discrete trait models for host and location, as identified using a BSSVS procedure.
@@
 %The visualisation options for integer or categorical parameters are complemented with
- popular visualisations for continuous parameters, including scatter plots (see Figure \ref{fig:4tabs}:c) for two parameters,
- and recent extensions for correlations between $\ge\hspace{-0.32em}2$ continuous parameters (Figure \ref{fig:4tabs}:d; \citet{Murdoch}).
+ popular visualisations for continuous parameters, including scatter plots for two parameters (see Figure \ref{fig:4tabs}:c),
+ and extensions for correlations between $\ge\hspace{-0.32em}2$ continuous parameters (Figure \ref{fig:4tabs}:d; \citet{Murdoch}).
  %, which indicates a strong positive correlation between the clock rates of both discrete trait models.
@@
 more than two) continuous variables (see Figure \ref{fig:4tabs}:d; \citet{Murdoch}), which makes use of coloured ellipses to indicate the strength of correlation between each pair of continuous variables.
-Colour gradients indicate the strength and direction of the correlation, from red (strong negative-correlation) to blue (strong positive-correlation).
+Colour gradients indicate strength and direction of the correlation, from red (strong negative) to blue (strong positive).
 %with a colour gradient from white to dark blue indicating increasingly strong correlation and a colour gradient from white to dark red indicating increasingly strong anti-correlation.
-The shape of ellipse drawn re-enforces the strength of correlation, with no correlation appearing as a circle and perfect (anti-)correlation as a line.
+Ellipse shapes re-enforce the strength of correlation, with no correlation appearing as a circle and perfect (anti-)correlation as a line.
 %Finally, in order to display the joint marginal distribution between a continuous variable and an integer or categorical variable, we make use of box-and-whisker plots.

### 2018-02-02 manuscript/tracer17.tex
last edits to fit onto 3 pages - please check text

@@
 
-Other packages are available for the post-processing of MCMC samples and are largely complementary to \tracer\ .
-The R package `coda' \citep{plummer2006coda} provides some of the functionality of \tracer\ within the R programming environment.
-`AWTY' \citep{nylander2007awty} explores the convergence of the phylogenetic tree parameter itself across multiple MCMC runs.
+%Other packages for post-processing of MCMC samples are
+
+Other packages are available for the post-processing of MCMC samples. % and are largely complementary to \tracer\ .
+ `Coda' \citep{plummer2006coda} provides some of the functionality of \tracer\ within the R programming environment, while `AWTY' \citep{nylander2007awty} explores the convergence of the phylogenetic tree parameter itself across multiple MCMC runs.
 %% AR - anything else? The MDS tree-space approaches of Hillis etc.?
 
-\vspace{-0.25cm}
+\vspace{-0.35cm}
 
@@
 
-\vspace{-0.25cm}
+\vspace{-0.35cm}
 

### 2018-02-02 manuscript/tracer17.tex
coda starts with an small c

@@
 Other packages are available for the post-processing of MCMC samples. % and are largely complementary to \tracer\ .
- `Coda' \citep{plummer2006coda} provides some of the functionality of \tracer\ within the R programming environment, while `AWTY' \citep{nylander2007awty} explores the convergence of the phylogenetic tree parameter itself across multiple MCMC runs.
+ `coda' \citep{plummer2006coda} provides some of the functionality of \tracer\ within the R programming environment, while `AWTY' \citep{nylander2007awty} explores the convergence of the phylogenetic tree parameter itself across multiple MCMC runs.
 %% AR - anything else? The MDS tree-space approaches of Hillis etc.?

### 2018-02-09 manuscript/syst_biol/tracer17_systbiol.tex
reformat for Syst Biol

--- 2018-02-02 manuscript/tracer17.tex
@@
-\documentclass{bioinfo_tracer}
-\copyrightyear{2017}
-\pubyear{2017}
+\documentclass[webpdf,mynatbib,nosurname,nogrid,noCE,noMSC]{SYS}
 
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-
+\usepackage{cpl-sft}
+\usepackage{url}
 \usepackage{graphicx}
 \usepackage{subfigure}
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
+\usepackage{enumitem}
 
-\usepackage{url}
+\setcounter{page}{1}
+
+\pvtitle{Software for Systematics and Evolution}
+\volname{}
+\jvolume{66}
+\jshort{syw}
+\jvol{VOL. 66}
+\jissue{2}
+\cename{PR}
+
+\mstype{Software for Systematics and Evolution}
+
+\pubyear{2018}
+
+\artid{087}
 
@@
 \begin{document}
-\firstpage{1}
 
-% Title must be 150 characters or less
-\title[Tracer]{Posterior summarisation in Bayesian phylogenetics using Tracer 1.7}
+\begin{document}
 
-\author[Rambaut \textit{et~al.}]{
-Andrew Rambaut\,$^{1}$,
-Alexei J.~Drummond\,$^{2,3}$,
-Dong Xie\,$^{2,3}$,
-Guy Baele$^{4}$,
-Marc A.~Suchard\,$^{5,6}$
+\title[RAMBAUT ET~AL.---TRACER 1.7]{
+Posterior summarisation in Bayesian phylogenetics using Tracer 1.7
+}
+
+\author[]{
+\textsc{Andrew} \surname{\textsc{Rambaut}}\,$^{1}$, 
+\textsc{Alexei J.} \surname{\textsc{Drummond}}\,$^{2,3}$, 
+\textsc{Dong} \surname{\textsc{Xie}}\,$^{2,3}$,
+\textsc{Guy} \surname{\textsc{Baele}}\,$^{4}$,
+\textsc{Marc A.} \surname{\textsc{Suchard}}\,$^{5,6}$
 }
@@
 \address{
-$^{1}$Institute of Evolutionary Biology, University of Edinburgh, Edinburgh, UK \\
-$^{2}$Department of Computer Science, University of Auckland, Auckland, NZ \\
-$^{3}$Centre for Computational Evolution, University of Auckland, Auckland, NZ \\
-$^{4}$Department of Microbiology and Immunology, Rega Institute, KU Leuven, Leuven, Belgium\\
-$^{5}$Department of Human Genetics, University of California, Los Angeles, USA \\
-$^{6}$Department of Biostatistics, University of California, Los Angeles, USA \\
-%$^{7}$Department of Human Genetics, David Geffen School of Medicine at UCLA, University of California, Los Angeles, USA\\
+$^{1}$Institute of Evolutionary Biology, University of Edinburgh, Edinburgh, UK;
+$^{2}$Department of Computer Science, University of Auckland, Auckland, NZ;
+$^{3}$Centre for Computational Evolution, University of Auckland, Auckland, NZ;
+$^{4}$Department of Microbiology and Immunology, Rega Institute, KU Leuven, Leuven, Belgium;
+$^{5}$Department of Human Genetics, University of California, Los Angeles, USA;
+$^{6}$Department of Biostatistics, University of California, Los Angeles, USA \\[6pt]
+$^{\ast}$Correspondence to: 
+\href{a.rambaut@ed.ac.uk}{\url{a.rambaut@ed.ac.uk}},
+\href{alexei@cs.auckland.ac.nz}{\url{alexei@cs.auckland.ac.nz}}
+and
+\href{msuchard@ucla.edu}{\url{msuchard@ucla.edu}}}
+
+\history{Received X February 2018; reviews returned X March 2018; accepted X April 2018}
+
+\editor{Associate Editor: To fill in}
+
+\abstract{
+Bayesian inference of phylogeny using Markov chain Monte Carlo (MCMC) plays a central role in understanding evolutionary history from molecular sequence data.
+Visualising and analysing the MCMC-generated samples from the posterior distribution is a key step in any non-trivial Bayesian inference. 
+We present the software package \tracer\ (version 1.7) for visualising and analysing the MCMC trace files generated through Bayesian phylogenetic inference.
+\tracer\ provides kernel density estimation, multivariate visualisation, demographic trajectory reconstruction, conditional posterior distribution summary and more.
+\tracer\ is open-source and available at \url{http://beast.community/tracer}.
+[Bayesian inference; phylogenetics; Markov chain Monte Carlo.]
+\vspace*{-25pt}
 }
-
-\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}
-
-\editor{Associate Editor: XXXXXXX}
 
@@
 
-% Please keep the abstract between 250 and 300 words
-\begin{abstract}
-
-\section{Motivation:}
-Bayesian inference of phylogeny using Markov chain Monte Carlo (MCMC) plays a central role in understanding evolutionary history from molecular sequence data.
-Visualising and analysing the MCMC-generated samples from the posterior distribution is a key step in any non-trivial Bayesian inference.
-%, and it is especially important for methods that employ MCMC.
-\section{Results:}
-We present the software package \tracer\ (version 1.7) for visualising and analysing the MCMC trace files generated through Bayesian phylogenetic inference.
-\tracer\ provides kernel density estimation, multivariate visualisation, demographic trajectory reconstruction, conditional posterior distribution summary and more.
-\section{Availability:}
-\tracer\ is open-source and available at \url{http://beast.community/tracer}.
-
-\section{Contact:}
-\href{a.rambaut@ed.ac.uk}{\url{a.rambaut@ed.ac.uk}},
-\href{alexei@cs.auckland.ac.nz}{\url{alexei@cs.auckland.ac.nz}}
-and
-\href{msuchard@ucla.edu}{\url{msuchard@ucla.edu}}
-
-\end{abstract}
-
-\section*{Introduction}
-
-Bayesian inference of phylogeny using Markov chain Monte Carlo (MCMC) \citep{rannala1996probability, mau1999bayesian, drummond2002estimating} flourishes as a popular approach to uncover the evolutionary relationships among taxa, such as genes, genomes, individuals or species.
+\indent Bayesian inference of phylogeny using Markov chain Monte Carlo (MCMC) \citep{rannala1996probability, mau1999bayesian, drummond2002estimating} flourishes as a popular approach to uncover the evolutionary relationships among taxa, such as genes, genomes, individuals or species.
 MCMC approaches generate samples of model parameter values - including the phylogenetic tree - drawn from their posterior distribution given molecular sequence data and a selection of evolutionary models.
@@
 
-\section*{Design and Implementation}
+\vspace{-2em}
+\section*{\textsc{Design and Implementation}}
 
@@
 which are intended in these other software packages.
-For each trace, \tracer\ examines the posterior samples from all the available parameters, treating continuous, integer and categorical parameters appropriately, and presents statistical summaries and visualizations.
+%For each trace,
+\tracer\ examines the posterior samples from all the available parameters, treating continuous, integer and categorical parameters appropriately,
+from a trace
+and presents statistical summaries and visualizations.
 Further, \tracer\ can analyse a single trace or combine samples from multiple files.
 Immediately apparent in the default \tracer\ view, the effective sample size (ESS) is one such statistic that
-allows users to assess the number of effectively independent draws from the posterior distribution the trace represents (Figure \ref{fig:overview}:a).
+allows users to assess the number of effectively independent draws from the posterior distribution the trace represents (Figure \ref{fig:overview}a).
 Colour coding assists the user in determining potential MCMC mixing problems, with cut-off values at 100 and 200.
@@
 
-Selecting multiple parameters from the ``Traces" panel on the left generates a side-by-side comparison or an overlay of the selected parameters' visualisations (Figure \ref{fig:overview}:b-e).
+Selecting multiple parameters from the ``Traces" panel on the left generates a side-by-side comparison or an overlay of the selected parameters' visualisations (Figure \ref{fig:overview}b-e).
 Multiple trace files can be selected in a similar fashion to compare posterior samples between different replicates of an analysis.
@@
 
-\begin{itemize}
-
+\begin{itemize}[leftmargin=*]
+\vspace{-0.5em}
 \item Estimates: Reports common summary statistics such as the sample mean, standard deviation, highest posterior density interval and ESS.
-Also presents a histogram of sample values for a single selected parameter (Figure \ref{fig:overview}:a) or  side-by-side boxplots for multiple continuous parameters (Figure \ref{fig:overview}:b).
-
-\item Marginal Density: Draws density plots for the selected parameter(s), including  kernel density estimates (Figure \ref{fig:overview}:c), histograms and violin plots (Figure \ref{fig:overview}:d) for continuous parameters and frequency plots for categorical or integer parameters.
-
+Also presents a histogram of sample values for a single selected parameter (Figure \ref{fig:overview}a) or  side-by-side boxplots for multiple continuous parameters (Figure \ref{fig:overview}b).
+\item Marginal Density: Draws density plots for the selected parameter(s), including  kernel density estimates (Figure \ref{fig:overview}c), histograms and violin plots (Figure \ref{fig:overview}d) for continuous parameters and frequency plots for categorical or integer parameters.
+\vspace*{-0.25em}
 \item Joint-Marginal: Visualisation in this panel appears after selecting two or more parameters and the plot form depends on the parameter types.
 We show several examples in the next section of the paper.
-
-\item Trace: Constructs line plots connecting the sequential samples of one or more selected parameters against state or generation number (Figure \ref{fig:overview}:e).
-Users typically use this plot to assess mixing, select a suitable burn-in and identify trends that suggest convergence issues.
-\end{itemize}
-%
-
 %
 \begin{figure}[t]
-\subfigure[Tracer overview]{\hspace{0.40cm}\includegraphics[width=0.432\textwidth]{./figures/rabv-overview-3.pdf}}
-\subfigure[Boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-boxplots.pdf}}
-\subfigure[Density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-density.pdf}}\\
-\subfigure[Violin]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-violin.pdf}}
-\subfigure[Trace]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-traces.pdf}}
+\subfigure[Tracer overview]{\hspace{0.40cm}\includegraphics[width=0.432\textwidth]{../figures/rabv-overview-3.pdf}} \\
+\subfigure[Boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{../figures/rabv-boxplots.pdf}}
+\subfigure[Density]{\includegraphics[width = .23\textwidth, height = 3cm]{../figures/rabv-density.pdf}}\\
+\subfigure[Violin]{\includegraphics[width = .23\textwidth, height = 3cm]{../figures/rabv-violin.pdf}}
+\subfigure[Trace]{\includegraphics[width = .23\textwidth, height = 3cm]{../figures/rabv-traces.pdf}}
 \vspace{-0.25cm}
@@
 \begin{figure}[t]
-\subfigure[Bubble chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integerjointmarginal-2.pdf}}
-\subfigure[Marginal density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integermarginaldensity.pdf}} \\
-\subfigure[Scatter plot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-correlation.pdf}}
-\subfigure[Correlation plot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-joint-real.pdf}}
+\subfigure[Bubble chart]{\includegraphics[width = .23\textwidth, height = 3cm]{../figures/rabv-integerjointmarginal-2.pdf}}
+\subfigure[Marginal density]{\includegraphics[width = .23\textwidth, height = 3cm]{../figures/rabv-integermarginaldensity.pdf}} \\
+\subfigure[Scatter plot]{\includegraphics[width = .23\textwidth, height = 3cm]{../figures/rabv-correlation.pdf}}
+\subfigure[Correlation plot]{\includegraphics[width = .23\textwidth, height = 3cm]{../figures/rabv-joint-real.pdf}}
 \vspace{-0.25cm}
-%\subfigure[Box-and-whisker plot]{\includegraphics[width = .465\textwidth, height = 3cm]{./figures/rabv-catreal.pdf}}
+%\subfigure[Box-and-whisker plot]{\includegraphics[width = .465\textwidth, height = 3cm]{../figures/rabv-catreal.pdf}}
 \caption{Multi-parameter visualisations of: (a) the joint probability distribution of two integer variables through a bubble chart; (b) the marginal density of multiple integer or categorical variables through frequency plots; (c) two continuous variables through a classic scatter or correlation plot; (d) multiple ($> 2$) continuous variables using large correlation matrices.
@@
 \end{figure}
+%
+\item Trace: Constructs line plots connecting the sequential samples of one or more selected parameters against state or generation number (Figure \ref{fig:overview}e).
+Users typically use this plot to assess mixing, select a suitable burn-in and identify trends that suggest convergence issues.
+\end{itemize}
 %
@@
 
-\vspace{-0.5cm}
+\vspace{-0.7cm}
 
-\section*{Example}
+\section*{\textsc{Example}}
 
-\subsection*{Cross-species dynamics of rabies in North American bats}
+\subsection*{Dynamics of rabies across North American bats}
 
@@
 %While a single parameter of such types can easily be visualised using a frequency plot,
-In \tracer, a bubble chart visualises the joint probability distribution between two integer or categorical traces (see Figure \ref{fig:4tabs}:a).
+In \tracer, a bubble chart visualises the joint probability distribution between two integer or categorical traces (see Figure \ref{fig:4tabs}a).
 Circle area is proportional to the joint probability, %the blue coloured circle is in the credible set of given a probability threshold default to 0.95, the red is in the set not in the credible set.
@@
 %Multiple integer parameters can also be visualised using
-Marginal density plots can also display multiple integer parameters, each with unique colour scales (see Figure \ref{fig:4tabs}:b).
+Marginal density plots can also display multiple integer parameters, each with unique colour scales (see Figure \ref{fig:4tabs}b).
 With approximately equal numbers of transition rates, both figures suggest similar host and location trait model complexity.
@@
 %The visualisation options for integer or categorical parameters are complemented with
- popular visualisations for continuous parameters, including scatter plots for two parameters (see Figure \ref{fig:4tabs}:c),
- and extensions for correlations between $\ge\hspace{-0.32em}2$ continuous parameters (Figure \ref{fig:4tabs}:d; \citet{Murdoch}).
+ popular visualisations for continuous parameters, including scatter plots for two parameters (see Figure \ref{fig:4tabs}c),
+ and extensions for correlations between $\ge\hspace{-0.32em}2$ continuous parameters (Figure \ref{fig:4tabs}d; \citet{Murdoch}).
  %, which indicates a strong positive correlation between the clock rates of both discrete trait models.
 %We also provide an extension of such a scatter plot for multiple (i.e.
-more than two) continuous variables (see Figure \ref{fig:4tabs}:d; \citet{Murdoch}), which makes use of coloured ellipses to indicate the strength of correlation between each pair of continuous variables.
+more than two) continuous variables (see Figure \ref{fig:4tabs}d; \citet{Murdoch}), which makes use of coloured ellipses to indicate the strength of correlation between each pair of continuous variables.
 Colour gradients indicate strength and direction of the correlation, from red (strong negative) to blue (strong positive).
@@
 %Finally, in order to display the joint marginal distribution between a continuous variable and an integer or categorical variable, we make use of box-and-whisker plots.
-%Using such a plot (Figure \ref{fig:4tabs}:e), we can infer that the range of the non-zero transition rates can depend on the ancestral location state inferred at the root of the underlying phylogeny.
+%Using such a plot (Figure \ref{fig:4tabs}e), we can infer that the range of the non-zero transition rates can depend on the ancestral location state inferred at the root of the underlying phylogeny.
 
@@
 \centerline{
-\includegraphics[width=0.46\textwidth]{./figures/rabv-skygrid.pdf}
+\includegraphics[width=0.46\textwidth]{../figures/rabv-skygrid.pdf}
 }
@@
 
-\vspace{-0.35cm}
+\vspace{-0.55cm}
 
-\section*{Availability}
+\section*{\textsc{Availability}}
 
@@
 
-%\section*{Old Text}
-%
-%%You can also select the "Demographic Analysis" from the Analysis menu - This plots the distribution of demographic population sizes over time for a number of models (constant size, exponential growth \& logistic growth) that are available in BEAST.
-This involves you selecting the traces for each parameter of the model.
-You should only select the model that was actually run under BEAST (e.g., if you ran an exponential growth model, you shouldn't plot the constant population size model).
-%
-%%The "Analysis" menu also contains options for performing Bayesian Skyline reconstructions and for calculating Bayes Factors between runs.
-%
-%Feature list:
-%\begin{itemize}
-%\item KDEs
-%\item cheap marginal likelihood estimators
-%\item demographic trajectory reconstruction
-%\end{itemize}
-%
-%%Model selection HME and AICM have been removed
-%New features:
-%\begin{itemize}
-%\item Three trace types: real, integer, string
-%\item Kernel density estimates
-%\item Conditional posterior distribution
-%%\item HME
-%%\item AICM
-%\end{itemize}
-%
-%%\begin{figure}[H]
-%%\includegraphics[width=.5\textwidth]{./figures/comp-95HPD.pdf}
-%%\caption{The comparison of  95\% HPD intervals from multi-trace}
-%%\label{fig:comp95HPD}
-%%\end{figure}
-%
-%%\begin{figure}[H]
-%%\includegraphics[width=.5\textwidth]{./figures/jointPrInt.png}
-%%\caption{The Joint probability table of two integer traces}
-%%\label{fig:int:jointpr}
-%%\end{figure}
-%
-%%\begin{figure}[ht]
-%%\includegraphics[width=.5\textwidth]{./figures/frequency.pdf}
-%%\caption{A frequency distribution for a continuous trace}
-%%\label{fig:freq}
-%%\end{figure}
-%%\begin{figure}[ht]
-%%\includegraphics[width=.5\textwidth]{./figures/multiKDE.pdf}
-%%\caption{KDE of multiple traces}
-%%\label{fig:multiKDE}
-%%\end{figure}
-%%\begin{figure}[ht]
-%%\includegraphics[width=.5\textwidth]{./figures/joint-marginal.pdf}
-%%\caption{The joint-marginal distribution of two selected tes}
-%%\label{fig:trace}
-%%\end{figure}
-%%\begin{figure}[ht]
-%%\includegraphics[width=.5\textwidth]{./figures/trace.pdf}
-%%\caption{The selected traces against state}
-%%\label{fig:trace}
-%%\end{figure}
-%
-%
-%\subsubsection*{Diagnostics:} Blah blah
-%
-%\subsubsection*{Demographic reconstruction:}
-%
-%%\subsubsection*{Model selection:}
-%
-%\subsubsection*{Conditional posterior distribution:}
+\vspace{-0.55cm}
 
-%Fairly general solution to looking at conditional posterior distributions;
-
-%Specifically, support for BSSVS forms of model averaging, in which some parameters are only in the likelihood when their submodel is "indicated" by some indicator function that is usually a discrete, integer or boolean variable.
-In that case the posterior of the parameter should not include the states when it was sampled only in the prior because the submodel it belonged to was "turned off".
-This is relevant for EBSP, Random Local Clocks model, microsatellite model averaging and relaxed clock model averaging methods all from my group in last couple of years.
-Also relevant for BSSVS in phylogeography as well depending on how the state is logged.
-%\section*{Taken from the Tracer website}
-
-%Tracer is a program for analysing the trace files generated by Bayesian MCMC runs (that is, the continuous parameter values sampled from the chain).
-It can be used to analyse runs of BEAST \citep{drummond2012bayesian,drummond2012bayesian}, BEAST2 \citep{bouckaert2014beast2}, MrBayes \citep{ronquist2012mrbayes}, RevBayes \citep{hohna2016revbayes}, LAMARC \citep{kuhner2006lamarc}, Migrate \citep{beerli2006comparison} and possibly other MCMC programs.
-
-%Although Tracer can be used with programs other than BEAST, users may find it useful to join the BEAST users mailing list.
-This is used to announce new versions and advise users about bugs and problems.
-
-%You can join the mailing list here:
-%http://groups.google.com/group/beast-users
-
-%The website for BEAST (and Tracer) is here:
-%http://beast.bio.ed.ac.uk/
-
-%At present there is no detailed manual for this application, you will simply have to play around and see what happens.
-Basically you can select the trace file in the top left of the window, the individual parameter in the bottom left and the analysis appears on the right.
-
-%There are 4 analysis tabs to choose from:
-
-%Estimates - this shows the mean, stdev, confidence intervals and other statistics about the selected parameter. A frequency distribution will also be plotted.
-%Density - this shows the Bayesian posterior density plot for the selected parameter.
-%Joint-Marginal - this only appears if exactly 2 parameters are chosen (hold down shift to select multiple parameters).
-It then plots one against the other to look at their joint-marginal distribution.
-%Trace - this shows the trace of the parameter against state or generation number. Use this to check mixing, choose a suitable burn-in and look for trends that might suggest problems with convergence.
-%Multiple parameters can be selected by holding down the shift key.
-This will overlay the plots for the different parameters allowing comparisons to be made.
-You can also select multiple trace files as well to compare different runs.
-If multiple trace files have the same trace names then a "Combined" trace will automatically appear.
-This can be selected as well as the individual trace files.
-
-%You can also select the "Demographic Analysis" from the Analysis menu - This plots the distribution of demographic population sizes over time for a number of models (constant size, exponential growth \& logistic growth) that are available in BEAST.
-This involves you selecting the traces for each parameter of the model.
-You should only select the model that was actually run under BEAST (e.g., if you ran an exponential growth model, you shouldn't plot the constant population size model).
-
-%The "Analysis" menu also contains options for performing Bayesian Skyline reconstructions and for calculating Bayes Factors between runs.
-
-%The "Print" function in the "File" menu will print the current graph or table and the "Export Data" function can be used to export the data from the plots for use in another graphic package.
-
-%To export the currently displayed graphic use the "Export PDF" function in the "File" menu.
-
-\vspace{-0.35cm}
-
-\section*{Acknowledgements}
+\section*{\textsc{Acknowledgements}}
 
@@
 
-\vspace{-0.35cm}
+\vspace*{-2em}
 
-%\section*{References}
-% The bibtex filename
-\bibliographystyle{natbib}
-\bibliography{tracer17}
+\bibliographystyle{chicago}
+\bibliography{../tracer17}
 

### 2018-02-10 manuscript/syst_biol/tracer17_systbiol.tex
small tweaks

@@
 \tracer\ is open-source and available at \url{http://beast.community/tracer}.
-[Bayesian inference; phylogenetics; Markov chain Monte Carlo.]
+[Bayesian inference; phylogenetics; Markov chain Monte Carlo; visualisation]
 \vspace*{-25pt}
@@
 
-\subsection*{Dynamics of rabies across North American bats}
+\subsection*{Cross-species dynamics of North American bat rabies}
 
@@
  popular visualisations for continuous parameters, including scatter plots for two parameters (see Figure \ref{fig:4tabs}c),
- and extensions for correlations between $\ge\hspace{-0.32em}2$ continuous parameters (Figure \ref{fig:4tabs}d; \citet{Murdoch}).
+ and extensions for correlations between $\ge\hspace{-0.08em}2$ continuous parameters (Figure \ref{fig:4tabs}d; \citet{Murdoch}).
  %, which indicates a strong positive correlation between the clock rates of both discrete trait models.

### 2018-04-12 src/tracer/application/TracerApp.java
Updating packaging and versions

@@ -126,7 +126,7 @@ public class TracerApp extends MultiDocApplication {
             }
 
             final String nameString = "Tracer";
-            final String versionString = "v1.7-Pre20180225";
+            final String versionString = "v1.7-Pre20180412";
             String aboutString = "<html><font face=\"helvetica,san-serif\"><center><p>MCMC Trace Analysis Tool<br>" +
                     "Version " + versionString + ", 2003-2018</p>" +
                     "<p>by<br>" +

### 2018-04-12 manuscript/response.md
Create response.md

# Reviewer 1

*This is the first paper describing Tracer, the most frequently used software for convergence analysis of phylogenetic MCMC. Thus this paper certainly merits publication, and will no doubt get thousands of citations. Tracer was already very useful, and this new version brings additional useful functionality. The paper is mostly an announcement, and is well written though we have a few comments.*

Thanks

## Regarding the software:

*S1: How does the conditional posterior visualization work? That seems like a major improvement, but it’s not shown anywhere or described beyond a mention in the intro. The documentation basically says “try things out!”-- this mostly works except for here.*

** To do **

*S2: We were happy to see that the HME is no longer an option for model comparison (given its notorious bias).*

No response required

*S3: Box plots and violin plots are a fantastic improvement for viewing multiple distributions at the same time! The >2 dimensional correlation plots are also a welcome addition to the toolkit. *

No response required

*S4: We miss the ability to adjust the number of bins for histograms of univariate marginal distributions.*

This feature has been restored [Issue #146](https://github.com/beast-dev/tracer/issues/146).

*S5: Please update the link on http://beast.community/tracer , which currently links to 1.6 (even though it says it’s linking to 1.7)*

** To do **

* P1: People mostly use Tracer to diagnose convergence problems. I think the best use of a paper about Tracer would be to walk readers through a side-by-side comparison of two runs, one of which has converged and another that hasn’t. This is additional work, but I think it would be very valuable to guide the community in making such calls. At least, there should be a link in the paper to some material giving such an example. The current Figures 1 b-e and 2a-c feel a little generic-- we know what these plot types look like. *

** To do **

Details:

* Title: We cannot resist the temptation to suggest to the more active “Summarizing Bayesian phylogenetic posterior distributions using Tracer 1.7”*

*L68: the BSSVS description doesn’t really feel like it comes to a satisfactory conclusion. Can you more clearly describe how Tracer helps with this type of analysis? Is the sentence at line 80 a continuation of this thought, or starting something new?*

*L197: the proper citation is “...finds its rootS”*

*Fig 1a: the parameter name being open for editing is distracting*

*L108: I am not sure I understand the comparison (“model complexity”) being made between the marginal and joint marginal plots. If it is just the number of nonzero parameters, it could be stated more clearly.*

*L130: coda also offers diagnostics absent from Tracer, such as Gelman-Rubin statistics.*

Please cite RWTY https://academic.oup.com/mbe/article/34/4/1016/2900564 , which has some features (e.g. topology diagnostics) that are not present in Tracer.
Figure 3 has no axis labels of any kind

### 2018-04-12 manuscript/response.md
Update response.md

@@
 
-Please cite RWTY https://academic.oup.com/mbe/article/34/4/1016/2900564 , which has some features (e.g. topology diagnostics) that are not present in Tracer.
-Figure 3 has no axis labels of any kind
+*Please cite RWTY https://academic.oup.com/mbe/article/34/4/1016/2900564 , which has some features (e.g. topology diagnostics) that are not present in Tracer.*
+
+*Figure 3 has no axis labels of any kind*
+
+# Reviewer 2
+
+*Line 36: Are there any references or reasons you can provide why should use cut-off values of 100 and 200? Or are these arbitrary?*
+
+*Line 43: If I'm correct, then the file names don't need to be same but the parameter name. Could you please correct this.*
+
+*Line 82-86: I was wondering if the demographic plotting are truly restricted to these models? There seem to be several more options available specifically for the "Demographic Reconstruction". *
+
+### Feature suggestions (these are rather general and not necessary for the manuscript):
+
+* 1) It would be nice to export a csv file with the parameter estimates (mean, 95% HPD,  etc) and ESS values so that this can be used easier in reports.*
+
+This has been implemented already under the File menu -> Export Data Table. ** This needs to be tested ***
+
+*2) There seems to be some memory issue in Tracer when multiple large files are opened. Closing and opening gets around the issue. It would be nice if that is fixed in some future release.*
+
+Test loading a too large file.
+
+# Editor's comments
+
+*Almost all of the material is clear and understandable, even to readers who have never used the software.
+For me the exception was the paragraph starting at 66.
+Since this appears to be a new feature, it probably could be elaborated on.
+Some general description of the types of things that can be conditioned on (models, the presence of variables in models, ...) and how the software does this, possibly with a figure, might help before or with some of the specifics on lines 68-79; I didn't actually see much about this in briefly looking at the tutorials on beast.community/tracer.*
+
+*-Fig 1(b)-(d): If you want to make space for additions, two of these could be cut and simply described as a list of (very familiar) available graphics for comparing multiple (marginal) distributions.
+By contrast, Fig 2(b)-(d) are less familiar and really need a figure to accompany the text description.*
+
+*-l67: Something like the sentence starting on l71 could be moved here to indicate generally how the conditional distributions are constructed from the sampled parameter values.
+It would be valuable to say something about how generally the indicator functions can be defined, how they are selected...*
+
+*-l79: `Finally,...' Should this be a new paragraph?*

### 2018-04-12 manuscript/response.md
Update response.md

@@
 
-* P1: People mostly use Tracer to diagnose convergence problems.
+*P1: People mostly use Tracer to diagnose convergence problems.
 I think the best use of a paper about Tracer would be to walk readers through a side-by-side comparison of two runs, one of which has converged and another that hasn’t.
@@
 
-Details:
-
-* Title: We cannot resist the temptation to suggest to the more active “Summarizing Bayesian phylogenetic posterior distributions using Tracer 1.7”*
+*Title: We cannot resist the temptation to suggest to the more active “Summarizing Bayesian phylogenetic posterior distributions using Tracer 1.7”*
 

### 2018-04-12 manuscript/response.md
Update response.md

@@
 
-** To do **
+**To do**
 
@@
 
-* 1) It would be nice to export a csv file with the parameter estimates (mean, 95% HPD,  etc) and ESS values so that this can be used easier in reports.*
+*1) It would be nice to export a csv file with the parameter estimates (mean, 95% HPD,  etc) and ESS values so that this can be used easier in reports.*
 
@@
 
-Test loading a too large file.
+**Test loading a too large file.**
 

### 2018-04-12 packaging_tools/mac/createdmg.sh
Create createdmg.sh

@@ -1,0 +1,58 @@
+#!/bin/sh
+cd release/Mac
+
+export source=Tracer
+export applicationName=Tracer
+export version=`perl -e 'while($s=<>) {if ($s=~/"version"/) {$s =~ /value="([^"]*)"/; print $1;}}' < ../../build.xml`
+export title="Tracer v${version}"
+export size=1g
+export backgroundPictureName=install.png
+export finalDMGName=${title}
+
+
+mkdir "BEAST/BEAST ${version}"
+mv BEAST/* "BEAST/BEAST ${version}"
+
+cp install.png ${source}/${backgroundPictureName}
+
+hdiutil create -srcfolder "${source}" -volname "${title}" -fs HFS+ \
+      -fsargs "-c c=64,a=16,e=16" -format UDRW -size ${size}k pack.temp.dmg
+
+export device=$(hdiutil attach -readwrite -noverify -noautoopen "pack.temp.dmg" | \
+         egrep '^/dev/' | sed 1q | awk '{print $1}')
+
+
+echo '
+   tell application "Finder"
+     tell disk "'${title}'"
+           open
+           set current view of container window to icon view
+           set toolbar visible of container window to false
+           set statusbar visible of container window to false
+           set the bounds of container window to {400, 100, 885, 430}
+           set theViewOptions to the icon view options of container window
+           set arrangement of theViewOptions to not arranged
+           set background picture of theViewOptions to file "'${backgroundPictureName}'"
+           set icon size of theViewOptions to 72
+           make new alias file at container window to POSIX file "/Applications" with properties {name:"Applications"}
+           set position of item "'${applicationName}' '${version}'" of container window to {100, 100}
+           set position of item "'${backgroundPictureName}'" of container window to {800, 100}
+           set position of item "Applications" of container window to {375, 100}
+			close
+			open
+           update without registering applications
+           delay 5
+#           eject
+     end tell
+   end tell
+' | osascript
+
+echo "OSASCRIPT DONE"
+ 
+ 
+chmod -Rf go-w /Volumes/"${title}"
+sync
+sync
+hdiutil detach ${device}
+hdiutil convert "pack.temp.dmg" -format UDZO -imagekey zlib-level=9 -o "${finalDMGName}"
+rm -f /pack.temp.dmg 

### 2018-04-12 packaging_tools/mac/createdmg.sh
Mac packaging now creates a drag to Applications folder type disk image. Issue #86

@@ -1,19 +1,15 @@
 #!/bin/sh
 cd release/Mac
 
-export source=Tracer
-export applicationName=Tracer
+export title="Tracer"
 export version=`perl -e 'while($s=<>) {if ($s=~/"version"/) {$s =~ /value="([^"]*)"/; print $1;}}' < ../../build.xml`
-export title="Tracer v${version}"
+export applicationName="${title} v${version}"
 export size=1g
 export backgroundPictureName=install.png
-export finalDMGName=${title}
+export finalDMGName="${title} v${version}"
+export source="${title} v${version}"
 
-
-mkdir "BEAST/BEAST ${version}"
-mv BEAST/* "BEAST/BEAST ${version}"
-
-cp install.png ${source}/${backgroundPictureName}
+echo ${version}
 
 hdiutil create -srcfolder "${source}" -volname "${title}" -fs HFS+ \
       -fsargs "-c c=64,a=16,e=16" -format UDRW -size ${size}k pack.temp.dmg
@@ -33,11 +29,11 @@ echo '
            set theViewOptions to the icon view options of container window
            set arrangement of theViewOptions to not arranged
            set background picture of theViewOptions to file "'${backgroundPictureName}'"
-           set icon size of theViewOptions to 72
+           set icon size of theViewOptions to 96
            make new alias file at container window to POSIX file "/Applications" with properties {name:"Applications"}
-           set position of item "'${applicationName}' '${version}'" of container window to {100, 100}
-           set position of item "'${backgroundPictureName}'" of container window to {800, 100}
-           set position of item "Applications" of container window to {375, 100}
+           set position of item "'${applicationName}'" of container window to {100, 120}
+           set position of item "'${backgroundPictureName}'" of container window to {800, 120}
+           set position of item "Applications" of container window to {375, 120}
 			close
 			open
            update without registering applications
@@ -54,5 +50,5 @@ chmod -Rf go-w /Volumes/"${title}"
 sync
 sync
 hdiutil detach ${device}
-hdiutil convert "pack.temp.dmg" -format UDZO -imagekey zlib-level=9 -o "${finalDMGName}"
-rm -f /pack.temp.dmg 
+hdiutil convert "pack.temp.dmg" -format UDZO -imagekey zlib-level=9 -o "../${finalDMGName}"
+rm -f /pack.temp.dmg

### 2018-04-12 manuscript/syst_biol/tracer17_systbiol.tex
depreciate old file
fix compilation error and added citation

@@
 \newcommand{\tracer}{\emph{Tracer}}
-
-\begin{document}
 
@@
 Other packages are available for the post-processing of MCMC samples. % and are largely complementary to \tracer\ .
- `coda' \citep{plummer2006coda} provides some of the functionality of \tracer\ within the R programming environment, while `AWTY' \citep{nylander2007awty} explores the convergence of the phylogenetic tree parameter itself across multiple MCMC runs.
+ `coda' \citep{plummer2006coda} provides some of the functionality of \tracer\ within the R programming environment, while `AWTY' \citep{nylander2007awty} and `RWTY' \citep{warren2017rwty} explore the convergence of the phylogenetic tree parameter itself across multiple MCMC runs.
 %% AR - anything else? The MDS tree-space approaches of Hillis etc.?
@@
 \end{document}
+
+% COVER
+% Please find attached our manuscript introducing the Tracer software tool for publication in Systematic Biology.
+While Tracer finds common use amongst phylogenetics and carries several thousand citations via GoogleScholar, there is no publication that announces its availability nor describes its features.
+We believe our manuscript will become the de facto citation for Tracer and will quickly emerge as an impactful publication in Systematic Biology.

### 2018-04-12 src/tracer/analysis/BayesianSkylineDialog.java
Fixing issue #9

@@ -203,6 +203,7 @@ public class BayesianSkylineDialog {
 
         setArguments(temporalAnalysisFrame);
 
+        rootHeightCombo.removeAllItems();
         for (int j = 0; j < traceList.getTraceCount(); j++) {
             String statistic = traceList.getTraceName(j);
             rootHeightCombo.addItem(statistic);

### 2018-04-12 src/tracer/analysis/ExtendedBayesianSkylineDialog.java

@@ -203,6 +203,7 @@ public class ExtendedBayesianSkylineDialog {
 
         setArguments(temporalAnalysisFrame);
 
+        rootHeightCombo.removeAllItems();
         for (int j = 0; j < traceList.getTraceCount(); j++) {
             String statistic = traceList.getTraceName(j);
             rootHeightCombo.addItem(statistic);

### 2018-04-12 src/tracer/analysis/GMRFSkyrideDialog.java

@@ -195,6 +195,7 @@ public class GMRFSkyrideDialog {
 
         setArguments(temporalAnalysisFrame);
 
+        rootHeightCombo.removeAllItems();
         for (int j = 0; j < traceList.getTraceCount(); j++) {
             String statistic = traceList.getTraceName(j);
             rootHeightCombo.addItem(statistic);

### 2018-04-12 src/tracer/analysis/LineagesThroughTimeDialog.java

@@ -111,6 +111,7 @@ public class LineagesThroughTimeDialog {
 
         setArguments(temporalAnalysisFrame);
 
+        rootHeightCombo.removeAllItems();
         for (int j = 0; j < traceList.getTraceCount(); j++) {
             String statistic = traceList.getTraceName(j);
             rootHeightCombo.addItem(statistic);

### 2018-04-12 src/tracer/analysis/SkyGridDialog.java

@@ -192,7 +192,8 @@ public class SkyGridDialog {
         }
 
         setArguments(temporalAnalysisFrame);
-
+        
+        rootHeightCombo.removeAllItems();
         for (int j = 0; j < traceList.getTraceCount(); j++) {
             String statistic = traceList.getTraceName(j);
             rootHeightCombo.addItem(statistic);

### 2018-04-12 src/tracer/analysis/TimeDensityDialog.java

@@ -72,6 +72,7 @@ public class TimeDensityDialog {
 
         setArguments();
 
+        traceCombo.removeAllItems();
         for (int j = 0; j < traceList.getTraceCount(); j++) {
             String statistic = traceList.getTraceName(j);
             traceCombo.addItem(statistic);

### 2018-04-12 src/tracer/analysis/TraitThroughTimeDialog.java

@@ -115,6 +115,7 @@ public class TraitThroughTimeDialog {
 
         setArguments(temporalAnalysisFrame);
 
+        rootHeightCombo.removeAllItems();
         for (int j = 0; j < traceList.getTraceCount(); j++) {
             String statistic = traceList.getTraceName(j);
             rootHeightCombo.addItem(statistic);

### 2018-04-12 manuscript/tracer17_systbiol.tex
move into main directory

--- 2018-04-12 manuscript/syst_biol/tracer17_systbiol.tex
@@
 Multiple trace files can be selected in a similar fashion to compare posterior samples between different replicates of an analysis.
-If multiple trace files have the same file name,  then a ``Combined'' trace appears automatically.
+If multiple trace files have the same file name,  then a ``Combined'' trace appears automatically. %GB: or have the same collection of logged parameters
 % , which can be selected as well as the individual trace files.
@@
 \begin{figure}[t]
-\subfigure[Tracer overview]{\hspace{0.40cm}\includegraphics[width=0.432\textwidth]{../figures/rabv-overview-3.pdf}} \\
-\subfigure[Boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{../figures/rabv-boxplots.pdf}}
-\subfigure[Density]{\includegraphics[width = .23\textwidth, height = 3cm]{../figures/rabv-density.pdf}}\\
-\subfigure[Violin]{\includegraphics[width = .23\textwidth, height = 3cm]{../figures/rabv-violin.pdf}}
-\subfigure[Trace]{\includegraphics[width = .23\textwidth, height = 3cm]{../figures/rabv-traces.pdf}}
+\subfigure[Tracer overview]{\hspace{0.40cm}\includegraphics[width=0.432\textwidth]{./figures/rabv-overview-3.pdf}} \\
+\subfigure[Boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-boxplots.pdf}}
+\subfigure[Density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-density.pdf}}\\
+\subfigure[Violin]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-violin.pdf}}
+\subfigure[Trace]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-traces.pdf}}
 \vspace{-0.25cm}
@@
 \begin{figure}[t]
-\subfigure[Bubble chart]{\includegraphics[width = .23\textwidth, height = 3cm]{../figures/rabv-integerjointmarginal-2.pdf}}
-\subfigure[Marginal density]{\includegraphics[width = .23\textwidth, height = 3cm]{../figures/rabv-integermarginaldensity.pdf}} \\
-\subfigure[Scatter plot]{\includegraphics[width = .23\textwidth, height = 3cm]{../figures/rabv-correlation.pdf}}
-\subfigure[Correlation plot]{\includegraphics[width = .23\textwidth, height = 3cm]{../figures/rabv-joint-real.pdf}}
+\subfigure[Bubble chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integerjointmarginal-2.pdf}}
+\subfigure[Marginal density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integermarginaldensity.pdf}} \\
+\subfigure[Scatter plot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-correlation.pdf}}
+\subfigure[Correlation plot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-joint-real.pdf}}
 \vspace{-0.25cm}
-%\subfigure[Box-and-whisker plot]{\includegraphics[width = .465\textwidth, height = 3cm]{../figures/rabv-catreal.pdf}}
+%\subfigure[Box-and-whisker plot]{\includegraphics[width = .465\textwidth, height = 3cm]{./figures/rabv-catreal.pdf}}
 \caption{Multi-parameter visualisations of: (a) the joint probability distribution of two integer variables through a bubble chart; (b) the marginal density of multiple integer or categorical variables through frequency plots; (c) two continuous variables through a classic scatter or correlation plot; (d) multiple ($> 2$) continuous variables using large correlation matrices.
@@
 \centerline{
-\includegraphics[width=0.46\textwidth]{../figures/rabv-skygrid.pdf}
+\includegraphics[width=0.46\textwidth]{./figures/rabv-skygrid.pdf}
 }
@@
 \bibliographystyle{chicago}
-\bibliography{../tracer17}
+\bibliography{./tracer17}
 

### 2018-04-12 manuscript/response.md
Update response.md
fixed citation
link to tracer download updated
convergence is boring

@@
+Dear Prof. Near,
+
+Attached you will find our revised manuscript USYB-2018-044.
+We thank the Associate Editor and Reviewers for their thorough reading, constructive comments and helpful insight.
+Below you will find our point-by-point response to each comment and details on how these comments have improved our manuscript.
+
+Yours sincerely,
+
+Andrew Rambaut, Alexei Drummond, Dong Xie, Guy Baele and Marc Suchard
+
 # Reviewer 1
@@
 
-Thanks
+We thank the Reviewer for this encouragement.
 
@@
 
-No response required
+No response required.
 
@@
 
-This feature has been restored [Issue #146](https://github.com/beast-dev/tracer/issues/146).
+At the request of the Reviewer, we have restored this feature in the software. See [Issue #146](https://github.com/beast-dev/tracer/issues/146).
 
@@
 
-**To do**
+Done.
 
@@
 
-** To do **
+We believe that Figures 1b-e and 2a-c feel generic to the Reviewer because the Reviewer is likely to be a long-term user of the Tracer software.
+A major goal of this manuscript is to raise the awareness for current non-users about the convenient graphical user interface of Tracer as opposed to command-line driven approaches of alternatives, like `coda`.
+We further disagree that most people employ Tracer simply to diagnose convergence problems; we, for example, commonly use Tracer to quickly calculate summary statistics (posterior means, 95% highest probably density intervals).
+However, we agree with the Reviewer that readers should be aware that Tracer is useful in diagnosing convergence problems and have added associated text highlighting a new tutorial on convergence to the manuscript.
 
@@
 
+Fixed.
+
 *Fig 1a: the parameter name being open for editing is distracting*
@@
 *Please cite RWTY https://academic.oup.com/mbe/article/34/4/1016/2900564 , which has some features (e.g. topology diagnostics) that are not present in Tracer.*
+
+Done.
 

### 2018-04-12 manuscript/tracer17_systbiol.tex
convergence is boring

@@
 and the non-parametric Bayesian skyline \citep{drummond2005bayesian,heledDrummond2008}, skyride \citep{minin2008smooth} and skygrid \citep{gill2012improving}.
-%Not sure if the statement below is correct ...
-%All of these demographic models are available in BEAST \citep{drummond2012bayesian} and BEAST2 \citep{bouckaert2014beast2}.
-%Besides these, ``Lineages Through Time'' (?) is also available to plot the quantiles of the number of lineages against time, which describes the change of diversification of all organisms over time.
-%citation?
 
@@
 
-%Other packages for post-processing of MCMC samples are
-
-Other packages are available for the post-processing of MCMC samples. % and are largely complementary to \tracer\ .
+Other packages are available for the post-processing of MCMC samples.
  `coda' \citep{plummer2006coda} provides some of the functionality of \tracer\ within the R programming environment, while `AWTY' \citep{nylander2007awty} and `RWTY' \citep{warren2017rwty} explore the convergence of the phylogenetic tree parameter itself across multiple MCMC runs.
+%These alternative packages focus on MCMC convergence diagnostics, such as computing Gelman-Rubin statistics across multiple chains
 %% AR - anything else? The MDS tree-space approaches of Hillis etc.?
@@
 \tracer\ is open-source under the GNU lesser general public license and available in both source code (\url{https://github.com/beast-dev/tracer}) and executable (\url{http://beast.community/tracer}) forms.
-This latter page also serves up self-contained, step-by-step tutorials covering basic to advanced usage of \tracer\ to summarise posteriors under a variety of phylogenetic models using BEAST.
+This latter page also serves up self-contained, step-by-step tutorials covering basic to advanced usage of \tracer\ to summarise posteriors under a variety of phylogenetic models using BEAST and diagnose MCMC chain convergence.
 Popular tutorials employ \tracer\ to generate marginal parameter summaries and to infer population dynamics trajectories over time.

### 2018-04-12 manuscript/response.md
GR diagnostics
fix figure 3

@@
 
+TODO (by GB)
+
 *L108: I am not sure I understand the comparison (“model complexity”) being made between the marginal and joint marginal plots.
@@
 
+Model complexity does refer to the number of non-zero parameters.
+We highlight this connection through the sentence "With approximately equal numbers of transition rates, both figures suggest similar host and location trait model complexity."
+
 *L130: coda also offers diagnostics absent from Tracer, such as Gelman-Rubin statistics.*
+
+We now acknowledge this limitation in the text and introduce the Gelman and Rubin (1992) diagnostic.
 
@@
 *Figure 3 has no axis labels of any kind*
+
+Fixed.
 

### 2018-04-12 manuscript/tracer17_systbiol.tex
GR diagnostics
fix figure 3
adding replacement figure 1a

@@
 \begin{figure}[t]
-\subfigure[Tracer overview]{\hspace{0.40cm}\includegraphics[width=0.432\textwidth]{./figures/rabv-overview-3.pdf}} \\
+\subfigure[Tracer overview]{\includegraphics[width=0.485\textwidth]{./figures/rabv-overview-3.pdf}} \\
 \subfigure[Boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-boxplots.pdf}}
@@
 \centerline{
-\includegraphics[width=0.46\textwidth]{./figures/rabv-skygrid.pdf}
+\includegraphics[width=0.46\textwidth]{./figures/rabv-skygrid2.pdf}
 }
@@
  `coda' \citep{plummer2006coda} provides some of the functionality of \tracer\ within the R programming environment, while `AWTY' \citep{nylander2007awty} and `RWTY' \citep{warren2017rwty} explore the convergence of the phylogenetic tree parameter itself across multiple MCMC runs.
-%These alternative packages focus on MCMC convergence diagnostics, such as computing Gelman-Rubin statistics across multiple chains
-%% AR - anything else? The MDS tree-space approaches of Hillis etc.?
+These alternative packages compute, for example, Gelman-Rubin diagnostics \citep{gelman1992inference} that \tracer\ currently does not provide.
 

### 2018-04-12 manuscript/response.md
more changes

@@
 
+The short answer is that these values are arbitrary and we have updated the manuscript to reflect this.
+The long answer is that (1) these are nice, round numbers that users are likely to remember and (2) are sufficient to provide relatively accurate estimates of posterior means.
+
 *Line 43: If I'm correct, then the file names don't need to be same but the parameter name. Could you please correct this.*
+
+Fixed.  The text now reads, "[i]f multiple trace files contain the same collection of parameters..."
 

### 2018-04-12 manuscript/tracer17_systbiol.tex
more changes

@@
 allows users to assess the number of effectively independent draws from the posterior distribution the trace represents (Figure \ref{fig:overview}a).
-Colour coding assists the user in determining potential MCMC mixing problems, with cut-off values at 100 and 200.
+Colour coding assists the user in determining potential MCMC mixing problems, with arbitrary cut-off values at 100 and 200.
 
@@
 Multiple trace files can be selected in a similar fashion to compare posterior samples between different replicates of an analysis.
-If multiple trace files have the same file name,  then a ``Combined'' trace appears automatically. %GB: or have the same collection of logged parameters
+If multiple trace files contain the same collection of parameters,  then a ``Combined'' trace appears automatically. %GB: or have the same collection of logged parameters
 % , which can be selected as well as the individual trace files.
@@
 \begin{figure}[t]
-\subfigure[Tracer overview]{\includegraphics[width=0.485\textwidth]{./figures/rabv-overview-3.pdf}} \\
+\subfigure[Tracer overview]{\hspace{0.40cm}\includegraphics[width=0.432\textwidth]{./figures/rabv-overview-3.pdf}} \\
 \subfigure[Boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-boxplots.pdf}}

### 2018-04-12 manuscript/tracer17_systbiol.tex
more comments

@@
 Under BSSVS, a posterior estimate of the parameter should only sample values from states where the indicator equals one.
-Discrete phylogeographic analyses frequently employ BSSVS due to the potentially large amount of transition rates that need to be estimated \citep{Lemey2009}, but this is also relevant when employing model averaging approaches such as for relaxed molecular clocks for example \citep{Li2012}.
+Discrete phylogeographic analyses frequently employ BSSVS due to the potentially large amount of transition rates that need to be estimated \citep{Lemey2009}, but this is also relevant when employing model averaging approaches, e.g., over relaxed molecular clocks \citep{Li2012}.
 %
@@
 %The currently
-Available models are constant size, exponential and logistic growth \citep{drummond2002estimating},
+Available models include, e.g., constant size, exponential and logistic growth \citep{drummond2002estimating},
 and the non-parametric Bayesian skyline \citep{drummond2005bayesian,heledDrummond2008}, skyride \citep{minin2008smooth} and skygrid \citep{gill2012improving}.
@@
  `coda' \citep{plummer2006coda} provides some of the functionality of \tracer\ within the R programming environment, while `AWTY' \citep{nylander2007awty} and `RWTY' \citep{warren2017rwty} explore the convergence of the phylogenetic tree parameter itself across multiple MCMC runs.
-These alternative packages compute, for example, Gelman-Rubin diagnostics \citep{gelman1992inference} that \tracer\ currently does not provide.
+These alternative packages compute, e.g., Gelman-Rubin diagnostics \citep{gelman1992inference} that \tracer\ currently does not provide.
 

### 2018-04-12 manuscript/response.md
more comments
ignore new title

@@
 
+We elect to keep our original title "Posterior summarisation in Bayesian phylogenetic using Tracer 1.7" as it is shorter and reads more naturally in the form (adjective-noun) -> conjunction -> (adjective-noun) -> conjunction -> (noun-modifier).
+
 *L68: the BSSVS description doesn’t really feel like it comes to a satisfactory conclusion.
@@
 
-TODO (by GB)
+Fixed.
 
@@
 
+The Reviewer is correct and we apologize for this oversight; many more parametric models are available.  This is the available list is quite long, we have modified the text to include an "e.g." cause.
+
 ### Feature suggestions (these are rather general and not necessary for the manuscript):
@@
 
-This has been implemented already under the File menu -> Export Data Table. ** This needs to be tested ***
+We thank the Reviewer for this suggestion.  It is implemented under the File menu -> Export Data Table.
 
@@
 
+We again thank the Reviewer for this bug alert and will attempt to fix in a future release.
 **Test loading a too large file.**
 
-# Editor's comments
+# Associate Editor's comments
 
@@
 
+Please see our response to Reviewer #1.
+
 *-l67: Something like the sentence starting on l71 could be moved here to indicate generally how the conditional distributions are constructed from the sampled parameter values.
@@
 *-l79: `Finally,...' Should this be a new paragraph?*
+
+Absolutely!

### 2018-04-12 manuscript/tracer17_systbiol.tex
ignore new title
move figures into same directory for submission

@@
 \begin{figure}[t]
-\subfigure[Tracer overview]{\includegraphics[width=0.485\textwidth]{./figures/rabv-overview-3.pdf}} \\
-\subfigure[Boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-boxplots.pdf}}
-\subfigure[Density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-density.pdf}}\\
-\subfigure[Violin]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-violin.pdf}}
-\subfigure[Trace]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-traces.pdf}}
+\subfigure[Tracer overview]{\includegraphics[width=0.485\textwidth]{figure_1a_rabv-overview.pdf}} \\
+\subfigure[Boxplot]{\includegraphics[width = .23\textwidth, height = 3cm]{figure_1b_rabv-boxplots.pdf}}
+\subfigure[Density]{\includegraphics[width = .23\textwidth, height = 3cm]{figure_1c_rabv-density.pdf}}\\
+\subfigure[Violin]{\includegraphics[width = .23\textwidth, height = 3cm]{figure_1d_rabv-violin.pdf}}
+\subfigure[Trace]{\includegraphics[width = .23\textwidth, height = 3cm]{figure_1e_rabv-traces.pdf}}
 \vspace{-0.25cm}
@@
 \begin{figure}[t]
-\subfigure[Bubble chart]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integerjointmarginal-2.pdf}}
-\subfigure[Marginal density]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-integermarginaldensity.pdf}} \\
-\subfigure[Scatter plot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-correlation.pdf}}
-\subfigure[Correlation plot]{\includegraphics[width = .23\textwidth, height = 3cm]{./figures/rabv-joint-real.pdf}}
+\subfigure[Bubble chart]{\includegraphics[width = .23\textwidth, height = 3cm]{figure_2a_rabv-integerjointmarginal.pdf}}
+\subfigure[Marginal density]{\includegraphics[width = .23\textwidth, height = 3cm]{figure_2b_rabv-integermarginaldensity.pdf}} \\
+\subfigure[Scatter plot]{\includegraphics[width = .23\textwidth, height = 3cm]{figure_2c_rabv-correlation.pdf}}
+\subfigure[Correlation plot]{\includegraphics[width = .23\textwidth, height = 3cm]{figure_2d_rabv-joint-real.pdf}}
 \vspace{-0.25cm}
-%\subfigure[Box-and-whisker plot]{\includegraphics[width = .465\textwidth, height = 3cm]{./figures/rabv-catreal.pdf}}
 \caption{Multi-parameter visualisations of: (a) the joint probability distribution of two integer variables through a bubble chart; (b) the marginal density of multiple integer or categorical variables through frequency plots; (c) two continuous variables through a classic scatter or correlation plot; (d) multiple ($> 2$) continuous variables using large correlation matrices.
-%; (e) box-and-whisker plots for the joint marginal distribution of a continuous and an integer or categorical variable.
 }
@@
 Discrete phylogeographic analyses frequently employ BSSVS due to the potentially large amount of transition rates that need to be estimated \citep{Lemey2009}, but this is also relevant when employing model averaging approaches, e.g., over relaxed molecular clocks \citep{Li2012}.
-%
+
 Finally, \tracer\ provides demographic reconstruction resulting in a graphical plot, often applied to reconstruct epidemic dynamics.
@@
 \centerline{
-\includegraphics[width=0.46\textwidth]{./figures/rabv-skygrid2.pdf}
+\includegraphics[width=0.46\textwidth]{figure_3_rabv-skygrid.pdf}
 }

### 2018-04-12 manuscript/response.md
remove commented out text

@@
 We further disagree that most people employ Tracer simply to diagnose convergence problems; we, for example, commonly use Tracer to quickly calculate summary statistics (posterior means, 95% highest probably density intervals).
-However, we agree with the Reviewer that readers should be aware that Tracer is useful in diagnosing convergence problems and have added associated text highlighting a new tutorial on convergence to the manuscript.
+However, we agree with the Reviewer that readers should be aware that Tracer is useful in diagnosing convergence problems and have added associated text to the manuscript highlighting a new tutorial on convergence.
 

### 2018-04-12 manuscript/tracer17_systbiol.tex
remove commented out text

@@
 
-\editor{Associate Editor: To fill in}
+\editor{Associate Editor: Edward Susko}
 
@@
 
-%A key feature of \tracer\ is its ability to link together samples from quantitative model parameter trace files with tree trace files. %% AR - only for the demographic reconstructions...
-%This linkage is not available in other popular post-processing MCMC software, such as coda \citep{plummer2006coda} and AWTY \citep{nylander2007awty}, and enables users to approximate the posterior distribution of scientifically relevant quantities that are functions of both the tree and other quantitative measures.
-%% AR - ...
-which are intended in these other software packages.
-%For each trace,
 \tracer\ examines the posterior samples from all the available parameters, treating continuous, integer and categorical parameters appropriately,
@@
 
-%\tracer\ is able to handle three trace types, also known as variable types \citep{mendenhall2012introduction}: real, integer, and categorical.
-%The real trace type is used to handle quantitative continuous variables, which commonly appear in phylodynamic analyses, whereas the integer trace type allows to deal with quantitative discrete variables and the categorical trace is meant for qualitative categorical variables.
-%The three trace types are automatically assigned when the trace file is imported but can also manually be set.
-
 Selecting multiple parameters from the ``Traces" panel on the left generates a side-by-side comparison or an overlay of the selected parameters' visualisations (Figure \ref{fig:overview}b-e).
 Multiple trace files can be selected in a similar fashion to compare posterior samples between different replicates of an analysis.
-If multiple trace files contain the same collection of parameters,  then a ``Combined'' trace appears automatically. %GB: or have the same collection of logged parameters
-% , which can be selected as well as the individual trace files.
+If multiple trace files contain the same collection of parameters,  then a ``Combined'' trace appears automatically.
 \tracer\ generates four display panels for the selected parameters:
@@
 Finally, \tracer\ provides demographic reconstruction resulting in a graphical plot, often applied to reconstruct epidemic dynamics.
-%Depending on the actual demographic model used in the analysis, an additional file containing samples from the posterior tree distribution will be required.
-%The currently
 Available models include, e.g., constant size, exponential and logistic growth \citep{drummond2002estimating},
@@
 The data set comprises 372 \textit{nucleoprotein} gene sequences from 17 bat species, sampled between 1997 and 2006 across 14 states in the United States \citep{Streicker,Faria2013}.
-%We estimate the ancestral locations of RABV and their host-jumping history using a Bayesian discrete phylogeographic approach with BSSVS \citep{Lemey2009}, while simultaneously estimating RABV effective population sizes over time through a Bayesian skygrid coalescent model \citep{gill2012improving}.
 We estimate RABV ancestral locations and host-jumping history using a Bayesian discrete phylogeographic approach with BSSVS, while simultaneously estimating effective population sizes over time through a Bayesian skygrid coalescent model \citep{gill2012improving}.
@@
 Phylogeographic BSSVS inference includes parameters of both integer (number of non-zero transition rates) and categorical (host or location-state) trace types.
-%While a single parameter of such types can easily be visualised using a frequency plot,
 In \tracer, a bubble chart visualises the joint probability distribution between two integer or categorical traces (see Figure \ref{fig:4tabs}a).
-Circle area is proportional to the joint probability, %the blue coloured circle is in the credible set of given a probability threshold default to 0.95, the red is in the set not in the credible set.
+Circle area is proportional to the joint probability,
 with a coloured tile background if this probability reaches a nominal threshold to enhance visibility.
-%The coloured background can also help to show the area of the credible set and non-credible set.
-%Multiple integer parameters can also be visualised using
 Marginal density plots can also display multiple integer parameters, each with unique colour scales (see Figure \ref{fig:4tabs}b).
 With approximately equal numbers of transition rates, both figures suggest similar host and location trait model complexity.
-%Both figures show a clear overlap in the number of non-zero transition rates in the discrete trait models for host and location, as identified using a BSSVS procedure.
-\tracer\ also provides
-%The visualisation options for integer or categorical parameters are complemented with
- popular visualisations for continuous parameters, including scatter plots for two parameters (see Figure \ref{fig:4tabs}c),
+\tracer\ also provides popular visualisations for continuous parameters, including scatter plots for two parameters (see Figure \ref{fig:4tabs}c),
  and extensions for correlations between $\ge\hspace{-0.08em}2$ continuous parameters (Figure \ref{fig:4tabs}d; \citet{Murdoch}).
- %, which indicates a strong positive correlation between the clock rates of both discrete trait models.
-%We also provide an extension of such a scatter plot for multiple (i.e.
-more than two) continuous variables (see Figure \ref{fig:4tabs}d; \citet{Murdoch}), which makes use of coloured ellipses to indicate the strength of correlation between each pair of continuous variables.
 Colour gradients indicate strength and direction of the correlation, from red (strong negative) to blue (strong positive).
-%with a colour gradient from white to dark blue indicating increasingly strong correlation and a colour gradient from white to dark red indicating increasingly strong anti-correlation.
 Ellipse shapes re-enforce the strength of correlation, with no correlation appearing as a circle and perfect (anti-)correlation as a line.
-%Finally, in order to display the joint marginal distribution between a continuous variable and an integer or categorical variable, we make use of box-and-whisker plots.
-%Using such a plot (Figure \ref{fig:4tabs}e), we can infer that the range of the non-zero transition rates can depend on the ancestral location state inferred at the root of the underlying phylogeny.
 
 \tracer\ reconstructs the demographic history of RABV by drawing the effective population sizes over time (Figure \ref{fig:rabv}).
-%Given the properties of this coalescent prior, which requires to specify a grid of intervals over the timeframe of the epidemic, \tracer\ is able to infer this history without requiring a posterior sample of trees.
-%According to \tracer's skygrid plot, it is clear that
 RABV has successfully established itself in North American bat species, with its effective population size rising steadily throughout recent centuries.
 Following a rapid decline at the end of last century, we observe a recent sharp increase in size.
-%, reflecting [INSERT SOMETHING].
 

### 2018-04-12 manuscript/response.md
briefly handle condition distributions

@@
 
+We agree that our original description of visualising conditional posteriors was confusing, chiefly because we failed to provide instructions on how a user can generate the plots.
+We have rectified this short-coming by adding these instructions.
+The beginning of the paragraph now reads:
+
+"Tracer offers a solution of visualising conditional posterior distributions as well.
+Selecting one continuous and one categorical parameter generates side-by-side violin or boxplots under the Joint-Marginal panel.
+These plots present the continuous parameter distribution conditioned on the observed categorical values.
+A typical use case involves Bayesian stochastic search variable selection..."
+
+The remainder of the paragraph provides several use-cases, in line with the Associate Editor's suggestion.
+
 *-Fig 1(b)-(d): If you want to make space for additions, two of these could be cut and simply described as a list of (very familiar) available graphics for comparing multiple (marginal) distributions.

### 2018-04-12 manuscript/response.md
move around response

@@
 
-**To do**
+We agree that our original description of visualising conditional posteriors was confusing, chiefly because we failed to provide instructions on how a user can generate the plots.
+We have rectified this short-coming by adding these instructions.
+The beginning of the paragraph now reads:
+
+"Tracer offers a solution of visualising conditional posterior distributions as well.
+Selecting one continuous and one categorical parameter generates side-by-side violin or boxplots under the Joint-Marginal panel.
+These plots present the continuous parameter distribution conditioned on the observed categorical values.
+A typical use case involves Bayesian stochastic search variable selection..."
 
@@
 
-We agree that our original description of visualising conditional posteriors was confusing, chiefly because we failed to provide instructions on how a user can generate the plots.
-We have rectified this short-coming by adding these instructions.
-The beginning of the paragraph now reads:
-
-"Tracer offers a solution of visualising conditional posterior distributions as well.
-Selecting one continuous and one categorical parameter generates side-by-side violin or boxplots under the Joint-Marginal panel.
-These plots present the continuous parameter distribution conditioned on the observed categorical values.
-A typical use case involves Bayesian stochastic search variable selection..."
-
-The remainder of the paragraph provides several use-cases, in line with the Associate Editor's suggestion.
+Please see our response to Reviewer #1 (S1).  The remainder of the quoted paragraph provides several use-cases, in line with the Associate Editor's suggestion.
 
@@
 
-Please see our response to Reviewer #1.
+Please see our response to Reviewer #1 (P1).
 

### 2018-04-14 src/tracer/traces/ContinuousDensityPanel.java
Increasing max heap memory configuration.

@@ -233,7 +233,7 @@ public class ContinuousDensityPanel extends TraceChartPanel {
     }
 
     protected Plot createHistogramPlot(List values) {
-        return new NumericalDensityPlot(values, currentSettings.minimumBins);
+        return new NumericalDensityPlot(values, getSettings().minimumBins);
     }
 
     protected Plot createKDEPlot(List values) {

### 2018-04-14 manuscript/response.md

@@
 We have rectified this short-coming by adding these instructions.
-The beginning of the paragraph now reads:
+The paragraph now reads:
 
@@
 These plots present the continuous parameter distribution conditioned on the observed categorical values.
-A typical use case involves Bayesian stochastic search variable selection..."
+A typical use case involves Bayesian stochastic search variable selection (BSSVS), a form of model averaging, in which parameters influence the likelihood function only when a specific model is selected by a random indicator function."
 
@@
 At least, there should be a link in the paper to some material giving such an example.
-The current Figures 1 b-e and 2a-c feel a little generic-- we know what these plot types look like.
-*
+The current Figures 1 b-e and 2a-c feel a little generic-- we know what these plot types look like.*
 
@@
 Is the sentence at line 80 a continuation of this thought, or starting something new?*
+
+As stated in a previous response, we have rewritten this paragraph to better describe the option in Tracer to visualise conditional posteriors.
+The sentence at line 80, starting a section on demographic plot generation/visualisation, has been moved to a new paragraph.
 
@@
 
-We again thank the Reviewer for this bug alert and will attempt to fix in a future release.
-**Test loading a too large file.**
+We again thank the Reviewer for this bug alert, which is related to the default memory allocation in Tracer.
+We have increased this default substantially, to decrease the odds the Tracer users will encounter this problem.
+Should users still encounter this problem, they can increase memory allocation as well, which we now discuss in one of the Tracer tutorials.
 
@@
 
+As stated in a previous response, we now provide proper instructions on how to visualise conditional posteriors.
+
 *-l79: `Finally,...' Should this be a new paragraph?*

### 2018-04-14 manuscript/tracer17_systbiol.tex
briefly handle condition distributions
conditional plots can use integers as well.

@@
 $^{3}$Centre for Computational Evolution, University of Auckland, Auckland, NZ;
-$^{4}$Department of Microbiology and Immunology, Rega Institute, KU Leuven, Leuven, Belgium;
+$^{4}$Department of Microbiology and Immunology, Rega Institute, KU Leuven - University of Leuven, Leuven, Belgium;
 $^{5}$Department of Human Genetics, University of California, Los Angeles, USA;
@@
 
-\tracer\ examines the posterior samples from all the available parameters, treating continuous, integer and categorical parameters appropriately,
-from a trace
-and presents statistical summaries and visualizations.
+\tracer\ examines the posterior samples from all the available parameters - treating continuous, integer and categorical parameters appropriately - from a trace and presents statistical summaries and visualisations.
 Further, \tracer\ can analyse a single trace or combine samples from multiple files.
@@
 \vspace{-0.25cm}
-\caption{Overview of \tracer\ functionality and individual parameter visualisations: (a) Main \tracer\ panel upon loading a single trace file; (b) boxplot representation of two continuous parameters; (c) kernel density estimates; (d) violin plots; (e) the actual traces connecting the parameter values visited by the Markov chain.}
+\caption{Overview of \tracer\ functionality and individual parameter visualisations: (a) Main \tracer\ panel upon loading a single trace file; (b) boxplot representation of two continuous parameters; corresponding (c) kernel density estimates; (d) violin plots; and (e) the actual traces connecting the parameter values visited by the Markov chain.}
 \label{fig:overview}
@@
 \tracer\ offers a solution of visualising conditional posterior distributions as well.
+Selecting one continuous and one integer or categorical parameter generates side-by-side violin or boxplots under the Joint-Marginal panel.
+These plots present the continuous parameter distribution conditioned on the unique integer or categorical values.
 A typical use case involves Bayesian stochastic search variable selection (BSSVS), a form of model averaging, in which parameters influence the likelihood function only when a specific model is selected by a random indicator function.
